### 2016-05-10 Paper_v1.tex
setting up latex files
basic outline of manuscript with Figure 1 included

\documentclass[12pt]{article}
% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{\arg\!\min}

\usepackage{algorithm2e}
\usepackage{algpseudocode}
%\usepackage[space]{cite}
\urlstyle{rm}

%\textwidth = 6.50 in
%\textheight = 9.5 in
%\oddsidemargin =  0.0 in
%\evensidemargin = 0.0 in
%\topmargin = -0.50 in
%\headheight = 0.0 in
%\headsep = 0.25 in
%\parskip = 0.15in
%\linespread{1.75}
\doublespace

%\bibliographystyle{chicago}
\bibliographystyle{plos2009}

\makeatletter
\renewcommand\subsection{\@startsection
	{subsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\bfseries}}
\renewcommand\subsubsection{\@startsection
	{subsubsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}
\renewcommand\section{\@startsection
	{subsection}{2}{0mm}
	{-0.2in}
	{0.05\baselineskip}
	{\normalfont\large\bfseries}}
\renewcommand\paragraph{\@startsection
	{paragraph}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}
\makeatother

%Review style settings
%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings

% Single space'd bib -
\setlength\bibsep{0pt}

\renewcommand{\rmdefault}{phv}\renewcommand{\sfdefault}{phv}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

% Change the number format in the ref list -
\renewcommand{\bibnumfmt}[1]{#1.}

% Change Figure to Fig.
\renewcommand{\figurename}{Fig.}

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Dynamic Modeling of Human Complement System using Reduced Ordered Models}}}
\vspace{0.05in}
{\par \centering \large{Adithya Sagar, Wei Dai, Mason Minor, and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {School of Chemical and Biomolecular Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~Dynamic Modeling of Human Complement System using Reduced Ordered Models}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{???????}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Professor, School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Fill me in.

%We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements. Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems. For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.

\vspace{0.1in}
{\noindent \textbf{Keywords:}~Biochemical engineering, systems biology, reduced order models, complement system}

% Extra abstract
% Mathematical modeling of biological systems with multiple feed back loops is one such area where parameter estimation is a difficult non-linear optimization problem. This difficulty is further compounded when dealing with parameter vectors of high dimensions.

%In this study, we present the dynamically dimensioned particle
%a novel meta-heuristic approach that combines a variant of particle swarm optimization (PSO) along with dynamically dimensioned search (DDS) to obtain near optimal solutions of high dimensional biochemical networks within a relatively few function evaluations.
%We use a particle swarm optimization technique that uses multi-swarms to generate candidate vectors which are then greedily updated using DDS by dynamically varying the perturbed parameter dimensions. We tested this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and also against DDS alone. The new algorithm outperforms all the other meta-heuristics on the coagulation model. The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data. We also performed this comparison on commonly used test functions (Ackley and Rosenbrock) for global optimization and found the same behavior. Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach. We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.

\pagebreak

\setcounter{page}{1}

% Uncomment in production -
\linenumbers

\section*{Introduction}
The introduction has three paragraphs (introduction no longer than 3 pages):
\begin{enumerate}
	\item{\textbf{First~paragraph}: Introduce human complement system, history, role in adaptive/innate immunity. }
	\item{\textbf{Second~paragraph}: Introduce mathematical models of complement system, current place in the field, our work was not possible without xyz who pioneered abc. Address shortcomings in the field .}
	\item{\textbf{Third~paragraph}: In this study, [Repeat the abstract with some additional detail]. Taken together, [killer statement].}
\end{enumerate}

% Extra text from the introduction -
%With the capacity to construct very large models using various mathematical formalisms \cite{chen2009input,tasseff2011modeling,luan2007computationally,mo2007genome,orth2011comprehensive,karr2012whole,buchel2013path2models,smallbone2010towards}, more often than not,

%this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics  The new algorithm outperforms all the other meta-heuristics on the coagulation model. The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data. We also performed this comparison on  and found the same behavior. Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach. We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.
%took into cognizance the fact that it may not be necessary to obtain the exact solution for high dimensional biological systems and that good enough solutions can be quickly obtained without expending a lot of objective function evaluations. Hence our current approach uses the power of population based heuristics along with DDS to obtain near optimal or good solutions. Though in theory we can combine any meta-heuristic with DDS, for the current purpose we used Particle Swarm Optimization (PSO). PSO unlike Genetic Algorithm (GA) or Differential Evolution (DE) does not have complex operations like cross over, mutation or recombination. It is simple to use and does not have a lot of parameters associated with other heuristics. However PSO is known to rapidly converge to a local optimum and thus several variants \cite{peer2003using,zhan2009adaptive,li2007fast} have been developed of which Multi swarm particle swarm optimization approaches (MLSPSO) \cite{zhao2008dynamic,liang2005dynamic} are one of the effective ones. We used MLSPSO to preclude bad regions of search and thereafter search within these regions using DDS.

\clearpage

\section*{Results}

The results are presented in \textbf{past~tense}. Each paragraph starts with a statement of the result in that paragraph in active voice.
Each results paragraph ends with a Taken together type statement followed by a link statement e.g., Next we considered etc. When referring to figures, state what the figures shows (Fig. ZZ).

\clearpage

%As the dimensionality of  increases, the search region gets widened and thus the problem becomes more challenging.
%

\section*{Discussion}

The discussion has three (sometimes four) paragraphs:
\begin{enumerate}
	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
	In this case, we will have a scaling issue if we extend to genome scale. We should extend to dynamic cases, and we need to experimentally validate the findings. }
\end{enumerate}

\clearpage

\section*{Materials and Methods}

\begin{enumerate}
	\item{\textbf{Model~formulation}: Present the reduced order modeling approach of the human complement network, . Outline parameter estimation, and state all parameter assumptions, sensitivity analysis}
\end{enumerate}

\clearpage

%Need to talk more about biochemical benefits and importance for of biochemical problems

\section*{Acknowledgements}
This study was supported by an award from [FILL ME IN].
\clearpage

\bibliography{References_v1}

\clearpage

% Figures and captions go here ...
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.00\textwidth]{./figs/<Filename>.pdf}
%\caption{Captiontext goes here}
%}\label{fig:<label_name>}
%\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./Figures/ComplementSystemSchematic_v1.pdf}
\caption{The biochemical reactions of the complement system.
}\label{fig1-schematic}
\end{figure}

\clearpage

% Supplemental figures -
% Set the S-
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{T\arabic{table}}
\renewcommand\thepage{S-\arabic{page}}
\renewcommand\theequation{S\arabic{equation}}

% Reset the counters -
\setcounter{equation}{0}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{page}{1}

% Supplemental figures go here ...
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.00\textwidth]{./figs/<Filename>.pdf}
%\caption{Captiontext goes here}
%}\label{fig:<label_name>}
%\end{figure}

\end{document}

### 2016-05-14 Paper.tex
Figures and Figure captions in progress
added figure captions for Fig 1-3, working on 4

\documentclass[12pt]{article}
% Load packages
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
\usepackage[round,comma,sort&compress]{natbib}
\usepackage{supertabular}
\usepackage{multirow}

\urlstyle{rm}

% set the size of the system -
% set the size of the system -
% Set the size
\textwidth = 6.50 in
\textheight = 9.5 in
\oddsidemargin =  0.0 in
\evensidemargin = 0.0 in
\topmargin = -0.50 in
\headheight = 0.0 in
\headsep = 0.25 in
%\parskip = 0.15in
%\linespread{1.75}
\doublespace

\bibliographystyle{plos2009}

\makeatletter
\renewcommand\subsection{\@startsection
	{subsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\bfseries}}
\renewcommand\subsubsection{\@startsection
	{subsubsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}	
\renewcommand\paragraph{\@startsection
	{paragraph}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}
\makeatother

%Review style settings
%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings

% Single space'd bib -
\setlength\bibsep{0pt}

\renewcommand{\rmdefault}{phv}\renewcommand{\sfdefault}{phv}

% Change the number format in the ref list -
\renewcommand{\bibnumfmt}[1]{#1.}

% Change Figure to Fig.
\renewcommand{\figurename}{Fig.}

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\large Analysis of the Human Alternative Complement Pathway Architecture Identified Globally Fragile and Robust Components.}}
\vspace{0.05in}
{\par \centering Adithya Sagar and Jeffrey D. Varner$^{*}$}
\vspace{0.05in}
{\par \centering School of Chemical and Biomolecular Engineering}
{\par \centering Cornell University, Ithaca NY 14853}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~Analysis of the alternative complement cascade}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~PLoS~ONE}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Associate Professor, School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853} 
{\par \centering Email: jdv27@cornell.edu} 
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}

{\noindent \textbf{Keywords:}~Complement activation, Systems~biology}

\pagebreak

\setcounter{page}{1}
\section*{Introduction}
Complement is the central subsystem of innate immunity activated as part of host defense.
Complement destroys invading microorganisms and eliminates immune complexes \cite{WALPORT1}.
Complement can also be activated during trauma and is correlated with negative clinical outcomes [REFME].
Complement, discovered more than a century ago [REFME], consists of more than 30 plasma and cell surface proteins and is organized into three pathways, 
the classical, lectin binding and alternative pathways \cite{WALPORT1}.
In this study, we explored only the alternative activation pathway. 
The alternative pathway is considered to be the primary amplification mechanism for complement activation [REFME].
The alternative pathway consists of complement proteins numbered C3-C9, plasma factors B,~D,~H and I and several regulatory proteins restraining activation. 
C3 undergoes spontaneous hydrolysis in plasma to form a reactive protein intermediate C3(H$_2$O). 
C3(H$_2$O) is almost immediately inactivated by the regulatory proteins Factor H and I, however, before inactivation C3(H$_2$O) can form a complex with Factor B thereby protecting itself against Factor H/I mediated inactivation. The C3(H$_2$O)-Factor B complex is cleaved by Factor D to form the fluid phase C3-convertase, denoted as C3Bb. C3Bb cleaves a 9-kDa fragment from the amino terminus of the $\alpha-$chain of free C3 to form two fragments, C3b and C3a. The C3b fragment tags foreign surfaces for destruction by the Membrane Attack
Complex (MAC). Surface bound C3b binds circulating Factor B, forming a complex which is cleaved by Factor D to produce the surface C3-convertase, denoted by C3bBb.
Fluid-phase and surface  C3-converatase amplifies the complement response; C3Bb and C3bBb cleave C3 to form C3b and C3a. Binding of
an addition C3b by C3bBb yields the C5-converatase, denoted as C3bBbC3b. The C5-convertase cleaves circulating C5 to form C5a and C5b fragments; C5b anchors itself on
hydrophobic regions of the invading surface. C5b binding is the first step in the formation of the MAC; surface bound C5b sequentially binds C6 and C7 forming the
C5b67 complex. C8 binds C5b67 via the $\beta-$chain of C5b. Binding of C8 triggers the polymerization of C9 proteins and the insertion of the poly-C9 structure
into the membrane of the invading pathogen. Insertion of the poly-C9 structure forms a pore that disrupts membrane integrity resulting in an influx of ions
and water into the cell leading to cell death and lysis \cite{WALPORT2}.

In this study, we developed and analyzed a mechanistic mathematical model of the alternative complement pathway. 
The objective of our analysis was to better understand the fragility and robustness of the components making up C3-converatase amplification loop.
The central computational challenge was the identification of model parameters.

\section*{Results}

\subsection*{Identification of an ensemble of alternative complement models.} 
We assembled a canonical model of the molecular architecture of the alternative complement pathway (Fig. Z and given in detail in Table Z) from primary literature.
The model, which described 53 molecular species interconnected by 104 interactions, was encoded as a system of nonlinear ordinary differential equations.
The rate of each of the 104 interactions contained in the model was modeled using elementary mass action kinetics; thus, each rate equation had a single rate constant.
A master set of model parameters (rate constants and initial conditions) was taken from literature, 
or adjusted to make model simulations \emph{qualitatively} consistent with observations. 
The master parameter set was used to generate synthetic data from which we identified a low-correlation ensemble of model parameters using Particle Swarm Optimization (PSO).

First-order sensitivity coefficients were calculated using the ODE15s routine of Matlab (The Mathworks, Natick MA)
over a family of random parameter sets ($N=100$) constructed by perturbing each nominal parameter by up $\pm${50}\%. The ODE15s solutions
were used as high-order approximations of the correct senstivity solutiuon. Overall State Sensitivity Coefficients (OSSC) calculated using the ODE15s solutions were compared 
with sensitvity coefficients calculated using four additional methods over the same random parameter family (see Methods); Low-order Foward Finite Difference, FFD; 
low-order Backward Difference Formulation, BDF; high-order Interpolating Collocation, Collocation and a high-order Approximate Spectral Expansion, ASE. 
Complement mechanisms were rank-ordered based upon their OSSC values generated by each method. The Spearman Rank Correlation was used to quantify the similarity 
between each of the rank-ordered lists and the ordering produced by the ODE15s method. Fixed-step methods produced results that were not consistent with
ODE15s and required several iterations to correctly identify the proper perurbation size. The conclusions drawn from fixed-step methods were also found to be sensitive to model
stiffness; both the classification of mechanisms as fragile and robust and the identifiablity of model parameters were effected by method choice.

\section*{Discussion}

\clearpage

\section*{Materials and Methods}
\subsection*{Formulation and solution of model equations.}
Mass balance equations were written around each protein or protein complex in the complement model yielding the set of nonlinear ordinary differential equations (vector-form):
\begin{equation}\label{eq-compact-form}
\frac{d\mathbf{x}}{dt}=\mathbf{S}\mathbf{r}\left(\mathbf{x},\mathbf{k}\right)\qquad\mathbf{x}\left(t_{o}\right)=\mathbf{x}_{o}
\end{equation}where $\mathbf{x}$ denotes the abundance vector (54$\times$1),
and $\mathbf{r}\left(\mathbf{x},\mathbf{k}\right)$ denotes the vector of reaction rates (104$\times$1).
The matrix $\mathbf{S}$, which denotes the stoichiometric matrix (53$\times$104), describes the biological connectivity of the model. 
Each row in $\mathbf{S}$ describes a particular protein or protein complex while each column describes the stoichiometry associated with a specific interaction in the network. 
The $(i,j)$ element of $\mathbf{S}$, denoted by $\sigma_{ij}$, describes how protein $i$ is connected to rate process $j$. 
If $\sigma_{ij}<0$, then protein $i$ is consumed in $r_{j}$, conversely, if $\sigma_{ij}>0$ protein $i$ is produced by $r_{j}$ and if $\sigma_{ij}=0$ 
there is no connection between protein $i$ and rate process $j$.
We have assumed mass action kinetics for each interaction in Table [RXNTABLE]. 
For the general reaction $q$:
\begin{equation}
\sum_{j\in\left\{\mathbf{R}_{q}\right\}}\sigma_{jq}x_{j}\rightarrow\sum_{k\in\left\{\mathbf{P}_{q}\right\}}\sigma_{kq}x_{k}
\end{equation}is given by:
\begin{equation}\label{eq-mass-action}
r_{q}\left(\mathbf{x},k_{q}\right)=k_{q}\prod_{j\in\left\{\mathbf{R}_{q}\right\}}x_{j}^{-\sigma_{jq}}
\end{equation}where $\left\{\mathbf{R}_{q}\right\}$ denotes the set of reactants for reaction $q$, $\left\{\mathbf{P}_{q}\right\}$
denotes the product set for reaction $q$, $k_{q}$ denotes the rate constant governing the qth reaction 
and $\sigma_{jq},\sigma_{kq}$ denote stoichiometric coefficients (elements of the matrix $\mathbf{S}$). 
We have treated every rate as non-negative; all reversible reactions were split into two irreversible reaction steps. 
Thus, every element of the reaction rate vector $\mathbf{r}\left(\mathbf{x},\mathbf{k}\right)$ takes the form shown in Eq. \eqref{eq-mass-action}.

\subsection*{Sensitivity and robustness analysis of the complement model population.}
Sensitivity coefficients were calculated over the model ensemble (N = 500). 
First-order sensitivity coefficients at time $t_{q}$:  
\begin{equation}\label{eqn_sen}
s_{ij}\left(t_{q}\right)=\frac{\partial{x_{i}}}{\partial{k_{j}}}\Biggr|_{t_{q}}
\end{equation}
were computed by solving the kinetic-sensitivity equations \cite{dickinson1976sensitivity}:
\begin{equation}\label{eqn_extended}
\begin{pmatrix}
d{\mathbf{x}}/dt\\
d{\mathbf{s}}_{j}/dt
\end{pmatrix}=
\begin{bmatrix}
\mathbf{S}\cdot\mathbf{r}\left(\mathbf{x},\mathbf{k}\right) \\
\mathbf{A}\left(t\right)\mathbf{s}_{j}+\mathbf{b}_{j}\left(t\right)
\end{bmatrix}\qquad j=1,2,\hdots,P
\end{equation}
subject to the initial condition $\textbf{s}_{j}(t_{0}) = \textbf{0}$.
The quantity $j$ denotes the parameter index, $P$ denotes the number of parameters in the model, $\mathbf{A}$ denotes the Jacobian matrix, and $\mathbf{b}_j$ denotes the $j$th column of the matrix of first-derivatives of the mass balances with respect to the parameters.
Sensitivity coefficients were calculated by repeatedly solving the extended kinetic-sensitivity system for forty parameters sets selected from the final 400 member ensemble. 
These sets were chosen to be comparable to the final 400 member ensemble on the basis of parametric coefficient of variation (CV); the sets selected for sensitivity analysis had a 
mean CV of 0.85 $\pm$ 0.5 and a mean correlation of approximately 0.6. Thus, there were diverse and uncorrelated.
The Jacobian $\mathbf{A}$ and the $\mathbf{b}_{j}$ vector were calculated at each time step using their analytical expressions generated by UNIVERSAL.

The resulting sensitivity coefficients were scaled and time-averaged (Trapezoid rule):
\begin{equation}\label{eqn_INsen}
	\mathcal{N}_{ij} \equiv \frac{1}{T}\int^{T}_{0} dt \cdot |s_{ij}(t)|
\end{equation}
where $T$ denotes the final simulation time.
The time-averaged sensitivity coefficients were then organized into an array for each ensemble member:
\begin{equation}
	\mathcal{N}^{\left(\epsilon\right)} = 
	\begin{pmatrix}
		\mathcal{N}_{11}^{\left(\epsilon\right)} & \mathcal{N}_{12}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{1j}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{1P}^{\left(\epsilon\right)} \\
		\mathcal{N}_{21}^{\left(\epsilon\right)} & \mathcal{N}_{22}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{2j}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{2P}^{\left(\epsilon\right)} \\
		\vdots & \vdots & & \vdots & & \vdots \\
		\mathcal{N}_{M1}^{\left(\epsilon\right)} & \mathcal{N}_{M2}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{Mj}^{\left(\epsilon\right)} & \hdots & \mathcal{N}_{MP}^{\left(\epsilon\right)} \\
	\end{pmatrix}\qquad\epsilon = 1,2,\hdots,N_{\epsilon}
\end{equation}
where $\epsilon$ denotes the index of the ensemble member, $P$ denotes the number of parameters, 
$N_{\epsilon}$ denotes the number of ensemble samples and $M$ denotes the number of model species.
To estimate the relative fragility or robustness of species and reactions in the network, we decomposed the $\mathcal{N}^{\left(\epsilon\right)}$ matrix 
using Singular Value Decomposition (SVD):
\begin{equation}
	\mathcal{N}^{\left(\epsilon\right)} = \mathbf{U}^{\left(\epsilon\right)}\Sigma^{\left(\epsilon\right)}\mathbf{V}^{T,{\left(\epsilon\right)}}
\end{equation}
Coefficients of the left (right) singular vectors corresponding to largest $\theta\leq{15}$ singular values of $\mathcal{N}^{\left(\epsilon\right)}$ 
were rank-ordered to estimate important species (reaction) combinations. Only coefficients with magnitude greater than a threshold ($\delta$ = 0.001) were considered. 
The fraction of the $\theta$ vectors in which a reaction or species index occurred was used to determine its importance (sensitivity ranking). 
The sensitivity ranking was compared between different conditions to understand how control in the network shifted as a function of perturbation or time.

\section*{Acknowledgement}

\bibliography{Paper}
\end{document}

### 2016-05-14 Paper_v1.tex
Figures and Figure captions in progress
added figure captions for Fig 1-3, working on 4

@@
 \centering
-\includegraphics[width=1.0\textwidth]{./Figures/ComplementSystemSchematic_v1.pdf}
-\caption{The biochemical reactions of the complement system.
-}\label{fig1-schematic}
+\includegraphics[width=1.0\textwidth]{./Figures/ComplementSystemSchematic_v2.pdf}
+\caption{Simplified schematic representation of the human complement system.
+The complement cascade is activated through any one, or more, of the three pathways: classical, lectin, and alternate pathway.
+The classical pathway is activated by the complex formation of $C1q$, $C1r$, and $C1s$ by the recogniztion of antibody:antigen complexes.
+Similarly, the lecin pathway is initiation by binding mannan-binding lectin to mannose on pathogen surfaces.
+Lastly, the alternative pathway is activated when a complement component is spontaneously bound to the surface of the pathogen of virus.
+The activation from the three pathways creates a cascades of reactions that forms the proteases, $C3$ Convertase that cleaves $C3$ into $C3a$, and $C3b$, the main effector molecule of the complement system.
+$C3b$ can find to a $C3$ convertase and form a $C5$ convertase that cleaves $C5$ into $C5a$, and $C5b$ that undergoes a series of reactions to form the membrane attach complex ($MAC$).}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./Figures/Figure2_Fit.pdf}
+\caption{Reduced order complement model training simulation for lectin and alternative pathway in presence of zymosan.
+Reduced order complement model parameters were estimated using dynamically dimensioned search (DDS) [Tolson and Shoemaker,2007,WRR] using the availability of zymosan as a function of lectin pathway initiation.
+Only parameters that govern the behavior of alternative pathway were allowed to vary when zymosan was not present.
+Our model training was conducted in a hierarchal fashion where the alternate parameters were trained and then used and fixed in estimating the lectin parameters.
+The red line shows the best-fit parameter, the black lines denotes the simulated mean value of $C3a$ or $C5a$ for a 50 parameter set ensemble.
+The shaded region denotes the distribution of $C3a$ and $C5a$ of the ensemble.}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./Figures/Figure3_Predictions_2.pdf}
+\caption{Reduced order complement model predictions of lectin and alternative pathway in presence of zymosan.
+(A-F) Simulation of complement dynamics in the presence of zymosan were conducted for a range of trigger values ($0.1$, $0.01$, and $0.001$ grams of zymosan).
+The time-course profiles of $C3a$ and $C5a$ under three different zymosan concentrations were simulated using 50 ensembles of trained parameter sets against experimental data of Shaw et al [REF].
+The red curve represents the best fit parameter, grey shaded region denotes the prediction results from 50 ensembles of parameter sets, and the black curve is the mean of the ensemble.
+All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./Figures/Figure4_Sensitivity_Analysis_v2.pdf}
+\caption{Sensitivity Analysis}
 \end{figure}

### 2016-05-14 Paper_v1.tex
Updated Paper and Figures

@@
 \vspace{0.05in}
-{\par \centering \large{Adithya Sagar, Wei Dai, Mason Minor, and Jeffrey D. Varner$^{*}$}}
+{\par \centering \large{Adithya Sagar, Wei Dai$^{\#}$, Mason Minot$^{\#}$, and Jeffrey D. Varner$^{*}$}}
 \vspace{0.10in}
@@
 \vspace{0.5in}
+{\par \centering $^{\#}$ Denotes equal contribution}
 {\par \centering $^{*}$Corresponding author:}
@@
 \section*{Introduction}
-The introduction has three paragraphs (introduction no longer than 3 pages):
+%The introduction has three paragraphs (introduction no longer than 3 pages):
 \begin{enumerate}
 	\item{\textbf{First~paragraph}: Introduce human complement system, history, role in adaptive/innate immunity. }
+
+	Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
+Complement was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
+Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases.
+The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
+C5a acts as a bridge between innate and adaptive immunity and plays a very important role in regulating inflammation and coagulation.
+Complement activation takes places through three different pathways: the alternate, the classical and the lectin.
+Each of these pathways involves a different initiator signal that leads to the formation of a serine protease called C5 convertase which cleaves an inactive protein called C5 to form C5a and C5b.
+The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
+A multimeric protein complex C1 binds to the antigen-antibody complex and undergoes a conformational change.
+This activated complex cleaves proteins C4 and C2 to C4a, C4b, C2a and C2b respectively.
+C4a and C2b combine to form a protease C4bC2a also known as the classical C3 convertase.
+The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to the carbohydrates on the surfaces of bacterial pathogens.
+This bound complex in turn cleaves C4 and C2 and leads to the production of C4bC2a.
+The alternate pathway involves a 'tickover' mechanism in which a protein called C3 is hydrolyzed to form C3b.
+In presence of foreign pathogens C3b binds to these surfaces and recruits  additional factors called factor B and factor D that lead to the formation of alternate C3 convertase - C3bBb.
+The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
+The classical and alternate C3 convertases recruit C3, Factor B and Factor D to form classical C5 convertase (C4bC2aC3b) and alternate C5 convertase (C3bBbc3B) respectively.
+The C5 convertases then cleave C5 to form C5a and C5b respectively.
+	The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the MAC complex.
+The activation of complement and formation of C5a and MAC complex is regulated at different points through a number of plasma and host cell proteins.
+C4 binding protein (C4BP), ---- (NEEDS TO BE FILLED)
+
+	-- Stuff about importance of complement (NEEDS TO BE FILLED)
+
 	\item{\textbf{Second~paragraph}: Introduce mathematical models of complement system, current place in the field, our work was not possible without xyz who pioneered abc.
 Address shortcomings in the field .}
+
+	Given the complexity and importance of complement, developing mathematical models of complement are crucial to understanding its dynamics.
+Complement models have typically been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
+Hirayama  et al.
+(ref) used a system of linear ODEs to model the classical pathway of complement.
+Korotaevskiy and co-workers (ref) built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways.
+However both these studies involve no validation studies with experimental data.
+Liu et al used analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters.
+Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters (Ref).
+This model delineated the response of complement on a host cell and a foreign antigen.
+However, these previous models were largely based upon mechanistic knowledge.
+However given the complexity of complement and its interactions with other networks like coagulation, autonomous nervous system, adaptive immunity it is unfeasible and computationally expensive to build such large mechanistic models.
+In addition is much more difficult to experimentally interrogate the response of various complement proteins under different conditions.
+This also presents with the problem of estimation of a large number of parameters with little or no experimental data.
+Thus there exists a need to reduce the mechanistic complexity while capturing dynamics of complement accurately.
+
     \item{\textbf{Third~paragraph}: In this study, [Repeat the abstract with some additional detail]. Taken together, [killer statement].}
+
+    Towards this unmet need, we formulated a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes.
+This circumvents the need to have detailed mechanistic knowledge or use large number of ODEs to model complex dynamics.
+
 \end{enumerate}
@@
 
+\subsection*{2.1 Formulation of Reduced Order Complement Models}
+
+We developed a reduced order extrinsic human complement network consisting of the most crucial steps of the human complement system (Fig.
+\ref{fig-schematic}).
+The core of our model was based upon the experimental measurements of Morad and coworker's earlier work \cite{morad2015time}, we only consider the activation of complement system through the alternate and the lectin pathways.
+In doing so we aim to capture a complex biological phenomenon using a few simple ordinary differential equations.
+A trigger event initiates the lectin pathway in the presence of zymosan, which activates the cleavage of $C2$ and $C4$ into $C2a$ and $C2b$, and $C4a$ and $C4b$ respectively.
+Classical Pathway ($CP$) $C3 convertase$ ($C4aC2b$) is by the combination of $C4a$ and $C2b$, which catalyzes the cleavage of $C3$ into $C3a$ and $C3b$.
+Similarly, the activation of the alternative pathways happens through the spontaneous hydrolysis of $C3$ which facilitates the cleavage of $C3$.
+$C3b$ then could combine with with $C3$ to form alternate pathway ($AP$) $C3 convertase$.
+The both  versions of the $C3 convertase$ catalyzes the cleavage of $C3$ into $C3a$ and $C3b$, and $C3b$ can then combine with either $CP$ or $AP C3 convertase$ to form $C5 convertase$, $CP or AP$ respectively that is responsible for the cleavage of $C5$ to $C5a$ and $C5b$.
+Lectin pathway activation was approximated using a combination of saturation kinetics and Hill-like function control functions.These control coefficients then modified the rates of model processes at each time step.
+Hill-like transfer functions $0 \le \mathbf{f (Z)}  \le 1$ quantified the contribution of components upon a target process, in this study, $\mathbf{Z}$ represents the abundance of the initiator.
+Taken together, while the reduced order human complement model encodes significant biological complexity, it is highly compact (consisting of only 18 differential equations).
+Thus, it will serve as an excellent proof
+of principle example to study the reduction of a highly complex human subsystem.
+
+\subsection*{2.2 Identification of Model Parameters Using Dynamically Dimensioned Searched (DDS)}
+
+A critical challenge for any dynamic model is the estimation of kinetic parameters. We estimated
+kinetic and control parameters in a hierarchical fashion using  two $in~vitro$ time-series human complement data sets with and without zymosan present.
+The residual between simulation and experimental measurements were minimized  using dynamically dimensioned search (DDS).
+An initial parameter set was initialized with randomized kinetic and control parameters and allowed to search for parameter vectors that minimized the residual.
+Knowing that the kinetic and control parameters of the lectin pathway does not affect the dynamics of the alternate pathway, we used a hierarchal approach that estimated the parameters for the alternative pathway and lectin pathway separately.
+For the alternative pathway, we utilized the time-course experimental measurements of Morad and coworkers \cite{morad2015time} of $C3a$ and $C5a$ in the absence of zymosan and only allowed the alternative parameters to vary (Fig.
+\ref{fig-fit} A and B).
+The estimated alternate parameters was then fixed for the determination of lectin pathway parameters.
+The training for the lectin parameters, we used the experimental measurements of $C3a$ and $C5a$ in the presence of $1$ g of zymosan published by Morad et al \cite{morad2015time} (Fig.
+\ref{fig-fit} C and D).
+
+The reduced human complement model captured the behavior of the alternative and lectin pathways through the time-course abundance of $C3a$ and $C5a$ (Fig.
+\ref{fig-fit}).
+However we were not able to capture the curvature of the $C5a$ alternate (Fig.
+\ref{fig-fit}), the decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
+Taken together, the model identification results suggested that our reduced order approach could reproduce a panel of lectin pathway intiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
+However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
+
+\subsection*{2.3 Validation of the Reduced Order Human Complement Model}
+IN PROGRESS
+
+We tested the predictive power of the reduced order human complement model with validation data sets not used during model training.
+Six validation data sets were used, three for $C3a$ and $C5a$ respectively at different zymosan concentrations.
+All kinetic and control parameters were
+fixed for the validation simulations. The reduced order model predicted the $C3a$ and $C5a$ time-course profiles at a qualitative level (Fig. \ref{fig-prediction}).
+
+\subsection*{2.4 Global Sensitivity Analysis of the Reduced Order Human Complement Model}
+
+We conducted a Sobol's sensitivity analysis to estimate which parameters controlled the performance of the reduced order model.
+We calculated the sensitivity of the change in $C3a$ and $C5a$ profiles using the residuals between simulation and experimentally measured data for the cases of $0$ and $1 g$ zymosan (Fig.
+\ref{fig-SA}.
+For the cases in absence of zymosan where only the alternative pathway is active, we observed that only a few variables are responsible for the system response.
+For $C3a$ alternate, the sensitivity analysis found that $k_{c3b~basal}$ and $k_{degradationC3a}$ are the only sensitive parameters.
+This gives us new insight in which of the parameters play a role in complement activation.
+Even though $AP~C3~Convertase$ is also responsible in the conversion of $C3$ and the production of $C3a$, the kinetic parameters that govern the equation was not sensitive at all.
+This elucidated that the activation of alternative pathway is more heavily governed by the spontaneous hydrolysis of $C3$ rather than the activity of $AP~C3~Convertase$.
+Surprisingly, closely examining the sensitive parameters that control $C5a$, in addition to the expected kinetic and control parameters related to the formation of $AP~C5~Convertase$, we observed that $k_C3~Convertase2$, the was previously not sensitive to $C3a$, to be sensitive in the formation of $C5a$.
+The $AP~C3~Convertase$ is a substrate required for the formation of $AP~C5~Convertase$ and the formation of $C3b$.
+The change in activity of $AP C3 Convertase$  will not drastically change the $C3a$ dynamics, but will effect $AP~C5a~Convertase$ formation and $C5a$ formation.
+The our reduced order human complement model in combination with Sobol's sensitivity analysis was able to unravel important indirect parameter interaction.
+
+Our sensitivity analysis yielded expected results for the lectin pathway analyzes (Fig.
+\ref{fig-SA} (C and D)).
+One key difference that was observed between the sensitivity of the parameters between $C3a$ an $C5a$ was their respective degradation terms.
+The degradation constant of $C3a$ was sensitive between the two different cases of zymosan that was tested while the degradation constant of the $C5a$ was not sensitive.
+We believe this different is attributed to the magnitude of the parameters and their respective concentrations.
+
 \clearpage
@@
 \section*{Materials and Methods}
+We used ordinary differential equations (ODEs) to model the time evolution of proteins ($x_{i}$) in our reduced order complement model:
+\begin{equation}
+	\frac{dx_{i}}{dt}  =  \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
+\end{equation}
+where $\mathcal{R}$ denotes the number of reactions, $\mathcal{M}$ denotes the number of protein species in the model. 
+The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$. 
+Typically, reaction $j$ is a non-linear function of biochemical species abundance, as well as unknown kinetic parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$). 
+The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$. 
+If $\sigma_{ij}>0$, species $i$ is produced by reaction $j$. 
+Conversely, if $\sigma_{ij}<0$, species $i$ is consumed by reaction $j$, while $\sigma_{ij} = 0$ indicates species $i$ is not connected with reaction $j$. 
+The system material balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$, which were specified by the experimental setup and human serum complement protein levels.
 
-\begin{enumerate}
-	\item{\textbf{Model~formulation}: Present the reduced order modeling approach of the human complement network, .
-Outline parameter estimation, and state all parameter assumptions, sensitivity analysis}
-\end{enumerate}
+Each reaction rate was written as either the product of two terms, a kinetic term ($\bar{r}_{j}$) and a control term ($f_{j}$), or $\bar{r}_{j}$ for the case $f_{j}$ = 1:
+\begin{equation}\label{eqn:rate-factor}
+	r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right) = \bar{r}_{j}f_{j}
+\end{equation}
+
+Each control term, $f_{j}\left(\mathcal{Z}\right)$ took the form:
+\begin{equation}\label{eqn:control-factor}
+	f_{j}\left(\mathcal{Z}_{j},\alpha_{j}\right) = \frac{\mathcal{Z}_{j}^{\eta}}{\mathcal{Z}_{j}^{\eta} + \alpha_{j}^{\eta}}
+\end{equation}where $\mathcal{Z}_{j}$ denotes the abundance of the $j$ factor (e.g., initiator abundance),
+$\alpha_{j}$ denotes the species gain parameter, and $\eta$ a cooperativity parameter (similar to a Hill coefficient).
+
+We used saturation kinetics to model the lectin pathway activation and C3 and C5 convertase activity reaction terms $\bar{r}_{j}$:
+\begin{equation}\label{eqn:rate-saturation}
+	\bar{r}_{j} = k_{j}^{max}\epsilon_{i}\left(\frac{x_{s}^{\eta}}{K_{js}^{\eta} + x_{s}^{\eta}}\right)
+\end{equation}
+where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$, $\eta$ denotes a cooperativity parameter (similar to a Hill coefficient),
+and $K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
+
+We used mass action kinetics to model the protein conversion reactions within the network $\bar{r}_{j}$:
+\begin{equation}\label{eqn:rate-action}
+	\bar{r}_{j} = k_{j}^{max}\epsilon_{i}\prod_{s\in{m_{j}^{-}}}x_{s}
+\end{equation}
+where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$.
+The product in\linebreak Equation \eqref{eqn:rate-action} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).
+
+Applying the general framework to the reduced coagulation network resulted in 18 ordinary differential equations, 12 rate equations, and two control equations:
+\begin{eqnarray}
+	\frac{dx_{1}}{dt} & =& - r_{init,c4}f_{init,c4} \\ 								%C4
+	\frac{dx_{2}}{dt} &=& - r_{init,c2}f_{init,c2} \\  								%C2
+	\frac{dx_{3}}{dt} &=&  r_{init,c4}f_{init,c4} \\ 								%C4a
+	\frac{dx_{4}}{dt} &=& r_{init,c4}f_{init,c4} - r_{cp,c3c,form} \\ 					%C4b
+	\frac{dx_{5}}{dt} & = & r_{init,c2}f_{init,c2} - r_{cp,c3c,form} \\ 					%C2a
+	\frac{dx_{6}}{dt} &=& r_{init,c2}f_{init,c2} \\ 								%C2b
+	\frac{dx_{7}}{dt} &=& r_{c3,basal} - r_{cp,c3c,c3b} - r_{ap,c3c,c3b}\\ 				%C3
+	\frac{dx_{8}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  - k_{deg,c3a}*C3a \\		%C3a
+	\frac{dx_{9}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  -  r_{ap,c3c,form} \\		%C3b
+	\frac{dx_{10}}{dt} &=& r_{cp,c3c,form} - r_{cp,c5c,form} -  r_{cp,c3c,c4bp,inhib} \\			%CP C3C
+	\frac{dx_{11}}{dt} &=& r_{ap,c3c,form} - r_{ap,c5c,form} -  r_{ap,c3c,factorH,inhib} \\			%AP C3C
+	\frac{dx_{12}}{dt} &=& r_{cp,c5c,form} - r_{cp,c5c,c4bp,inhib} \\							%CP C5C
+	\frac{dx_{13}}{dt} &=& r_{cp,c5c,form}  \\											%AP C5C
+	\frac{dx_{14}}{dt} &=& - r_{cp,c5c,c5b} - r_{ap,c5c,c5b} \\								%C5
+	\frac{dx_{15}}{dt} &=&  r_{cp,c5c,c5b} + r_{ap,c5c,c5b} - k_{deg,c5a} \\					%C5a
+	\frac{dx_{16}}{dt} &=& r_{cp,c5c,c5b} + r_{ap,c5c,c5b} \\								%C5b
+	\frac{dx_{17}}{dt} &=& - r_{cp,c3c,c4bp,inhib} - r_{cp,c5c,c4bp,inhib} \\  				         %C4BP
+	\frac{dx_{18}}{dt} &=& - r_{ap,c3c,factorH,inhib} \\									%Factor H	
+\end{eqnarray}
+where \begin{multline}
+f{x} = f(Zymosan,C2,C2a,C2b,C4,C4a,C4b,C3,C3a,C3b,CP C3C Convertase,... \\
+           AP C3 Convertase,CP C5 Convertase,AP C5 Convertase,C5,C5a,... \\
+           ... C5b,C4BP,Factor H)^{T}
+\end{multline}
+
+The various rate and control equations are given by:
+\begin{eqnarray}
+	r_{init,c4} &=& \frac{k_{init,c4}*C4}{K_{init,c4,s} + C4} \\
+	r_{init,c2} &=& \frac{k_{init,c2}*C2}{K_{init,c2,s} + C2} \\
+	f_{init,c4} &=& \frac{Zymo^{\eta_{zymo,c4}}}{Zymo^{\eta_{zymo,c4}} + alpha_{c2}^{\eta_{zymo,c4}}} \\
+	f_{init,c2} &=& \frac{Zymo^{\eta_{zymo,c2}}}{Zymo^{\eta_{zymo,c2}} + alpha_{c4}^{\eta_{zymo,c2}}} \\
+	r_{c3,basal} &=& k_{c3b,basal}*C3 \\
+	r_{cp,c3c,c3b} &=& \frac{k_{cp,c3c}*CP,C3C*C3^{\eta_{cp,c3c}}}{K_{cp,c3c,s}^{\eta_{cp,c3c}} + C3^{\eta_{cp,c3c}}}  \\
+	r_{ap,c3c,c3b} &=& \frac{k_{ap,c3c}*AP,C3C*C3}{K_{ap,c3c,s} + C3} \\
+	r_{cp,c3c,form} &=& k_{cp,c3c,form}*C4b*C2a \\
+	r_{ap,c3c,form} &=& k_{ap,c3c,form}*C4b*C2a \\
+	r_{cp,c3c,c4bp,inhib} &=& k_{cp,c3c,inhib}*CP,C3C*C4b*C4BP \\
+	r_{ap,c3c,factorH,inhib} &=& k_{ap,c3c,inhib}*AP,C3C*FactorH \\ 
+	r_{cp,c5c,form} &=& k_{cp,c5c,form}*CP,C3C*C3b \\
+	r_{ap,c5c,form} &=& k_{ap,c5c,form}*AP,C3C*C3b \\
+	 r_{cp,c5c,c5b} &=& \frac{k_{cp,c5c}*CP,C5C*C5^{\eta_{cp,c5c}}}{K_{cp,c5c,s}^{\eta_{cp,c5c}} + C5^{\eta_{cp,c5c}}} \\
+	 r_{ap,c5c,c5b} &=& \frac{k_{ap,c5c}*AP,C5C*C5}{K_{ap,c5c,s} + C5} \\
+	 r_{cp,c5c,c4bp,inhib} &=& k_{cp,c5,inhib}*CP,C5C*C4BP
+\end{eqnarray}
+
+\subsection{Estimation of Model Parameters From Experimental Data - \textbf{Needs to be completed}}
+
+Model parameters were estimated by minimizing the difference between simulations and \linebreak experimental C3a and C5a measurements (squared residual):
+\begin{equation}\label{eqn:objective-function}
+	\min_{\mathbf{k}} \sum_{\tau=1}^{\mathcal{T}}\sum_{j=1}^{\mathcal{S}}\left(\frac{\hat{x}_{j}\left(\tau\right) - x_{j}\left(\tau,\mathbf{k}\right)}{\omega_{j}\left(\tau\right)}\right)^{2}
+\end{equation}where $\hat{x}_{j}\left(\tau\right)$ denotes the measured value of species $j$ at time $\tau$, $x_{j}\left(\tau,\mathbf{k}\right)$ denotes the simulated 
+value for species $j$ at time $\tau$, and $\omega_{j}\left(\tau\right)$ denotes the experimental measurement variance for species $j$ at time $\tau$. The outer summation is with respect to
+time, while the inner summation is with respect to state. We minimized the model residual using Particle swarm optimization (PSO) \citep{PSO}.
+PSO uses a \textit{swarming} metaheuristic to explore parameter spaces. 
+A strength of PSO is its ability to find the global minimum, even in the presence of potentially many local minima, by communicating the local
+error landscape experienced by each particle collectively to the swarm. Thus, PSO acts both as a local and a global search algorithm. 
+For each iteration, particles in the swarm compute their local error by evaluating the model equations using their specific parameter vector realization.
+From each of these local points, a globally best error is identified. Both the local and global error 
+are then used to update the parameter estimates of each particle using the rules:
+\begin{eqnarray}
+	\mathbf{\Delta}_{i} &=&\theta_{1}\mathbf{\Delta}_{i} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{k}_{i}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G} - \mathbf{k}_{i}\right) \\
+	\mathbf{k}_{i} &=& \mathbf{k}_{i} + \mathbf{\Delta}_{i}
+\end{eqnarray}where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the local best solution found by particle $i$, and
+$\mathcal{G}$ denotes the best solution found over the entire population of particles.
+The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
+parameters \linebreak($\mathcal{K}\times{1}$). In thus study, we used $\left(\theta_{1},\theta_{2},\theta_{3}\right) = \left(1.0, 0.05564, 0.02886\right)$. The quality of parameter
+estimates was measured using goodness of fit (model residual). The particle swarm optimization routine was implemented in the Python programming language. 
+All plots were made using the Matplotlib module of Python \citep{Matplotlib}.
+%The optimization and plotting code can be downloaded from http://www.varnerlab.org.
+%Experimental data,simulation code,parameter ensembles and initial conditions for each experimental simulation can be viewed and downloaded from https://github.com/jeffreyvarner/HybridCoagulationModel\_v1.
+
+\subsection{Global Sensitivity Analysis of Model Performance}
+
+We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}. 
+We computed the total sensitivity index of each parameter relative to four performance objectives, each objective was based on the sum of squared errors between model and experimental data for C3a alternate, C5a alternate, C3a lectin, and C5a lectin simulations.
+We established the sampling bounds for each parameter from the minimum and maximum value of that parameter in the parameter set ensemble. 
+We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges, 
+where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 42, so the total sensitivity indices were computed from
+11,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \citep{SALIB}.
 
@@
 %\end{figure}
+
 \begin{figure}[h]
@@
 $C3b$ can find to a $C3$ convertase and form a $C5$ convertase that cleaves $C5$ into $C5a$, and $C5b$ that undergoes a series of reactions to form the membrane attach complex ($MAC$).}
+\label{fig-schematic}
 \end{figure}
@@
 The shaded region denotes the distribution of $C3a$ and $C5a$ of the ensemble.}
+\label{fig-fit}
 \end{figure}
@@
 All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}
+\label{fig-prediction}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./Figures/Figure4_Sensitivity_Analysis_v2.pdf}
-\caption{Sensitivity Analysis}
+\caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
+Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at $0$ zymosan, (B) C5a $0$ zymosan, (C) C3a $1$ g zymosan, and (D) C5a $1$ g zymosan.
+The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions.
+The error bars are the 95 percent confidence interval.
+$k$ represents association rate, $km$ denote Michaelis-Menten saturation constants, and $alpha$ and $n$ refers to the exponentials of the control functions.}
+\label{fig-SA}
 \end{figure}

### 2016-05-16 Paper_v2.tex
Updated files and some content

--- 2016-05-14 Paper_v1.tex
@@
 \section*{Introduction}
-%The introduction has three paragraphs (introduction no longer than 3 pages):
-\begin{enumerate}
-	\item{\textbf{First~paragraph}: Introduce human complement system, history, role in adaptive/innate immunity. }
-
 Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
@@
 
-	-- Stuff about importance of complement (NEEDS TO BE FILLED)
-
-	\item{\textbf{Second~paragraph}: Introduce mathematical models of complement system, current place in the field, our work was not possible without xyz who pioneered abc.
-Address shortcomings in the field .}
-
 Given the complexity and importance of complement, developing mathematical models of complement are crucial to understanding its dynamics.
@@
 
-    \item{\textbf{Third~paragraph}: In this study, [Repeat the abstract with some additional detail]. Taken together, [killer statement].}
-
-    Towards this unmet need, we formulated a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes.
-This circumvents the need to have detailed mechanistic knowledge or use large number of ODEs to model complex dynamics.
-
-\end{enumerate}
-
-% Extra text from the introduction -
-%With the capacity to construct very large models using various mathematical formalisms \cite{chen2009input,tasseff2011modeling,luan2007computationally,mo2007genome,orth2011comprehensive,karr2012whole,buchel2013path2models,smallbone2010towards}, more often than not,
-
-%this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics  The new algorithm outperforms all the other meta-heuristics on the coagulation model.
-The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data.
-We also performed this comparison on  and found the same behavior.
-Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach.
-We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.
-%took into cognizance the fact that it may not be necessary to obtain the exact solution for high dimensional biological systems and that good enough solutions can be quickly obtained without expending a lot of objective function evaluations.
-Hence our current approach uses the power of population based heuristics along with DDS to obtain near optimal or good solutions.
-Though in theory we can combine any meta-heuristic with DDS, for the current purpose we used Particle Swarm Optimization (PSO).
-PSO unlike Genetic Algorithm (GA) or Differential Evolution (DE) does not have complex operations like cross over, mutation or recombination.
-It is simple to use and does not have a lot of parameters associated with other heuristics.
-However PSO is known to rapidly converge to a local optimum and thus several variants \cite{peer2003using,zhan2009adaptive,li2007fast} have been developed of which Multi swarm particle swarm optimization approaches (MLSPSO) \cite{zhao2008dynamic,liang2005dynamic} are one of the effective ones.
-We used MLSPSO to preclude bad regions of search and thereafter search within these regions using DDS.
+In this study, we formulated a hybrid modeling of the human complement system. [FILL ME IN].
 
@@
 
-The results are presented in \textbf{past~tense}. Each paragraph starts with a statement of the result in that paragraph in active voice.
-Each results paragraph ends with a Taken together type statement followed by a link statement e.g., Next we considered etc. When referring to figures, state what the figures shows (Fig. ZZ).
-
-\subsection*{2.1 Formulation of Reduced Order Complement Models}
-
+\subsection*{Formulation of a reduced order complement model.}
 We developed a reduced order extrinsic human complement network consisting of the most crucial steps of the human complement system (Fig.
@@
 
-\subsection*{2.2 Identification of Model Parameters Using Dynamically Dimensioned Searched (DDS)}
-
+\subsection*{Model were estimated using dynamically dimensioned search.}
 A critical challenge for any dynamic model is the estimation of kinetic parameters. We estimated
@@
 
-\subsection*{2.3 Validation of the Reduced Order Human Complement Model}
-IN PROGRESS
-
+\subsection*{Validation of the reduced order human complement model.}
 We tested the predictive power of the reduced order human complement model with validation data sets not used during model training.
@@
 
-\subsection*{2.4 Global Sensitivity Analysis of the Reduced Order Human Complement Model}
-
+\subsection*{Global Sensitivity analysis of the reduced order complement model}
 We conducted a Sobol's sensitivity analysis to estimate which parameters controlled the performance of the reduced order model.
@@
 
-Model parameters were estimated by minimizing the difference between simulations and \linebreak experimental C3a and C5a measurements (squared residual):
+Model parameters were estimated by minimizing the difference between simulations and experimental C3a and C5a measurements (squared residual):
 \begin{equation}\label{eqn:objective-function}
@@
 \subsection{Global Sensitivity Analysis of Model Performance}
-
 We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./Figures/ComplementSystemSchematic_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/ComplementSystemSchematic_v2.pdf}
 \caption{Simplified schematic representation of the human complement system.
@@
 The activation from the three pathways creates a cascades of reactions that forms the proteases, $C3$ Convertase that cleaves $C3$ into $C3a$, and $C3b$, the main effector molecule of the complement system.
-$C3b$ can find to a $C3$ convertase and form a $C5$ convertase that cleaves $C5$ into $C5a$, and $C5b$ that undergoes a series of reactions to form the membrane attach complex ($MAC$).}
-\label{fig-schematic}
+$C3b$ can find to a $C3$ convertase and form a $C5$ convertase that cleaves $C5$ into $C5a$, and $C5b$ that undergoes a series of reactions to form the membrane attach complex ($MAC$).}\label{fig-schematic}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./Figures/Figure2_Fit.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fit.pdf}
 \caption{Reduced order complement model training simulation for lectin and alternative pathway in presence of zymosan.
@@
 The red line shows the best-fit parameter, the black lines denotes the simulated mean value of $C3a$ or $C5a$ for a 50 parameter set ensemble.
-The shaded region denotes the distribution of $C3a$ and $C5a$ of the ensemble.}
-\label{fig-fit}
+The shaded region denotes the distribution of $C3a$ and $C5a$ of the ensemble.}\label{fig-fit}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./Figures/Figure3_Predictions_2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure3_Predictions_2.pdf}
 \caption{Reduced order complement model predictions of lectin and alternative pathway in presence of zymosan.
@@
 The red curve represents the best fit parameter, grey shaded region denotes the prediction results from 50 ensembles of parameter sets, and the black curve is the mean of the ensemble.
-All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}
-\label{fig-prediction}
+All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}\label{fig-prediction}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./Figures/Figure4_Sensitivity_Analysis_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v2.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
-Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at $0$ zymosan, (B) C5a $0$ zymosan, (C) C3a $1$ g zymosan, and (D) C5a $1$ g zymosan.
+Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0 zymosan, (B) C5a 0 zymosan, (C) C3a 1 g zymosan, and (D) C5a $1$ g zymosan.
 The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions.
 The error bars are the 95 percent confidence interval.
-$k$ represents association rate, $km$ denote Michaelis-Menten saturation constants, and $alpha$ and $n$ refers to the exponentials of the control functions.}
-\label{fig-SA}
+$k$ represents association rate, $km$ denote Michaelis-Menten saturation constants, and $alpha$ and $n$ refers to the exponentials of the control functions.}\label{fig-SA}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./figs/Figure3_Predictions_2.pdf}
+\caption{Reduced order complement model predictions of lectin and alternative pathway in presence of zymosan.
+(A-F) Simulation of complement dynamics in the presence of zymosan were conducted for a range of trigger values ($0.1$, $0.01$, and $0.001$ g zymosan).
+The time-course profiles of $C3a$ and $C5a$ under three different zymosan concentrations were simulated using 50 ensembles of trained parameter sets against experimental data of Shaw et al [REF].
+The red curve represents the best fit parameter, grey shaded region denotes the prediction results from 50 ensembles of parameter sets, and the black curve is the mean of the ensemble.
+All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v2.pdf}
+\caption{Sensitivity Analysis}
 \end{figure}

### 2016-05-16 Paper_v2.tex
Initial updates

@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Dynamic Modeling of Human Complement System using Reduced Ordered Models}}}
+{\par\centering\textbf{\Large {Dynamic modeling of human complement}}}
 \vspace{0.05in}
@@
 {\par \centering $^{\#}$ Denotes equal contribution}
+\vspace{0.1in}
 {\par \centering $^{*}$Corresponding author:}
@@
 
-In this study, we formulated a hybrid modeling of the human complement system. [FILL ME IN].
+In this study, we formulated a model of the human complement system. [FILL ME IN].
 
@@
 Conversely, if $\sigma_{ij}<0$, species $i$ is consumed by reaction $j$, while $\sigma_{ij} = 0$ indicates species $i$ is not connected with reaction $j$.
-The system material balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$, which were specified by the experimental setup and human serum complement protein levels.
+Species balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$.
 
-Each reaction rate was written as either the product of two terms, a kinetic term ($\bar{r}_{j}$) and a control term ($f_{j}$), or $\bar{r}_{j}$ for the case $f_{j}$ = 1:
-\begin{equation}\label{eqn:rate-factor}
-	r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right) = \bar{r}_{j}f_{j}
-\end{equation}
+The reaction rate was written as the product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$), $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
+In this study, we used either saturation or mass action kinetics.
+The control term $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
+For each rate, we used a rule-based approach to select from competing control factors.
+If rate j was influenced by $1,\dots,m$ factors, we modeled this relationship as
+$v_{j} = \mathcal{I}_{j}\left(f_{1j}\left(\cdot\right),\hdots,f_{mj}\left(\cdot\right)\right)$
+where $0\leq f_{ij}\left(\cdot\right)\leq 1$ denotes a regulatory transfer function quantifying the influence of factor $i$ on rate $j$.
+The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
+variable. Each regulatory transfer function took the form:
+\begin{equation}\label{eqn:control-factor}
+	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right) = k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
+\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
+In this study, we used $\mathcal{I}_{j}\in\left\{\max,\min\right\}$ \cite{pr3010138}. If a process has no modifying factors, $v_{j} = 1$.
 
-Each control term, $f_{j}\left(\mathcal{Z}\right)$ took the form:
-\begin{equation}\label{eqn:control-factor}
-	f_{j}\left(\mathcal{Z}_{j},\alpha_{j}\right) = \frac{\mathcal{Z}_{j}^{\eta}}{\mathcal{Z}_{j}^{\eta} + \alpha_{j}^{\eta}}
-\end{equation}where $\mathcal{Z}_{j}$ denotes the abundance of the $j$ factor (e.g., initiator abundance),
-$\alpha_{j}$ denotes the species gain parameter, and $\eta$ a cooperativity parameter (similar to a Hill coefficient).
-
-We used saturation kinetics to model the lectin pathway activation and C3 and C5 convertase activity reaction terms $\bar{r}_{j}$:
+We used saturation kinetics to model the lectin pathway activation and C3 and C5 convertase activity $\bar{r}_{j}$:
 \begin{equation}\label{eqn:rate-saturation}
@@
 \end{equation}
-where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$, $\eta$ denotes a cooperativity parameter (similar to a Hill coefficient),
-and $K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
-
-We used mass action kinetics to model the protein conversion reactions within the network $\bar{r}_{j}$:
+where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$, $\eta$ denotes a cooperativity parameter (similar to a Hill coefficient), and $K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
+On the other hand,
+we used mass action kinetics to model the protein conversion reactions within the network $\bar{r}_{j}$:
 \begin{equation}\label{eqn:rate-action}
@@
 where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$.
-The product in\linebreak Equation \eqref{eqn:rate-action} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).
+The product in Eqn \eqref{eqn:rate-action} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).
 
-Applying the general framework to the reduced coagulation network resulted in 18 ordinary differential equations, 12 rate equations, and two control equations:
-\begin{eqnarray}
-	\frac{dx_{1}}{dt} & =& - r_{init,c4}f_{init,c4} \\ 								%C4
-	\frac{dx_{2}}{dt} &=& - r_{init,c2}f_{init,c2} \\  								%C2
-	\frac{dx_{3}}{dt} &=&  r_{init,c4}f_{init,c4} \\ 								%C4a
-	\frac{dx_{4}}{dt} &=& r_{init,c4}f_{init,c4} - r_{cp,c3c,form} \\ 					%C4b
-	\frac{dx_{5}}{dt} & = & r_{init,c2}f_{init,c2} - r_{cp,c3c,form} \\ 					%C2a
-	\frac{dx_{6}}{dt} &=& r_{init,c2}f_{init,c2} \\ 								%C2b
-	\frac{dx_{7}}{dt} &=& r_{c3,basal} - r_{cp,c3c,c3b} - r_{ap,c3c,c3b}\\ 				%C3
-	\frac{dx_{8}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  - k_{deg,c3a}*C3a \\		%C3a
-	\frac{dx_{9}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  -  r_{ap,c3c,form} \\		%C3b
-	\frac{dx_{10}}{dt} &=& r_{cp,c3c,form} - r_{cp,c5c,form} -  r_{cp,c3c,c4bp,inhib} \\			%CP C3C
-	\frac{dx_{11}}{dt} &=& r_{ap,c3c,form} - r_{ap,c5c,form} -  r_{ap,c3c,factorH,inhib} \\			%AP C3C
-	\frac{dx_{12}}{dt} &=& r_{cp,c5c,form} - r_{cp,c5c,c4bp,inhib} \\							%CP C5C
-	\frac{dx_{13}}{dt} &=& r_{cp,c5c,form}  \\											%AP C5C
-	\frac{dx_{14}}{dt} &=& - r_{cp,c5c,c5b} - r_{ap,c5c,c5b} \\								%C5
-	\frac{dx_{15}}{dt} &=&  r_{cp,c5c,c5b} + r_{ap,c5c,c5b} - k_{deg,c5a} \\					%C5a
-	\frac{dx_{16}}{dt} &=& r_{cp,c5c,c5b} + r_{ap,c5c,c5b} \\								%C5b
-	\frac{dx_{17}}{dt} &=& - r_{cp,c3c,c4bp,inhib} - r_{cp,c5c,c4bp,inhib} \\  				         %C4BP
-	\frac{dx_{18}}{dt} &=& - r_{ap,c3c,factorH,inhib} \\									%Factor H
-\end{eqnarray}
-where \begin{multline}
-f{x} = f(Zymosan,C2,C2a,C2b,C4,C4a,C4b,C3,C3a,C3b,CP C3C Convertase,... \\
-           AP C3 Convertase,CP C5 Convertase,AP C5 Convertase,C5,C5a,... \\
-           ... C5b,C4BP,Factor H)^{T}
-\end{multline}
-
-The various rate and control equations are given by:
-\begin{eqnarray}
-	r_{init,c4} &=& \frac{k_{init,c4}*C4}{K_{init,c4,s} + C4} \\
-	r_{init,c2} &=& \frac{k_{init,c2}*C2}{K_{init,c2,s} + C2} \\
-	f_{init,c4} &=& \frac{Zymo^{\eta_{zymo,c4}}}{Zymo^{\eta_{zymo,c4}} + alpha_{c2}^{\eta_{zymo,c4}}} \\
-	f_{init,c2} &=& \frac{Zymo^{\eta_{zymo,c2}}}{Zymo^{\eta_{zymo,c2}} + alpha_{c4}^{\eta_{zymo,c2}}} \\
-	r_{c3,basal} &=& k_{c3b,basal}*C3 \\
-	r_{cp,c3c,c3b} &=& \frac{k_{cp,c3c}*CP,C3C*C3^{\eta_{cp,c3c}}}{K_{cp,c3c,s}^{\eta_{cp,c3c}} + C3^{\eta_{cp,c3c}}}  \\
-	r_{ap,c3c,c3b} &=& \frac{k_{ap,c3c}*AP,C3C*C3}{K_{ap,c3c,s} + C3} \\
-	r_{cp,c3c,form} &=& k_{cp,c3c,form}*C4b*C2a \\
-	r_{ap,c3c,form} &=& k_{ap,c3c,form}*C4b*C2a \\
-	r_{cp,c3c,c4bp,inhib} &=& k_{cp,c3c,inhib}*CP,C3C*C4b*C4BP \\
-	r_{ap,c3c,factorH,inhib} &=& k_{ap,c3c,inhib}*AP,C3C*FactorH \\
-	r_{cp,c5c,form} &=& k_{cp,c5c,form}*CP,C3C*C3b \\
-	r_{ap,c5c,form} &=& k_{ap,c5c,form}*AP,C3C*C3b \\
-	 r_{cp,c5c,c5b} &=& \frac{k_{cp,c5c}*CP,C5C*C5^{\eta_{cp,c5c}}}{K_{cp,c5c,s}^{\eta_{cp,c5c}} + C5^{\eta_{cp,c5c}}} \\
-	 r_{ap,c5c,c5b} &=& \frac{k_{ap,c5c}*AP,C5C*C5}{K_{ap,c5c,s} + C5} \\
-	 r_{cp,c5c,c4bp,inhib} &=& k_{cp,c5,inhib}*CP,C5C*C4BP
-\end{eqnarray}
-
-\subsection{Estimation of Model Parameters From Experimental Data - \textbf{Needs to be completed}}
-
+\subsection*{Estimation of an ensemble of model parameters.}
 Model parameters were estimated by minimizing the difference between simulations and experimental C3a and C5a measurements (squared residual):
@@
 
-\subsection{Global Sensitivity Analysis of Model Performance}
+\subsection*{Global sensitivity analysis of model performance}
 We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}.
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure3_Predictions_2.pdf}
-\caption{Reduced order complement model predictions of lectin and alternative pathway in presence of zymosan.
-(A-F) Simulation of complement dynamics in the presence of zymosan were conducted for a range of trigger values ($0.1$, $0.01$, and $0.001$ g zymosan).
-The time-course profiles of $C3a$ and $C5a$ under three different zymosan concentrations were simulated using 50 ensembles of trained parameter sets against experimental data of Shaw et al [REF].
-The red curve represents the best fit parameter, grey shaded region denotes the prediction results from 50 ensembles of parameter sets, and the black curve is the mean of the ensemble.
-All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v2.pdf}
-\caption{Sensitivity Analysis}
-\end{figure}
-
 \clearpage
@@
 
+\section*{Supplemental materials.}
+\subsubsection*{Model equations.}
+Applying the general framework to the reduced coagulation network resulted in 18 ordinary differential equations, 12 rate equations, and two control equations:
+\begin{eqnarray}
+	\frac{dx_{1}}{dt} & =& - r_{init,c4}f_{init,c4} \\ 								%C4
+	\frac{dx_{2}}{dt} &=& - r_{init,c2}f_{init,c2} \\  								%C2
+	\frac{dx_{3}}{dt} &=&  r_{init,c4}f_{init,c4} \\ 								%C4a
+	\frac{dx_{4}}{dt} &=& r_{init,c4}f_{init,c4} - r_{cp,c3c,form} \\ 					%C4b
+	\frac{dx_{5}}{dt} & = & r_{init,c2}f_{init,c2} - r_{cp,c3c,form} \\ 					%C2a
+	\frac{dx_{6}}{dt} &=& r_{init,c2}f_{init,c2} \\ 								%C2b
+	\frac{dx_{7}}{dt} &=& r_{c3,basal} - r_{cp,c3c,c3b} - r_{ap,c3c,c3b}\\ 				%C3
+	\frac{dx_{8}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  - k_{deg,c3a}*C3a \\		%C3a
+	\frac{dx_{9}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  -  r_{ap,c3c,form} \\		%C3b
+	\frac{dx_{10}}{dt} &=& r_{cp,c3c,form} - r_{cp,c5c,form} -  r_{cp,c3c,c4bp,inhib} \\			%CP C3C
+	\frac{dx_{11}}{dt} &=& r_{ap,c3c,form} - r_{ap,c5c,form} -  r_{ap,c3c,factorH,inhib} \\			%AP C3C
+	\frac{dx_{12}}{dt} &=& r_{cp,c5c,form} - r_{cp,c5c,c4bp,inhib} \\							%CP C5C
+	\frac{dx_{13}}{dt} &=& r_{cp,c5c,form}  \\											%AP C5C
+	\frac{dx_{14}}{dt} &=& - r_{cp,c5c,c5b} - r_{ap,c5c,c5b} \\								%C5
+	\frac{dx_{15}}{dt} &=&  r_{cp,c5c,c5b} + r_{ap,c5c,c5b} - k_{deg,c5a} \\					%C5a
+	\frac{dx_{16}}{dt} &=& r_{cp,c5c,c5b} + r_{ap,c5c,c5b} \\								%C5b
+	\frac{dx_{17}}{dt} &=& - r_{cp,c3c,c4bp,inhib} - r_{cp,c5c,c4bp,inhib} \\  				         %C4BP
+	\frac{dx_{18}}{dt} &=& - r_{ap,c3c,factorH,inhib} \\									%Factor H
+\end{eqnarray}
+where \begin{multline}
+f{x} = f(Zymosan,C2,C2a,C2b,C4,C4a,C4b,C3,C3a,C3b,CP C3C Convertase,... \\
+           AP C3 Convertase,CP C5 Convertase,AP C5 Convertase,C5,C5a,... \\
+           ... C5b,C4BP,Factor H)^{T}
+\end{multline}
+
+The various rate and control equations are given by:
+\begin{eqnarray}
+	r_{init,c4} &=& \frac{k_{init,c4}*C4}{K_{init,c4,s} + C4} \\
+	r_{init,c2} &=& \frac{k_{init,c2}*C2}{K_{init,c2,s} + C2} \\
+	f_{init,c4} &=& \frac{Zymo^{\eta_{zymo,c4}}}{Zymo^{\eta_{zymo,c4}} + alpha_{c2}^{\eta_{zymo,c4}}} \\
+	f_{init,c2} &=& \frac{Zymo^{\eta_{zymo,c2}}}{Zymo^{\eta_{zymo,c2}} + alpha_{c4}^{\eta_{zymo,c2}}} \\
+	r_{c3,basal} &=& k_{c3b,basal}*C3 \\
+	r_{cp,c3c,c3b} &=& \frac{k_{cp,c3c}*CP,C3C*C3^{\eta_{cp,c3c}}}{K_{cp,c3c,s}^{\eta_{cp,c3c}} + C3^{\eta_{cp,c3c}}}  \\
+	r_{ap,c3c,c3b} &=& \frac{k_{ap,c3c}*AP,C3C*C3}{K_{ap,c3c,s} + C3} \\
+	r_{cp,c3c,form} &=& k_{cp,c3c,form}*C4b*C2a \\
+	r_{ap,c3c,form} &=& k_{ap,c3c,form}*C4b*C2a \\
+	r_{cp,c3c,c4bp,inhib} &=& k_{cp,c3c,inhib}*CP,C3C*C4b*C4BP \\
+	r_{ap,c3c,factorH,inhib} &=& k_{ap,c3c,inhib}*AP,C3C*FactorH \\
+	r_{cp,c5c,form} &=& k_{cp,c5c,form}*CP,C3C*C3b \\
+	r_{ap,c5c,form} &=& k_{ap,c5c,form}*AP,C3C*C3b \\
+	 r_{cp,c5c,c5b} &=& \frac{k_{cp,c5c}*CP,C5C*C5^{\eta_{cp,c5c}}}{K_{cp,c5c,s}^{\eta_{cp,c5c}} + C5^{\eta_{cp,c5c}}} \\
+	 r_{ap,c5c,c5b} &=& \frac{k_{ap,c5c}*AP,C5C*C5}{K_{ap,c5c,s} + C5} \\
+	 r_{cp,c5c,c4bp,inhib} &=& k_{cp,c5,inhib}*CP,C5C*C4BP
+\end{eqnarray}
+
 % Supplemental figures go here ...

### 2016-05-19 Paper_v2.tex
Updates, added text

@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Dynamic modeling of human complement}}}
+{\par\centering\textbf{\Large {Reduced order modeling and analysis of the human complement subsystem}}}
 \vspace{0.05in}
@@
 \vspace{0.1in}
-{\par \centering \textbf{Running Title:}~Dynamic Modeling of Human Complement System using Reduced Ordered Models}
+{\par \centering \textbf{Running Title:}~Reduced order model of complement}
 \vspace{0.1in}
-{\par \centering \textbf{To be submitted:}~\emph{???????}}
+{\par \centering \textbf{To be submitted:}~\emph{PLoS~ONE}}
 \vspace{0.5in}
@@
 \section*{Abstract}
-Fill me in.
+Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
+In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
+The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
+We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
+Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
+The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
+Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
+We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
+We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
+Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
+The reduced order hybrid approach produced a surprisingly predictive human complement model, much similar to the study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
+Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into roles of governing parameters that mediate the complement system.
+Deeply understanding how the main governing parameters change with respective to the different combinations of pathway activation will greatly aid in development of drugs for strategic therapeutic targets.
+Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
 
@@
 	The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the MAC complex.
+
 The activation of complement and formation of C5a and MAC complex is regulated at different points through a number of plasma and host cell proteins.
-C4 binding protein (C4BP), ---- (NEEDS TO BE FILLED)
+The initiation of the classical pathway through the attachment of C1 to an antibody is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
+C1-Inh irreversibly binds to and deactivates the active subunits of component C1 to prevent spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
+The serum and host-tissue regulation of the upstream elements of the complement system is also achieved through the binding of C4 binding protein (C4BP) to C4b and through the binding of factor H to C3b \cite{blom2001structural}.
+These proteins are also capable of binding their respective components in the convertase form.
+Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
+Delay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
+MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
+Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
 
-Given the complexity and importance of complement, developing mathematical models of complement are crucial to understanding its dynamics.
+Given the complexity and importance, developing mathematical models of complement are crucial to understanding its dynamics.
 Complement models have typically been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
@@
 
-In this study, we formulated a model of the human complement system. [FILL ME IN].
+In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
+The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
+We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
+Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
+The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
+Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
+We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
+We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
+Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
+The reduced order hybrid approach produced a surprisingly predictive human complement model, much similar to the study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
+Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into roles of governing parameters that mediate the complement system.
+Deeply understanding how the main governing parameters change with respective to the different combinations of pathway activation will greatly aid in development of drugs for strategic therapeutic targets.
+Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
 
@@
 
-\subsection*{Formulation of a reduced order complement model.}
-We developed a reduced order extrinsic human complement network consisting of the most crucial steps of the human complement system (Fig.
+\subsection*{Formulation of a reduced order complement model}
+We developed a reduced order human complement network consisting of the most crucial steps of the human complement system (Fig.
 \ref{fig-schematic}).
@@
 In doing so we aim to capture a complex biological phenomenon using a few simple ordinary differential equations.
-A trigger event initiates the lectin pathway in the presence of zymosan, which activates the cleavage of $C2$ and $C4$ into $C2a$ and $C2b$, and $C4a$ and $C4b$ respectively.
-Classical Pathway ($CP$) $C3 convertase$ ($C4aC2b$) is by the combination of $C4a$ and $C2b$, which catalyzes the cleavage of $C3$ into $C3a$ and $C3b$.
-Similarly, the activation of the alternative pathways happens through the spontaneous hydrolysis of $C3$ which facilitates the cleavage of $C3$.
-$C3b$ then could combine with with $C3$ to form alternate pathway ($AP$) $C3 convertase$.
-The both  versions of the $C3 convertase$ catalyzes the cleavage of $C3$ into $C3a$ and $C3b$, and $C3b$ can then combine with either $CP$ or $AP C3 convertase$ to form $C5 convertase$, $CP or AP$ respectively that is responsible for the cleavage of $C5$ to $C5a$ and $C5b$.
-Lectin pathway activation was approximated using a combination of saturation kinetics and Hill-like function control functions.These control coefficients then modified the rates of model processes at each time step.
+A trigger event initiates the lectin pathway in the presence of zymosan, which activates the cleavage of C2 and C4 into C2a and C2b, and C4a and C4b respectively.
+Classical Pathway (CP) C3 convertase (C4aC2b) is a combination of C4a and C2b, which catalyzes the cleavage of C3 into C3a and C3b.
+Similarly, the activation of the alternative pathways happens through the spontaneous hydrolysis of C3 which facilitates the cleavage of C3.
+C3b then could combine with with C3 to form alternate pathway (AP) C3 convertase.
+Both C3 convertases catalyze the cleavage of C3 into C3a and C3b, and C3b can then combine with either CP or AP C3 convertase to form C5 convertase, CP or AP respectively that is responsible for the cleavage of C5 to C5a and C5b.
+Lectin pathway activation was approximated using a combination of saturation kinetics and Hill-like function control functions.
+These control coefficients then modified the rates of model processes at each time step.
 Hill-like transfer functions $0 \le \mathbf{f (Z)}  \le 1$ quantified the contribution of components upon a target process, in this study, $\mathbf{Z}$ represents the abundance of the initiator.
 Taken together, while the reduced order human complement model encodes significant biological complexity, it is highly compact (consisting of only 18 differential equations).
-Thus, it will serve as an excellent proof
-of principle example to study the reduction of a highly complex human subsystem.
+Thus, it will serve as an excellent proof of principle example to study the reduction of a highly complex human subsystem.
 
-\subsection*{Model were estimated using dynamically dimensioned search.}
+\subsection*{An ensemble of complement models was estimated using dynamically dimensioned search.}
 A critical challenge for any dynamic model is the estimation of kinetic parameters. We estimated
@@
 \ref{fig-fit} C and D).
-
-The reduced human complement model captured the behavior of the alternative and lectin pathways through the time-course abundance of $C3a$ and $C5a$ (Fig.
-\ref{fig-fit}).
-However we were not able to capture the curvature of the $C5a$ alternate (Fig.
-\ref{fig-fit}), the decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
-Taken together, the model identification results suggested that our reduced order approach could reproduce a panel of lectin pathway intiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
+The reduced human complement model captured the behavior of the alternative and lectin pathways through the time-course abundance of C3a and C5a (Fig. \ref{fig-fit}).
+However we were not able to capture the curvature of the C5a alternative (Fig. \ref{fig-fit}).
+The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
+Taken together, the model identification results suggested that our reduced order approach could reproduce a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
 However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
 
-\subsection*{Validation of the reduced order human complement model.}
 We tested the predictive power of the reduced order human complement model with validation data sets not used during model training.
@@
 All kinetic and control parameters were
-fixed for the validation simulations. The reduced order model predicted the $C3a$ and $C5a$ time-course profiles at a qualitative level (Fig. \ref{fig-prediction}).
+fixed for the validation simulations.
+The reduced order model predicted the $C3a$ and $C5a$ time-course profiles at a qualitative level (Fig. \ref{fig-prediction}).
 
@@
 \end{enumerate}
+
+In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
+The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
+We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
+Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
+The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
+Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
+We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
+We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
+Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
+The reduced order hybrid approach produced a surprisingly predictive human complement model, much similar to the study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
+Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into roles of governing parameters that mediate the complement system.
+Deeply understanding how the main governing parameters change with respective to the different combinations of pathway activation will greatly aid in development of drugs for strategic therapeutic targets.
+Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
+
+The performance of the reduced order complement model was impressive given its limited size.
+However, there are several critical questions that should be explored following this study.
+A logical progression for this work would include expanding the network to include the classical pathway and the formation of the membrane attack complex (MAC).
+It is unclear whether the addition of the classical pathway will decrease the prediction of our existing model due to the cross-talk between the classical and lectin activation shown by Liu et al \cite{liu2011computational}.
+One potential approach in addressing such difficulties would be the incorporation of additional species such as C reactive proteins (CRP) and L-ficolin (LF) that involved in complement initiation of classical and lectin pathways.
+The influence of CRP, LF and the cross-talk can be captured through additional control functions that act upon the initiation pathways in a logical integration rule developed by Wayman and coworkers \cite{wayman2015dynamic}.
+Another issue with our reduced order model involve the omitted species that are implicitly lumped together with our effective kinetics and control parameters.
+Due to the reduction of parameters, the model cannot determine the dynamics or explicit impact of the omitted species on the system.
+However, we have created a hierarchy approach for parameter estimation that can be used to uncouple the kinetic parameter and contribution of any additional complement proteins and regulators.
+Using this simple and versatile modeling approach that we created, we took the first step in the development of a computation toolkit that can be readily used in a clinical setting.
+Our reduced order complement model is computationally inexpensive, and versatile so it could easily be incorporated into pre-existing or new pharmacokinetic models.
+Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
 
@@
 \section*{Supplemental materials.}
-\subsubsection*{Model equations.}
-Applying the general framework to the reduced coagulation network resulted in 18 ordinary differential equations, 12 rate equations, and two control equations:
+\subsection*{Model equations.}
+The reduced-order complement model consisted of 18 ordinary differential equations, 12 rate equations, and two control equations:
 \begin{eqnarray}
-	\frac{dx_{1}}{dt} & =& - r_{init,c4}f_{init,c4} \\ 								%C4
-	\frac{dx_{2}}{dt} &=& - r_{init,c2}f_{init,c2} \\  								%C2
-	\frac{dx_{3}}{dt} &=&  r_{init,c4}f_{init,c4} \\ 								%C4a
-	\frac{dx_{4}}{dt} &=& r_{init,c4}f_{init,c4} - r_{cp,c3c,form} \\ 					%C4b
-	\frac{dx_{5}}{dt} & = & r_{init,c2}f_{init,c2} - r_{cp,c3c,form} \\ 					%C2a
-	\frac{dx_{6}}{dt} &=& r_{init,c2}f_{init,c2} \\ 								%C2b
-	\frac{dx_{7}}{dt} &=& r_{c3,basal} - r_{cp,c3c,c3b} - r_{ap,c3c,c3b}\\ 				%C3
-	\frac{dx_{8}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  - k_{deg,c3a}*C3a \\		%C3a
-	\frac{dx_{9}}{dt} &=& r_{c3,basal}  + r_{cp,c3c,c3b} + r_{ap,c3c,c3b}  -  r_{ap,c3c,form} \\		%C3b
-	\frac{dx_{10}}{dt} &=& r_{cp,c3c,form} - r_{cp,c5c,form} -  r_{cp,c3c,c4bp,inhib} \\			%CP C3C
-	\frac{dx_{11}}{dt} &=& r_{ap,c3c,form} - r_{ap,c5c,form} -  r_{ap,c3c,factorH,inhib} \\			%AP C3C
-	\frac{dx_{12}}{dt} &=& r_{cp,c5c,form} - r_{cp,c5c,c4bp,inhib} \\							%CP C5C
-	\frac{dx_{13}}{dt} &=& r_{cp,c5c,form}  \\											%AP C5C
-	\frac{dx_{14}}{dt} &=& - r_{cp,c5c,c5b} - r_{ap,c5c,c5b} \\								%C5
-	\frac{dx_{15}}{dt} &=&  r_{cp,c5c,c5b} + r_{ap,c5c,c5b} - k_{deg,c5a} \\					%C5a
-	\frac{dx_{16}}{dt} &=& r_{cp,c5c,c5b} + r_{ap,c5c,c5b} \\								%C5b
-	\frac{dx_{17}}{dt} &=& - r_{cp,c3c,c4bp,inhib} - r_{cp,c5c,c4bp,inhib} \\  				         %C4BP
-	\frac{dx_{18}}{dt} &=& - r_{ap,c3c,factorH,inhib} \\									%Factor H
-\end{eqnarray}
-where \begin{multline}
-f{x} = f(Zymosan,C2,C2a,C2b,C4,C4a,C4b,C3,C3a,C3b,CP C3C Convertase,... \\
-           AP C3 Convertase,CP C5 Convertase,AP C5 Convertase,C5,C5a,... \\
-           ... C5b,C4BP,Factor H)^{T}
-\end{multline}
-
-The various rate and control equations are given by:
+	\frac{dx_{1}}{dt} & =& - r_{1}f_{1} \\ 								%C4
+	\frac{dx_{2}}{dt} &=& - r_{2}f_{2} \\  								%C2
+	\frac{dx_{3}}{dt} &=&  r_{1}f_{1} \\ 								%C4a
+	\frac{dx_{4}}{dt} &=& r_{1}f_{1} - r_{6} \\ 					%C4b
+	\frac{dx_{5}}{dt} & = & r_{2}f_{2} - r_{6} \\ 					%C2a
+	\frac{dx_{6}}{dt} &=& r_{2}f_{2} \\ 								%C2b
+	\frac{dx_{7}}{dt} &=& r_{3} - r_{4} - r_{5}\\ 				%C3
+	\frac{dx_{8}}{dt} &=& r_{3}  + r_{4} + r_{5}  - k_{deg,c3a}*C3a \\		%C3a
+	\frac{dx_{9}}{dt} &=& r_{3}  + r_{4} + r_{5}  -  r_{7} \\		%C3b
+	\frac{dx_{10}}{dt} &=& r_{6} - r_{10} -  r_{8} \\			%CP C3C
+	\frac{dx_{11}}{dt} &=& r_{7} - r_{11} -  r_{9} \\			%AP C3C
+	\frac{dx_{12}}{dt} &=& r_{10} - r_{14} \\							%CP C5C
+	\frac{dx_{13}}{dt} &=& r_{10}  \\											%AP C5C
+	\frac{dx_{14}}{dt} &=& - r_{12} - r_{13} \\								%C5
+	\frac{dx_{15}}{dt} &=&  r_{12} + r_{13} - k_{deg,c5a} \\					%C5a
+	\frac{dx_{16}}{dt} &=& r_{12} + r_{13} \\								%C5b
+	\frac{dx_{17}}{dt} &=& - r_{8} - r_{14} \\  				         %C4BP
+	\frac{dx_{18}}{dt} &=& - r_{9} \\									%Factor H
+\end{eqnarray}where the rate equations are given by:
 \begin{eqnarray}
-	r_{init,c4} &=& \frac{k_{init,c4}*C4}{K_{init,c4,s} + C4} \\
-	r_{init,c2} &=& \frac{k_{init,c2}*C2}{K_{init,c2,s} + C2} \\
-	f_{init,c4} &=& \frac{Zymo^{\eta_{zymo,c4}}}{Zymo^{\eta_{zymo,c4}} + alpha_{c2}^{\eta_{zymo,c4}}} \\
-	f_{init,c2} &=& \frac{Zymo^{\eta_{zymo,c2}}}{Zymo^{\eta_{zymo,c2}} + alpha_{c4}^{\eta_{zymo,c2}}} \\
-	r_{c3,basal} &=& k_{c3b,basal}*C3 \\
-	r_{cp,c3c,c3b} &=& \frac{k_{cp,c3c}*CP,C3C*C3^{\eta_{cp,c3c}}}{K_{cp,c3c,s}^{\eta_{cp,c3c}} + C3^{\eta_{cp,c3c}}}  \\
-	r_{ap,c3c,c3b} &=& \frac{k_{ap,c3c}*AP,C3C*C3}{K_{ap,c3c,s} + C3} \\
-	r_{cp,c3c,form} &=& k_{cp,c3c,form}*C4b*C2a \\
-	r_{ap,c3c,form} &=& k_{ap,c3c,form}*C4b*C2a \\
-	r_{cp,c3c,c4bp,inhib} &=& k_{cp,c3c,inhib}*CP,C3C*C4b*C4BP \\
-	r_{ap,c3c,factorH,inhib} &=& k_{ap,c3c,inhib}*AP,C3C*FactorH \\
-	r_{cp,c5c,form} &=& k_{cp,c5c,form}*CP,C3C*C3b \\
-	r_{ap,c5c,form} &=& k_{ap,c5c,form}*AP,C3C*C3b \\
-	 r_{cp,c5c,c5b} &=& \frac{k_{cp,c5c}*CP,C5C*C5^{\eta_{cp,c5c}}}{K_{cp,c5c,s}^{\eta_{cp,c5c}} + C5^{\eta_{cp,c5c}}} \\
-	 r_{ap,c5c,c5b} &=& \frac{k_{ap,c5c}*AP,C5C*C5}{K_{ap,c5c,s} + C5} \\
-	 r_{cp,c5c,c4bp,inhib} &=& k_{cp,c5,inhib}*CP,C5C*C4BP
+	r_{1} &=& \frac{k_{i1}(C4)}{(K_{1s} + C4)} \\
+	r_{2} &=& \frac{k_{2}(C2)}{(K_{2s} + C2)} \\
+	f_{1} &=& \frac{Zymo^{\eta_{1}}}{(Zymo^{\eta_{1}} + \alpha_{1}^{\eta_{1}})} \\
+	f_{2} &=& \frac{Zymo^{\eta_{2}}}{(Zymo^{\eta_{2}} + \alpha_{2}^{\eta_{2}})} \\
+	r_{3} &=& k_{3}(C3) \\
+	r_{4} &=& \frac{k_{4}(C3C_{L})(C3^{\eta_{3})}}{(K_{4s}^{\eta_{3}} + C3^{\eta_{3}})}  \\
+	r_{5} &=& \frac{k_{5}(C3C_{A})(C3)}{(K_{5s} + C3)} \\
+	r_{6} &=& k_{6}(C4b)(C2a) \\
+	r_{7} &=& k_{7}(C4b)(C2a) \\
+	r_{8} &=& k_{8}(C3C_{L})(C4b)(C4BP) \\
+	r_{9} &=& k_{9}(C3C_{A})(FactorH) \\
+	r_{10} &=& k_{10}(C3C_{L})(C3b) \\
+	r_{11} &=& k_{11}(C3C_{A})(C3b) \\
+	r_{12} &=& \frac{k_{12}(C5C_{L})(C5^{\eta_{4}})}{(K_{12s}^{\eta_{4}} + C5^{\eta_{4}})} \\
+	r_{13} &=& \frac{k_{13}(C5C_{A})(C5)}{(K_{13s} + C5)} \\
+	r_{14} &=& k_{14}(C5C_{L})(C4BP)
 \end{eqnarray}

### 2016-05-19 Paper_v2.tex
Updates, added text

@@
 Knowing that the kinetic and control parameters of the lectin pathway does not affect the dynamics of the alternate pathway, we used a hierarchal approach that estimated the parameters for the alternative pathway and lectin pathway separately.
-For the alternative pathway, we utilized the time-course experimental measurements of Morad and coworkers \cite{morad2015time} of $C3a$ and $C5a$ in the absence of zymosan and only allowed the alternative parameters to vary (Fig.
+For the alternative pathway, we utilized the time-course experimental measurements of Morad and coworkers \cite{morad2015time} of C3a and C5a in the absence of zymosan and only allowed the alternative parameters to vary (Fig.
 \ref{fig-fit} A and B).
 The estimated alternate parameters was then fixed for the determination of lectin pathway parameters.
-The training for the lectin parameters, we used the experimental measurements of $C3a$ and $C5a$ in the presence of $1$ g of zymosan published by Morad et al \cite{morad2015time} (Fig.
+The training for the lectin parameters, we used the experimental measurements of C3a and C5a in the presence of $1$ g of zymosan published by Morad et al \cite{morad2015time} (Fig.
 \ref{fig-fit} C and D).
@@
 We tested the predictive power of the reduced order human complement model with validation data sets not used during model training.
-Six validation data sets were used, three for $C3a$ and $C5a$ respectively at different zymosan concentrations.
+Six validation data sets were used, three for C3a and C5a respectively at different zymosan concentrations.
 All kinetic and control parameters were
-fixed for the validation simulations.
-The reduced order model predicted the $C3a$ and $C5a$ time-course profiles at a qualitative level (Fig. \ref{fig-prediction}).
+fixed for the validation simulations. The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig. \ref{fig-prediction}). [AND THEN WHAT??]
 
@@
 We conducted a Sobol's sensitivity analysis to estimate which parameters controlled the performance of the reduced order model.
-We calculated the sensitivity of the change in $C3a$ and $C5a$ profiles using the residuals between simulation and experimentally measured data for the cases of $0$ and $1 g$ zymosan (Fig.
+We calculated the sensitivity of the change in C3a and C5a profiles using the residuals between simulation and experimentally measured data for the cases of 0 and 1g zymosan (Fig.
 \ref{fig-SA}.
 For the cases in absence of zymosan where only the alternative pathway is active, we observed that only a few variables are responsible for the system response.
-For $C3a$ alternate, the sensitivity analysis found that $k_{c3b~basal}$ and $k_{degradationC3a}$ are the only sensitive parameters.
+For C3a alternate, the sensitivity analysis found that $k_{c3b~basal}$ and $k_{degradationC3a}$ are the only sensitive parameters.
 This gives us new insight in which of the parameters play a role in complement activation.
-Even though $AP~C3~Convertase$ is also responsible in the conversion of $C3$ and the production of $C3a$, the kinetic parameters that govern the equation was not sensitive at all.
-This elucidated that the activation of alternative pathway is more heavily governed by the spontaneous hydrolysis of $C3$ rather than the activity of $AP~C3~Convertase$.
-Surprisingly, closely examining the sensitive parameters that control $C5a$, in addition to the expected kinetic and control parameters related to the formation of $AP~C5~Convertase$, we observed that $k_C3~Convertase2$, the was previously not sensitive to $C3a$, to be sensitive in the formation of $C5a$.
-The $AP~C3~Convertase$ is a substrate required for the formation of $AP~C5~Convertase$ and the formation of $C3b$.
-The change in activity of $AP C3 Convertase$  will not drastically change the $C3a$ dynamics, but will effect $AP~C5a~Convertase$ formation and $C5a$ formation.
+Even though AP~C3~convertase is also responsible in the conversion of C3 and the production of C3a, the kinetic parameters that govern the equation was not sensitive at all.
+This elucidated that the activation of alternative pathway is more heavily governed by the spontaneous hydrolysis of C3 rather than the activity of AP~C3~Convertase.
+Surprisingly, closely examining the sensitive parameters that control C5a, in addition to the expected kinetic and control parameters related to the formation of AP~C5~Convertase, we observed that $k_C3~Convertase2$, the was previously not sensitive to C3a, to be sensitive in the formation of C5a.
+The AP~C3~Convertase is a substrate required for the formation of AP~C5~Convertase and the formation of C3b.
+The change in activity of AP C3 Convertase  will not drastically change the C3a dynamics, but will effect AP~C5a~Convertase formation and C5a formation.
 The our reduced order human complement model in combination with Sobol's sensitivity analysis was able to unravel important indirect parameter interaction.
-
 Our sensitivity analysis yielded expected results for the lectin pathway analyzes (Fig.
 \ref{fig-SA} (C and D)).
-One key difference that was observed between the sensitivity of the parameters between $C3a$ an $C5a$ was their respective degradation terms.
-The degradation constant of $C3a$ was sensitive between the two different cases of zymosan that was tested while the degradation constant of the $C5a$ was not sensitive.
+One key difference that was observed between the sensitivity of the parameters between C3a an C5a was their respective degradation terms.
+The degradation constant of C3a was sensitive between the two different cases of zymosan that was tested while the degradation constant of the C5a was not sensitive.
 We believe this different is attributed to the magnitude of the parameters and their respective concentrations.

### 2016-05-19 Paper_v2.tex
Updated part about ensembles

@@
 
-The reaction rate was written as the product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$), $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
-In this study, we used either saturation or mass action kinetics.
-The control term $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
-For each rate, we used a rule-based approach to select from competing control factors.
-If rate j was influenced by $1,\dots,m$ factors, we modeled this relationship as
-$v_{j} = \mathcal{I}_{j}\left(f_{1j}\left(\cdot\right),\hdots,f_{mj}\left(\cdot\right)\right)$
-where $0\leq f_{ij}\left(\cdot\right)\leq 1$ denotes a regulatory transfer function quantifying the influence of factor $i$ on rate $j$.
-The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
-variable. Each regulatory transfer function took the form:
+The reaction rates controlling formation C4a, C4b, C2a and C2b were written as a product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$) such that $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
+The kinetic term for these rates was modeled using saturation kinetics.
+The control term $0\leq v_{j}\leq 1$ for these reaction rates was modeled using regulatory transfer functions which took the form:
 \begin{equation}\label{eqn:control-factor}
@@
 \end{equation}where $\mathcal{Z}_{i}$ denotes the abundance factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
-In this study, we used $\mathcal{I}_{j}\in\left\{\max,\min\right\}$ \cite{pr3010138}. If a process has no modifying factors, $v_{j} = 1$.
 
@@
 \begin{equation}\label{eqn:rate-action}
-	\bar{r}_{j} = k_{j}^{max}\epsilon_{i}\prod_{s\in{m_{j}^{-}}}x_{s}
+	\bar{r}_{j} = k_{j}^{max}\prod_{s\in{m_{j}^{-}}}x_{s}
 \end{equation}
@@
 value for species $j$ at time $\tau$, and $\omega_{j}\left(\tau\right)$ denotes the experimental measurement variance for species $j$ at time $\tau$. The outer summation is with respect to
-time, while the inner summation is with respect to state. We minimized the model residual using Particle swarm optimization (PSO) \citep{PSO}.
-PSO uses a \textit{swarming} metaheuristic to explore parameter spaces.
-A strength of PSO is its ability to find the global minimum, even in the presence of potentially many local minima, by communicating the local
-error landscape experienced by each particle collectively to the swarm. Thus, PSO acts both as a local and a global search algorithm.
-For each iteration, particles in the swarm compute their local error by evaluating the model equations using their specific parameter vector realization.
-From each of these local points, a globally best error is identified. Both the local and global error
-are then used to update the parameter estimates of each particle using the rules:
-\begin{eqnarray}
-	\mathbf{\Delta}_{i} &=&\theta_{1}\mathbf{\Delta}_{i} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{k}_{i}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G} - \mathbf{k}_{i}\right) \\
-	\mathbf{k}_{i} &=& \mathbf{k}_{i} + \mathbf{\Delta}_{i}
-\end{eqnarray}where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the local best solution found by particle $i$, and
-$\mathcal{G}$ denotes the best solution found over the entire population of particles.
+time, while the inner summation is with respect to state.
+
+We minimized the model residual using Dynamic Optimization with Particle Swarms (DOPS).
+DOPS is a novel metaheuristic that combines multi swarm particle swarm optimization (PSO) with a greedy global optimization algorithm called dynamically dimensioned search (DDS).
+DOPS is much faster than conventional global optimizers and has the ability to find near optimal solutions for high dimensional systems within a relatively few function evaluations.
+It uses an adaptive switching strategy based on error convergence rates to switch from swarms search to DDS search.
+This enables it to find quickly, globally optimal or close to globally optimal solutions even in the presence of many local minima.
+In the swarm search, for each iteration the particles compute error within each sub-swarm by evaluating the model equations using their specific parameter vector realization.
+From each of these points within a sub-swarm a local best is identified.
+This along with the particle best within the sub-swarm $\mathcal{S}_{k}$ is used to update the parameter estimate for each particle using the following rules:
+
+\begin{equation}\label{eqn:update-rule}
+	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
+\end{equation}
+where ${z}_{i,j}$ is the parameter vector, $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm
+$\mathcal{S}_{k}$ for function evaluations $1\rightarrow j-1$, and
+$\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
+The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
+At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was used to initialize the DDS phase.
+For the DDS phase, the best parameter estimate was updated using the rule:
+\begin{equation}
+  \mathcal{G}_{new}({J})=\begin{cases}
+    \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
+    \mathcal{G}(\mathbf{J}), & \text{otherwise}.
+  \end{cases}
+\end{equation}
+where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
+and $\sigma$ denotes the perturbation amplitude:
+\begin{equation}
+	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
+\end{equation}
+
+where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
+The set $\mathbf{J}$ was constructed using a monotonically decreasing probability function $\mathcal{P}_{i}$ that represents a threshold for determining whether a specific dimension $j$ was perturbed or not.
+DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
+At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ for our model which we use for plotting best fits against the experimental data.
+We perturb this parameter vector to generate an ensemble of parameter vectors.
+The quality of parameter estimates was measured using goodness of fit (model residual).
+The DOPS routine was implemented in the MATLAB programming language.
+
+%From each of these local points, a globally best error is identified. Both the local and global error
+%are then used to update the parameter estimates of each particle using the rules:
+%\begin{eqnarray}
+%	\mathbf{\Delta}_{i} &=&\theta_{1}\mathbf{\Delta}_{i} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{k}_{i}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G} - \mathbf{k}_{i}\right) \\
+%	\mathbf{k}_{i} &=& \mathbf{k}_{i} + \mathbf{\Delta}_{i}
+%\end{eqnarray}where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the local best solution found by particle $i$, and
+%$\mathcal{G}$ denotes the best solution found over the entire population of particles.
 The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
-parameters \linebreak($\mathcal{K}\times{1}$). In thus study, we used $\left(\theta_{1},\theta_{2},\theta_{3}\right) = \left(1.0, 0.05564, 0.02886\right)$. The quality of parameter
-estimates was measured using goodness of fit (model residual). The particle swarm optimization routine was implemented in the Python programming language.
-All plots were made using the Matplotlib module of Python \citep{Matplotlib}.
-%The optimization and plotting code can be downloaded from http://www.varnerlab.org.
-%Experimental data,simulation code,parameter ensembles and initial conditions for each experimental simulation can be viewed and downloaded from https://github.com/jeffreyvarner/HybridCoagulationModel\_v1.
+%parameters \linebreak($\mathcal{K}\times{1}$). In thus study, we used $\left(\theta_{1},\theta_{2},\theta_{3}\right) = \left(1.0, 0.05564, 0.02886\right)$.
 
@@
 We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
-where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 42, so the total sensitivity indices were computed from
+where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from
 11,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \citep{SALIB}.

### 2016-05-21 Paper_v2.tex
Updated version

@@
 \section*{Abstract}
-Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
-In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
-The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
-We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
-Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
+Complement is a central part of innate immunity and plays a significant role in regulating the inflammatory response.
+In this study, we build a reduced order model of complement to study the human complement system.
+The key novelty of our approach is the use of Ordinary Differential Equations (ODEs) along with logical rules to capture the behavior of a complex biochemical network.
+Using this framework we constructed a model of complement that analyzed the dynamics of C3a and C5a when initiated through the alternate and lectin pathways.
 The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
-Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
-We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
-We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
+Thus, the model was an order of magnitude smaller that any existing model of complement that includes alternate and lectin pathways.
+
+%and includes more pathways than comparable purely ODE models in the literature.
+We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers, in the presence and absence of zymosan, and without the classical pathway.
+
+%We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
 Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
@@
 
-%We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
-Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems.
-For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.
-
 \vspace{0.1in}
 {\noindent \textbf{Keywords:}~Biochemical engineering, systems biology, reduced order models, complement system}
-
-% Extra abstract
-% Mathematical modeling of biological systems with multiple feed back loops is one such area where parameter estimation is a difficult non-linear optimization problem.
-This difficulty is further compounded when dealing with parameter vectors of high dimensions.
-
-%In this study, we present the dynamically dimensioned particle
-%a novel meta-heuristic approach that combines a variant of particle swarm optimization (PSO) along with dynamically dimensioned search (DDS) to obtain near optimal solutions of high dimensional biochemical networks within a relatively few function evaluations.
-%We use a particle swarm optimization technique that uses multi-swarms to generate candidate vectors which are then greedily updated using DDS by dynamically varying the perturbed parameter dimensions.
-We tested this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and also against DDS alone.
-The new algorithm outperforms all the other meta-heuristics on the coagulation model.
-The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data.
-We also performed this comparison on commonly used test functions (Ackley and Rosenbrock) for global optimization and found the same behavior.
-Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach.
-We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.
 
@@
 Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
-Delay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
+Decay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
 MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
@@
 However both these studies involve no validation studies with experimental data.
-Liu et al used analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters.
+Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters.
 Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters (Ref).
@@
 The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
-We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
+We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing or is complex.
 Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
@@
 Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
+
+%%%The start of the second paragraph
+
+This is the first ever model of complement that combines the alternate and non-alternate pathways and validates the model on experimental data.
+The formation of C3a without any initiator is controlled by the tick-over mechanism.
+
+Though it has been known that complement is a key player in regulating immune response and inflammation, very few mathematical models have been built to understand the dynamics of this network.
+To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of C3a and C5a on experimental data.
+The cleavage to C4 and C2 to form the classical/lectin C3 convertase involves a number of complex steps.
+The formation of this convertase and thereby the formation of C3a and C3b was modeled with  45 non-linear ODEs with 85 parameters by Liu and co-workers.
+The hybrid modeling framework allowed us to model the lectin mediated formation of C3 convertase using only 3 ODEs.
+We were able to successfully capture the formation of C3a at the maximum initiator (zymosan) concentration and also were able to qualitatively predict the dynamics of lectin mediated C3a formation at different initiator levels.
+The model is also surprisingly accurate in capturing the dynamics of C3a and C5a formed from the alternate pathway.
 

### 2016-05-21 Paper_v2.tex
Updated version

@@
 The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
-C5a acts as a bridge between innate and adaptive immunity and plays a very important role in regulating inflammation and coagulation.
 Complement activation takes places through three different pathways: the alternate, the classical and the lectin.
@@
 	The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the MAC complex.
-
 The activation of complement and formation of C5a and MAC complex is regulated at different points through a number of plasma and host cell proteins.
@@
 
-Given the complexity and importance, developing mathematical models of complement are crucial to understanding its dynamics.
-Complement models have typically been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
+Research over the past decade has shown that the role of complement extends beyond as a simple player in the immune system.
+It is an important network in maintaining homeostasis and malfunction of complement has been linked with several diseases like Alzheimers, acute macular degeneration (AMD), glaucoma, Parkinson's disease, multiple sclerosis, renal failure, schizophrenia, rheumatoid arthritis, cancer, sepsis and coagulopathy.
+The ability to maintain cross talk with other networks like coagulation, autonomous nervous response and the ability to regulate inflammation makes it important to understand complement in a more holistic perspective.
+[MORE LITERATURE]
+
+Thus given the complexity and importance of complement in influencing the human condition, developing models of complement within an integrative framework (that includes other biochemical networks) are crucial to understanding its role in human disease.
+Traditionally, complement models have been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
 Hirayama  et al.
@@
 However, these previous models were largely based upon mechanistic knowledge.
-However given the complexity of complement and its interactions with other networks like coagulation, autonomous nervous system, adaptive immunity it is unfeasible and computationally expensive to build such large mechanistic models.
+Given the complexity of complement and its interactions with other networks it is unfeasible and computationally expensive to build such large mechanistic models.
 In addition is much more difficult to experimentally interrogate the response of various complement proteins under different conditions.
@@
 
-In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
-The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
-We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing.
-Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
+In this study, we present a hybrid modeling approach to build a reduced order model of complement.
+The key innovation of this approach is the use of simple equations to capture the behavior of a complex biochemical network.
+The hybrid approach combines ODEs with logical rules to model biochemical processes that are complex or for which a complete mechanistic understanding is missing.
+We used this framework to capture dynamics of C3a and C5a formation in the lectin and alternative pathways.
 The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
-Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
-We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
-We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
-Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
-The reduced order hybrid approach produced a surprisingly predictive human complement model, much similar to the study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
-Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into roles of governing parameters that mediate the complement system.
-Deeply understanding how the main governing parameters change with respective to the different combinations of pathway activation will greatly aid in development of drugs for strategic therapeutic targets.
-Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
-
+Thus, the model was an order  of magnitude smaller and included more pathways than comparable ODE models in the literature.
+We estimated the model parameters from in vitro time series data of C3a and C5a from Morad and coworkers \cite{morad2015time}.
+Subsequently we validated the model on unseen C3a and C5a experimental data that were not used for model training.
+After validation, we performed a sensitivity analysis on the model to estimate which parameters were critical to model performance under different experimental conditions.
+Given its small size, the hybrid approach produced a surprisingly predictive human complement model, similar to an earlier study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
+Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into the roles of parameters that govern the complement system.
+A deeper understanding about how these parameters influence complement dynamics will greatly aid in the development of drugs for strategic therapeutic targets.
+Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.
 \clearpage
@@
 	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
+	In this study, we present a hybrid modeling approach to build a reduced order model of complement.
+The key innovation of this approach is the use of simple equations to capture the behavior of a complex biochemical network.
+The hybrid approach combines ODEs with logical rules to model biochemical processes that are complex or for which a complete mechanistic understanding is missing.
+We used this framework to capture dynamics of C3a and C5a formation in the lectin and alternative pathways.
+The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
+Thus, the model was an order  of magnitude smaller and included more pathways than comparable ODE models in the literature.
+We estimated the model parameters from in vitro time series data of C3a and C5a from Morad and coworkers \cite{morad2015time}.
+Subsequently we validated the model on unseen C3a and C5a experimental data that were not used for model training.
+After validation, we performed a sensitivity analysis on the model to estimate which parameters were critical to model performance under different experimental conditions.
+Given its small size, the hybrid approach produced a surprisingly predictive human complement model, similar to an earlier study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
+Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into the roles of parameters that govern the complement system.
+A deeper understanding about how these parameters influence complement dynamics will greatly aid in the development of drugs for strategic therapeutic targets.
+Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.
+
 	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
+Though the role of complement in immune response has been well known since long, there has been a paucity of mathematical models of complement.
+To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of downstream proteins like C5a using experimental data.
+The formation of C5a is preceded by formation of C3 convertases from the classical/lectin and alternate pathways.
+Liu and co-workers used 45 non-linear ODEs with 85 parameters to model the formation of C3a through the classical pathway.
+The hybrid modeling framework allowed us to model the lectin mediated formation of C3 convertase using only 4 ODEs.
+We were able to successfully capture the formation of C3a at the maximum initiator (zymosan) concentration and also were able to qualitatively predict the dynamics of lectin mediated C3a formation at different initiator levels.
+The model was also surprisingly accurate in capturing the quantitative dynamics of C3a and C5a formed from the alternate pathway with only 7 ODEs.
+The lag phase in the initiation of C5a followed by an accelerated production is also qualitatively similar to predicted C5a time profiles using a theoretical model of complement by Zewde et al with 107 equations.
+Similarly, our model was able to capture C3a formation from the alternate pathway that showed the same qualitative trends as Zewde et al.
+We also observe in our model simulation that the quantity of C3a produced in the alternate pathway is nearly 1000 times the quantity of  C5a produced.
+Though this is in agreement with the experimental data, it differs from the theoretical predictions by Zewde et al.
+who show that C3a is 1e8 times the C5a concentration.
+
 	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
@@
 \end{enumerate}
-
-In this study, we present an effective biochemical network modeling approach in building a reduced order model to study a complex biochemical network, the human complement system.
-The key innovation of our approach is the use of simple equations to capture the behavior of a complex biochemical network.
-We developed a hybrid modeling approach which combines ODEs and logical rules to model biochemical processes for which a complete mechanistic understanding is missing or is complex.
-Using this modeling framework we  incorporated computational simulation and experimental validation of the lectin and alternative pathway to examine C3 and C5 convertase formation and activity.
-The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
-Thus, the model was several orders of magnitude smaller and includes more pathways than comparable purely ODE models in the literature.
-We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers \cite{morad2015time}, in the presence and absence of zymosan, and without the classical pathway.
-We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
-Once validated, we performed sensitivity analysis on the model to estimate which parameters were critical to model performance in several conditions.
-The reduced order hybrid approach produced a surprisingly predictive human complement model, much similar to the study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
-Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into roles of governing parameters that mediate the complement system.
-Deeply understanding how the main governing parameters change with respective to the different combinations of pathway activation will greatly aid in development of drugs for strategic therapeutic targets.
-Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for treatment analysis or clinical screening in the medical field.
-
-%%%The start of the second paragraph
-
-This is the first ever model of complement that combines the alternate and non-alternate pathways and validates the model on experimental data.
-The formation of C3a without any initiator is controlled by the tick-over mechanism.
-
-Though it has been known that complement is a key player in regulating immune response and inflammation, very few mathematical models have been built to understand the dynamics of this network.
-To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of C3a and C5a on experimental data.
-The cleavage to C4 and C2 to form the classical/lectin C3 convertase involves a number of complex steps.
-The formation of this convertase and thereby the formation of C3a and C3b was modeled with  45 non-linear ODEs with 85 parameters by Liu and co-workers.
-The hybrid modeling framework allowed us to model the lectin mediated formation of C3 convertase using only 3 ODEs.
-We were able to successfully capture the formation of C3a at the maximum initiator (zymosan) concentration and also were able to qualitatively predict the dynamics of lectin mediated C3a formation at different initiator levels.
-The model is also surprisingly accurate in capturing the dynamics of C3a and C5a formed from the alternate pathway.
+%This suggests that our reduced order model's prediction for the alternate pathway have great potential application for the study of glomerular diseases that show strong correlation with disturbed alternative pathway activity (Pickering2011).
+In the diseases driven by dysfunctional alternative pathway, the activation of C5 appears to be the source that fueled the cycle by causing endothelial damage.
 

### 2016-05-22 Paper_v2.tex
More edits in introduction

@@
 	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
+
 	In this study, we present a hybrid modeling approach to build a reduced order model of complement.
@@
 	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
+
 Though the role of complement in immune response has been well known since long, there has been a paucity of mathematical models of complement.
 To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of downstream proteins like C5a using experimental data.
-The formation of C5a is preceded by formation of C3 convertases from the classical/lectin and alternate pathways.
-Liu and co-workers used 45 non-linear ODEs with 85 parameters to model the formation of C3a through the classical pathway.
-The hybrid modeling framework allowed us to model the lectin mediated formation of C3 convertase using only 4 ODEs.
-We were able to successfully capture the formation of C3a at the maximum initiator (zymosan) concentration and also were able to qualitatively predict the dynamics of lectin mediated C3a formation at different initiator levels.
+Liu and co-workers modeled formation of C3a through the classical pathway using 45 non-linear ODEs.
+The hybrid modeling framework, however allowed us to model the lectin mediated formation of C3a using only 5 ODEs.
+Though we do not capture all the interactions of initiation in detail, especially the cross-talk between lectin and classical pathways like Liu et al.
+we successfully captured C3a dynamics with respect to different initiator concentrations of the lectin pathway.
 The model was also surprisingly accurate in capturing the quantitative dynamics of C3a and C5a formed from the alternate pathway with only 7 ODEs.
@@
 We also observe in our model simulation that the quantity of C3a produced in the alternate pathway is nearly 1000 times the quantity of  C5a produced.
-Though this is in agreement with the experimental data, it differs from the theoretical predictions by Zewde et al.
-who show that C3a is 1e8 times the C5a concentration.
+Though this is in agreement with the experimental data  \cite{morad2015time}, it differs from the theoretical predictions by Zewde et al.
+who show that C3a is ${10^8}$ times the C5a concentration.
+The time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
+We see that the lag phase for generation progressively decreases with increasing concentration of initiator.
+Korotaevskiy et al.
+show a similar lag phase followed by accelerated production of C5a using a model of complement for much smaller time scales.
+Zewde et al.
+show a similar time profile for C5a generated in the alternate pathway as well.
+We do observe a similar trend in our model, however given the difference in time scales in our study we do not observe a very prominent lag phase.
+Taken together we surprisingly do very well in capturing the dynamics of key complement proteins and observe similar trends as in large mechanistic models.
 

### 2016-05-22 Paper_v2.tex
Changed refs

@@
 Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
-Complement was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
-Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases.
+It was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
+Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases \cite{}.
 The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
@@
 Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
-
-Research over the past decade has shown that the role of complement extends beyond as a simple player in the immune system.
+Research over the past decade has also shown that the role of complement extends beyond as a simple player in the immune system.
 It is an important network in maintaining homeostasis and malfunction of complement has been linked with several diseases like Alzheimers, acute macular degeneration (AMD), glaucoma, Parkinson's disease, multiple sclerosis, renal failure, schizophrenia, rheumatoid arthritis, cancer, sepsis and coagulopathy.
 The ability to maintain cross talk with other networks like coagulation, autonomous nervous response and the ability to regulate inflammation makes it important to understand complement in a more holistic perspective.
-[MORE LITERATURE]
 
@@
 The hybrid modeling framework, however allowed us to model the lectin mediated formation of C3a using only 5 ODEs.
-Though we do not capture all the interactions of initiation in detail, especially the cross-talk between lectin and classical pathways like Liu et al.
-we successfully captured C3a dynamics with respect to different initiator concentrations of the lectin pathway.
+Though we do not model all the interactions of initiation in detail, especially the cross-talk between lectin and classical pathways like Liu et al.
+we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
 The model was also surprisingly accurate in capturing the quantitative dynamics of C3a and C5a formed from the alternate pathway with only 7 ODEs.
@@
 who show that C3a is ${10^8}$ times the C5a concentration.
-The time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
+It is possible that the experimental system has trace amounts of C5a which leads to this difference.
+In our mode the time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
 We see that the lag phase for generation progressively decreases with increasing concentration of initiator.
 Korotaevskiy et al.
-show a similar lag phase followed by accelerated production of C5a using a model of complement for much smaller time scales.
+show a similar lag phase followed by accelerated production of C5a using a theoretical model of complement for much smaller time scales.
 Zewde et al.
 show a similar time profile for C5a generated in the alternate pathway as well.
-We do observe a similar trend in our model, however given the difference in time scales in our study we do not observe a very prominent lag phase.
+We do observe a similar trend in our model, however given the different time scale in our study we do not observe a very prominent lag phase.
+We also capture
+Our model also captures the max concentrations of C5a with different initiators and in the alternate pathway.
+However we do not capture the maximum concentration of C3a with low initiator levels.
+This is probably due to
+
 Taken together we surprisingly do very well in capturing the dynamics of key complement proteins and observe similar trends as in large mechanistic models.
@@
 \caption{Reduced order complement model training simulation for lectin and alternative pathway in presence of zymosan.
-Reduced order complement model parameters were estimated using dynamically dimensioned search (DDS) [Tolson and Shoemaker,2007,WRR] using the availability of zymosan as a function of lectin pathway initiation.
+Reduced order complement model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS) using the availability of zymosan as a function of lectin pathway initiation.
 Only parameters that govern the behavior of alternative pathway were allowed to vary when zymosan was not present.

### 2016-05-25 Paper_v2.tex
Working draft-Needs addition in discussion

@@
 To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of downstream proteins like C5a using experimental data.
-Liu and co-workers modeled formation of C3a through the classical pathway using 45 non-linear ODEs.
-The hybrid modeling framework, however allowed us to model the lectin mediated formation of C3a using only 5 ODEs.
+Liu and co-workers modeled formation of C3a through the classical pathway using 45 non-linear ODEs  \cite{liu2011computational}.
+The hybrid modeling framework, however, allowed us to model the lectin mediated formation of C3a using only 5 ODEs.
 Though we do not model all the interactions of initiation in detail, especially the cross-talk between lectin and classical pathways like Liu et al.
-we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
+\cite{liu2011computational} we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
 The model was also surprisingly accurate in capturing the quantitative dynamics of C3a and C5a formed from the alternate pathway with only 7 ODEs.
-The lag phase in the initiation of C5a followed by an accelerated production is also qualitatively similar to predicted C5a time profiles using a theoretical model of complement by Zewde et al with 107 equations.
-Similarly, our model was able to capture C3a formation from the alternate pathway that showed the same qualitative trends as Zewde et al.
+The lag phase in the initiation of C5a is followed by an acceleration in the production of C5a.
+This is also qualitatively similar to the predicted C5a time profiles using a theoretical model of complement with 107 equations \cite{zewde2016quantitative}.
+Similarly, our model was able to capture C3a formation from the alternate pathway that showed the same qualitative trends as Zewde et al \cite{zewde2016quantitative}.
 We also observe in our model simulation that the quantity of C3a produced in the alternate pathway is nearly 1000 times the quantity of  C5a produced.
-Though this is in agreement with the experimental data  \cite{morad2015time}, it differs from the theoretical predictions by Zewde et al.
-who show that C3a is ${10^8}$ times the C5a concentration.
+Though this is in agreement with the experimental data  \cite{morad2015time}, it differs from the theoretical predictions by Zewde et al.\cite{zewde2016quantitative} who show that C3a is ${10^8}$ times the C5a concentration.
 It is possible that the experimental system has trace amounts of C5a which leads to this difference.
-In our mode the time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
-We see that the lag phase for generation progressively decreases with increasing concentration of initiator.
+In our model the time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
+We see that the lag phase for generation progressively decreases with increasing concentration of the initiator.
 Korotaevskiy et al.
-show a similar lag phase followed by accelerated production of C5a using a theoretical model of complement for much smaller time scales.
+\cite{korotaevskiy2009non} show a similar lag phase followed by an accelerated production of C5a using a theoretical model of complement for much smaller time scales.
 Zewde et al.
-show a similar time profile for C5a generated in the alternate pathway as well.
-We do observe a similar trend in our model, however given the different time scale in our study we do not observe a very prominent lag phase.
-We also capture
-Our model also captures the max concentrations of C5a with different initiators and in the alternate pathway.
-However we do not capture the maximum concentration of C3a with low initiator levels.
-This is probably due to
+\cite{zewde2016quantitative} show a similar time profile for C5a generated in the alternate pathway as well.
+We do observe a similar trend in our model.
+However given the different time scale in our study we do not observe a very prominent lag phase.
+Our model also captures the maximum concentrations of C5a with different initiators and in the alternate pathway.
+However we do not capture the maximum concentration of C3a with low initiator levels though we capture the overall trend which includes the initiation and amplification phases.
+One of the possible reasons for this could be the non-inclusion of C2-by pass pathway which further accelerates production of C3a without the involvement of a C3 convertase.
+Currently the C3a in the model is generated only through the activity of a C3 convertase.
+Incorporating this additional step within the reduced order modeling framework would be a future direction that we need to consider.
+Taken together we surprisingly do very well in capturing the dynamics of key complement proteins and observe similar trends as in large mechanistic models.
 
-Taken together we surprisingly do very well in capturing the dynamics of key complement proteins and observe similar trends as in large mechanistic models.
+%%% We talk about complement therapeutics here in relation with sensitivity analysis
+
+The global sensitivity analysis of our reduced order model provides interesting insights into the regulation of C3a and C5a which are important therapeutic targets.
+We see that the formation of C5a in the absence of an initiator is sensitive to the formation of basal C3a and C3b.
+%Though it is known complement malfunction causes a number of diseases \cite{ricklin2007complement} developing complement specific therapeutics has been challenging\cite{ricklin2007complement}.
+Currently there are only four approved drugs targeting complement \cite{morgan2015complement}.
+Eculizumab targets C5, Berinert, Ruconest and Cinryze target C1.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v5.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.

### 2016-05-25 Paper_v2.tex
Paragraph updated about therapeutic targets

@@
 The global sensitivity analysis of our reduced order model provides interesting insights into the regulation of C3a and C5a which are important therapeutic targets.
-We see that the formation of C5a in the absence of an initiator is sensitive to the formation of basal C3a and C3b.
+We see that the formation of C5a in the absence of an initiator is the most sensitive to the formation of basal C3b through the tick over mechanism.
+This mechanism could be a potential target for C5a inhibitors especially in cases of autoimmune disorders where there is an absence of initiators of the lectin or the classical pathway.
+In presence of an initiator we see that C5a formation is primarily sensitive to the lectin initiation parameters and the kinetic parameters that govern the conversion of C5 to C5a and C5b.
+This result agrees well with the current protease inhibitors targeting initiating complexes that includes mannose-associated serine proteases 1 and  2 (MASP-1,2).
+The most commonly used anti-complement drug eculizumab, targets C5 protein which is cleaved to form C5a.
+Our sensitivity analysis shows that kinetic parameters governing C5 conversion are sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
+The formation of basal C3b is also a sensitive parameter in the formation of C3a through the alternate pathway.
+Thus this mechanism can act as a target for both C3a and C5a inhibitors.
+The lectin initiated formation of C3a shows a number of sensitive parameters.
+This includes the lectin initiation parameters that control the formation of C5a along with C3 convertase inhibition using C4BP and the parameter governing C3 convertase.
+All these mechanisms involve potential drug targets and literature shows that C4BP, C3 convertase inhibitors are the being developed as the next generation inhibitors of complement \cite{morgan2015complement}.
+Taken together it is surprising to see that reduced order model has the ability to identify mechanisms for potential drug targets that are in agreement with existing literature.
 %Though it is known complement malfunction causes a number of diseases \cite{ricklin2007complement} developing complement specific therapeutics has been challenging\cite{ricklin2007complement}.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_v5.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_may24.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.

### 2016-05-25 Paper_v0.tex
Merged changes to intro refs results and figs

--- 2016-05-22 Paper_v2.tex
@@
 Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
-It was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
-Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases \cite{}.
+Complement was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
+Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases.
 The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
+C5a acts as a bridge between innate and adaptive immunity and plays a very important role in regulating inflammation and coagulation.
 Complement activation takes places through three different pathways: the alternate, the classical and the lectin.
-Each of these pathways involves a different initiator signal that leads to the formation of a serine protease called C5 convertase which cleaves an inactive protein called C5 to form C5a and C5b.
+Each of these pathways involves a different initiator signal that leads to a cascade of downstream reactions in the complement system.
 The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
@@
 The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to the carbohydrates on the surfaces of bacterial pathogens.
-This bound complex in turn cleaves C4 and C2 and leads to the production of C4bC2a.
-The alternate pathway involves a 'tickover' mechanism in which a protein called C3 is hydrolyzed to form C3b.
+This bound complex in turn cleaves C4 and C2 and leads to the formation of C4bC2a.
+The alternate pathway involves a 'tickover' mechanism in which  C3 is hydrolyzed to form C3b.
 In presence of foreign pathogens C3b binds to these surfaces and recruits  additional factors called factor B and factor D that lead to the formation of alternate C3 convertase - C3bBb.
@@
 The C5 convertases then cleave C5 to form C5a and C5b respectively.
-	The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the MAC complex.
-The activation of complement and formation of C5a and MAC complex is regulated at different points through a number of plasma and host cell proteins.
+The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the membrane attack complex (MAC).
+The activation of complement and formation of C5a and MAC  is regulated at different points through a number of plasma and host cell proteins.
 The initiation of the classical pathway through the attachment of C1 to an antibody is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
@@
 Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
-Decay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
+Delay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
 MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
-Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
+Carboxypeptidase-N, a regulator of inflammation is a plasma zinc metalloprotease that cleaves carboxyl-terminal arginines and lysines of complement complement components and  anaphylatoxins such as C3a, C4a, and C5a.
+\cite{liszewski1995control}.
+
 Research over the past decade has also shown that the role of complement extends beyond as a simple player in the immune system.
-It is an important network in maintaining homeostasis and malfunction of complement has been linked with several diseases like Alzheimers, acute macular degeneration (AMD), glaucoma, Parkinson's disease, multiple sclerosis, renal failure, schizophrenia, rheumatoid arthritis, cancer, sepsis and coagulopathy.
-The ability to maintain cross talk with other networks like coagulation, autonomous nervous response and the ability to regulate inflammation makes it important to understand complement in a more holistic perspective.
+Complement has been shown to contribute to homeostasis by inducing growth factors involved in tissue repair \cite{ricklin2010complement}.
+The malfunction of complement has been linked with several diseases including Alzheimers, glaucoma, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+Complement has also been shown to play both a positive and negative role in certain cancers; attacking tumor cells with altered surface patterns in some cases potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
+Several networks maintain cross talk with complement including coagulation, autonomous nervous response and the ability to regulate inflammation \cite{ricklin2013complement}.
+The system's involvement in a variety of beneficial regulatory processes coupled with the role it can play in disease makes it important to understand complement in a more holistic perspective.
 
-Thus given the complexity and importance of complement in influencing the human condition, developing models of complement within an integrative framework (that includes other biochemical networks) are crucial to understanding its role in human disease.
+Thus given the complexity and importance of complement, developing models of the system within an integrative framework (that can include other biochemical networks) is crucial to understanding its role in human disease.
 Traditionally, complement models have been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
 Hirayama  et al.
-(ref) used a system of linear ODEs to model the classical pathway of complement.
-Korotaevskiy and co-workers (ref) built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways.
+used a system of linear ODEs to model the classical pathway of complement \cite{hirayama1996linear}.
+Korotaevskiy and co-workers built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways \cite{korotaevskiy2009non}.
 However both these studies involve no validation studies with experimental data.
-Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters.
-Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters (Ref).
+Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
+Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters \cite{zewde2016quantitative}.
 This model delineated the response of complement on a host cell and a foreign antigen.
@@
 All kinetic and control parameters were
-fixed for the validation simulations. The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig. \ref{fig-prediction}). [AND THEN WHAT??]
+fixed for the validation simulations.
+The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig.
+\ref{fig-prediction}).
+For the prediction of the $C3a$ at the three different inducer concentrations, our model was able to accurately capture the dynamic behaviors at a qualitative level, including the time the concentration of $C3a$ plateaus.
+For the prediction for $C5a$, our model was able to capture the behavior at a quantitative level for all three inducer concentrations.
+However, the our model was not able to capture the concave down curvature for the cases of $0.001~g$ and $0.01~g$ inducer.
+We believe some of the shortcomings of the prediction may be attributed by our reduced complexity in the modeling of $C4B$ interactions in our system.
+$C4BP$ interaction was modeled as irreversible mass action kinetics, in reality the binding of $C4BP$ onto other complement components are reversible.
+Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the $C4BP$ is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
+
+%This phenomenon may be a part of evolutionary adaptation since C5a is a key factor in the development of many inflammatory diseases such as sepsis, rheumatoid arthritis, and systemic lupus erythemotosis.
 
@@
 \subsection*{Global sensitivity analysis of model performance}
-We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}.
+We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{zhang2015sobol}.
 We computed the total sensitivity index of each parameter relative to four performance objectives, each objective was based on the sum of squared errors between model and experimental data for C3a alternate, C5a alternate, C3a lectin, and C5a lectin simulations.
 We established the sampling bounds for each parameter from the minimum and maximum value of that parameter in the parameter set ensemble.
-We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
+We used the sampling method of Saltelli \textit{et al.} \citep{saltelli2010variance} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
 where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from

### 2016-05-25 Paper_v2.tex
Edited legends

@@
 \section*{Introduction}
-Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
+
+%Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
 It was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
@@
 MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
-Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
+%Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
+
+Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
+Complement was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
+Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases \cite{walport2001complement}.
+The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
+C5a acts as a bridge between innate and adaptive immunity and plays a very important role in regulating inflammation and coagulation \cite{sarma2011complement}.
+Complement activation takes places through three different pathways: the alternate, the classical and the lectin.
+Each of these pathways involves a different initiator signal that leads to a cascade of downstream reactions in the complement system.
+The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
+A multimeric protein complex C1 binds to the antigen-antibody complex and undergoes a conformational change.
+This activated complex cleaves proteins C4 and C2 to C4a, C4b, C2a and C2b respectively.
+C4a and C2b combine to form a protease C4bC2a also known as the classical C3 convertase.
+The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to the carbohydrates on the surfaces of bacterial pathogens.
+This bound complex in turn cleaves C4 and C2 and leads to the formation of C4bC2a.
+The alternate pathway involves a 'tickover' mechanism in which  C3 is hydrolyzed to form C3b.
+In presence of foreign pathogens C3b binds to these surfaces and recruits  additional factors called factor B and factor D that lead to the formation of alternate C3 convertase - C3bBb \cite{pangburn1984alternative}.
+The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
+The classical and alternate C3 convertases recruit C3, Factor B and Factor D to form classical C5 convertase (C4bC2aC3b) and alternate C5 convertase (C3bBbc3B) respectively.
+The C5 convertases then cleave C5 to form C5a and C5b respectively.
+The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the membrane attack complex (MAC) \cite{ricklin2010complement}.
+
+The activation of complement and formation of C5a and MAC  is regulated at different points through a number of plasma and host cell proteins.
+The initiation of the classical pathway through the attachment of C1 to an antibody is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
+C1-Inh irreversibly binds to and deactivates the active subunits of component C1 to prevent spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
+The serum and host-tissue regulation of the upstream elements of the complement system is also achieved through the binding of C4 binding protein (C4BP) to C4b and through the binding of factor H to C3b \cite{blom2001structural}.
+These proteins are also capable of binding their respective components in the convertase form.
+Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
+Delay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
+MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
+Carboxypeptidase-N, a regulator of inflammation is a plasma zinc metalloprotease that cleaves carboxyl-terminal arginines and lysines of complement complement components and  anaphylatoxins such as C3a, C4a, and C5a.
+\cite{liszewski1995control}.
+
 Research over the past decade has also shown that the role of complement extends beyond as a simple player in the immune system.
-It is an important network in maintaining homeostasis and malfunction of complement has been linked with several diseases like Alzheimers, acute macular degeneration (AMD), glaucoma, Parkinson's disease, multiple sclerosis, renal failure, schizophrenia, rheumatoid arthritis, cancer, sepsis and coagulopathy.
-The ability to maintain cross talk with other networks like coagulation, autonomous nervous response and the ability to regulate inflammation makes it important to understand complement in a more holistic perspective.
+Complement has been shown to contribute to homeostasis by inducing growth factors involved in tissue repair \cite{ricklin2010complement}.
+The malfunction of complement has been linked with several diseases including Alzheimers, glaucoma, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+Complement has also been shown to play both a positive and negative role in certain cancers; attacking tumor cells with altered surface patterns in some cases potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
+Several networks maintain cross talk with complement including coagulation, autonomous nervous response and the ability to regulate inflammation \cite{ricklin2013complement}.
+The system's involvement in a variety of beneficial regulatory processes coupled with the role it can play in disease makes it important to understand complement in a more holistic perspective.
 
-Thus given the complexity and importance of complement in influencing the human condition, developing models of complement within an integrative framework (that includes other biochemical networks) are crucial to understanding its role in human disease.
+Thus given the complexity and importance of complement, developing models of complement within an integrative framework (that includes other biochemical networks) are crucial to understanding its role.
 Traditionally, complement models have been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
 Hirayama  et al.
-(ref) used a system of linear ODEs to model the classical pathway of complement.
-Korotaevskiy and co-workers (ref) built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways.
+used a system of linear ODEs to model the classical pathway of complement \cite{hirayama1996linear}.
+Korotaevskiy and co-workers built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways \cite{korotaevskiy2009non}.
 However both these studies involve no validation studies with experimental data.
-Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters.
-Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters (Ref).
+Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
+Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters \cite{zewde2016quantitative}.
 This model delineated the response of complement on a host cell and a foreign antigen.
@@
 Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.
+
 \clearpage
@@
 All kinetic and control parameters were
-fixed for the validation simulations. The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig. \ref{fig-prediction}). [AND THEN WHAT??]
+fixed for the validation simulations.
+The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig.
+\ref{fig-prediction}).
+For the prediction of the $C3a$ at the three different inducer concentrations, our model was able to accurately capture the dynamic behaviors at a qualitative level, including the time the concentration of $C3a$ plateaus.
+For the prediction for $C5a$, our model was able to capture the behavior at a quantitative level for all three inducer concentrations.
+However, the our model was not able to capture the concave down curvature for the cases of $0.001~g$ and $0.01~g$ inducer.
+We believe some of the shortcomings of the prediction may be attributed by our reduced complexity in the modeling of $C4B$ interactions in our system.
+$C4BP$ interaction was modeled as irreversible mass action kinetics, in reality the binding of $C4BP$ onto other complement components are reversible.
+Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the $C4BP$ is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
 
@@
 We calculated the sensitivity of the change in C3a and C5a profiles using the residuals between simulation and experimentally measured data for the cases of 0 and 1g zymosan (Fig.
-\ref{fig-SA}.
+\ref{fig-SA}).
 For the cases in absence of zymosan where only the alternative pathway is active, we observed that only a few variables are responsible for the system response.
@@
 In presence of an initiator we see that C5a formation is primarily sensitive to the lectin initiation parameters and the kinetic parameters that govern the conversion of C5 to C5a and C5b.
-This result agrees well with the current protease inhibitors targeting initiating complexes that includes mannose-associated serine proteases 1 and  2 (MASP-1,2).
-The most commonly used anti-complement drug eculizumab, targets C5 protein which is cleaved to form C5a.
+This result agrees well with the current protease inhibitors targeting initiating complexes that includes mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{morgan2015complement}.
+The most commonly used anti-complement drug eculizumab \cite{morgan2015complement}, targets C5 protein which is cleaved to form C5a.
 Our sensitivity analysis shows that kinetic parameters governing C5 conversion are sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
@@
 \subsection*{Global sensitivity analysis of model performance}
-We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}.
+We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \cite{SOBOL_METHOD}.
 We computed the total sensitivity index of each parameter relative to four performance objectives, each objective was based on the sum of squared errors between model and experimental data for C3a alternate, C5a alternate, C3a lectin, and C5a lectin simulations.
 We established the sampling bounds for each parameter from the minimum and maximum value of that parameter in the parameter set ensemble.
-We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
+We used the sampling method of Saltelli \textit{et al.} \cite{saltelli2010variance} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
 where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from
-11,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \citep{SALIB}.
+11,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/ComplementSystemSchematic_v2.pdf}
-\caption{Simplified schematic representation of the human complement system.
-The complement cascade is activated through any one, or more, of the three pathways: classical, lectin, and alternate pathway.
-The classical pathway is activated by the complex formation of $C1q$, $C1r$, and $C1s$ by the recogniztion of antibody:antigen complexes.
-Similarly, the lecin pathway is initiation by binding mannan-binding lectin to mannose on pathogen surfaces.
-Lastly, the alternative pathway is activated when a complement component is spontaneously bound to the surface of the pathogen of virus.
-The activation from the three pathways creates a cascades of reactions that forms the proteases, $C3$ Convertase that cleaves $C3$ into $C3a$, and $C3b$, the main effector molecule of the complement system.
-$C3b$ can find to a $C3$ convertase and form a $C5$ convertase that cleaves $C5$ into $C5a$, and $C5b$ that undergoes a series of reactions to form the membrane attach complex ($MAC$).}\label{fig-schematic}
+\includegraphics[width=1.0\textwidth]{./figs/Fig1_Schematic_v2.pdf}
+\caption{Simplified schematic of the human complement system.
+The complement cascade is activated through any one, or more, of the three pathways:  the classical, the lectin, and the alternate pathways.
+The classical pathway is activated by the binding of C1 complex through the C1q subunit to the IgG or IgM immune complex.
+This binding leads to conformational changes in the C1 complex that leads to the activation of C1r and C1s subunits.
+Activated C1-antibody complex cleaves C4 and C2 to form the classical C3 convertase.
+The lectin pathway is initiated by the binding mannose-binding lectins (MBL) and ficolins to carbohydrate moieties on the pathogen surfaces.
+This results in the formation mannose-binding lectin-associated serine proteases (MASPs).
+The MBL-MASP complex cleaves C4 and C2 to form the lectin C3 convertase.
+The alternative pathway is activated through a spontaneous tick-over mechanism by the hydrolysis of C3 to form fluid phase C3 convertase.
+The C3 convertases  cleaves C3 into C3a, and C3b.
+C3b combines with C4b and C2a to form classical C5 convertase (C4bC3aC3b).
+The C3b binds with Factor B to form the alternate C5 convertase (C3bBbC3b) .
+The C5 convertases cleave C5 into C5a, and C5b that undergoes a series of reactions to form the membrane attack complex (MAC).}\label{fig-schematic}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure2_Fit.pdf}
-\caption{Reduced order complement model training simulation for lectin and alternative pathway in presence of zymosan.
-Reduced order complement model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS) using the availability of zymosan as a function of lectin pathway initiation.
-Only parameters that govern the behavior of alternative pathway were allowed to vary when zymosan was not present.
-Our model training was conducted in a hierarchal fashion where the alternate parameters were trained and then used and fixed in estimating the lectin parameters.
-The red line shows the best-fit parameter, the black lines denotes the simulated mean value of $C3a$ or $C5a$ for a 50 parameter set ensemble.
-The shaded region denotes the distribution of $C3a$ and $C5a$ of the ensemble.}\label{fig-fit}
+\caption{Reduced order complement model training simulations.
+Reduced order complement model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS).
+The model was trained against experimental data from Shaw and co-workers \cite{morad2015time} in the presence and absence of zymosan.
+The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin initiated pathway with 1g zymosan (\textbf{C}--\textbf{D}).
+The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of $C3a$ or $C5a$ for 50 independent particles.
+The shaded region denotes 99 \% confidence interval on the simulated mean concentration of $C3a$ or $C5a$ (uncertainty in the model simulation).
+All initial concentrations of complement proteins are at human serum levels unless otherwise noted.}\label{fig-fit}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure3_Predictions_2.pdf}
-\caption{Reduced order complement model predictions of lectin and alternative pathway in presence of zymosan.
-(A-F) Simulation of complement dynamics in the presence of zymosan were conducted for a range of trigger values ($0.1$, $0.01$, and $0.001$ grams of zymosan).
-The time-course profiles of $C3a$ and $C5a$ under three different zymosan concentrations were simulated using 50 ensembles of trained parameter sets against experimental data of Shaw et al [REF].
-The red curve represents the best fit parameter, grey shaded region denotes the prediction results from 50 ensembles of parameter sets, and the black curve is the mean of the ensemble.
-All complement protein and factor initial concentrations coincided with human serum levels unless otherwise noted.}\label{fig-prediction}
+\caption{Reduced order complement model predictions vs experimental data for C3a and C5a generated in the lectin pathway.
+The reduced order coagulation model parameter estimates were tested against data not used during model training.
+Simulations of C3a and C5a generated in the lectin pathway using different levels of zymosan ($0.1$, $0.01$, and $0.001$ grams of zymosan) were compared with the corresponding experimental data (\textbf{A}--\textbf{F}).
+The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of $C3a$ or $C5a$ for 50 independent particles.
+The shaded region denotes 99 \% confidence interval on the simulated mean concentration of $C3a$ or $C5a$ (uncertainty in the model simulation).
+All initial concentrations of complement proteins are at human serum levels unless otherwise noted.
+}\label{fig-prediction}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_may24.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_may25.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.

### 2016-05-25 Paper_v2.tex
Abstract and introduction

@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Reduced order modeling and analysis of the human complement subsystem}}}
+{\par\centering\textbf{\Large {Reduced order modeling and analysis of the human complement system}}}
 \vspace{0.05in}
@@
 \vspace{0.1in}
-{\par \centering \textbf{Running Title:}~Reduced order model of complement}
+{\par \centering \textbf{Running Title:}~A reduced order model of complement}
 \vspace{0.1in}
@@
 \section*{Abstract}
-Complement is a central part of innate immunity and plays a significant role in regulating the inflammatory response.
-In this study, we build a reduced order model of complement to study the human complement system.
-The key novelty of our approach is the use of Ordinary Differential Equations (ODEs) along with logical rules to capture the behavior of a complex biochemical network.
-Using this framework we constructed a model of complement that analyzed the dynamics of C3a and C5a when initiated through the alternate and lectin pathways.
-The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
-Thus, the model was an order of magnitude smaller that any existing model of complement that includes alternate and lectin pathways.
+Complement is a central part of innate immunity which plays a significant role in the inflammatory response, and many other disease processes.
+In this study, we analyzed an ensemble of experimentally validated reduced order complement models.
+Our reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
+We used this framework to capture the dynamics of C3a and C5a formation in the lectin and alternative pathways.
+The reduced order model consisted of only 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
+Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
+After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
 
@@
 %Proteins C3a, C4a, and C5a are inactivated or reduced in activity by carboxypeptidase-N \cite{liszewski1995control}.
-
-Complement is a central part of innate immunity and plays a very significant role in regulating the inflammatory response.
-Complement was first discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies.
-Complement is mediated through a set of approximately 30-35 soluble and cell surface proteases \cite{walport2001complement}.
-The central process in complement activation involves the formation of Membrane Attack Complex (MAC) and a protein called C5a.
-C5a acts as a bridge between innate and adaptive immunity and plays a very important role in regulating inflammation and coagulation \cite{sarma2011complement}.
-Complement activation takes places through three different pathways: the alternate, the classical and the lectin.
-Each of these pathways involves a different initiator signal that leads to a cascade of downstream reactions in the complement system.
-The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
-A multimeric protein complex C1 binds to the antigen-antibody complex and undergoes a conformational change.
-This activated complex cleaves proteins C4 and C2 to C4a, C4b, C2a and C2b respectively.
-C4a and C2b combine to form a protease C4bC2a also known as the classical C3 convertase.
-The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to the carbohydrates on the surfaces of bacterial pathogens.
-This bound complex in turn cleaves C4 and C2 and leads to the formation of C4bC2a.
-The alternate pathway involves a 'tickover' mechanism in which  C3 is hydrolyzed to form C3b.
-In presence of foreign pathogens C3b binds to these surfaces and recruits  additional factors called factor B and factor D that lead to the formation of alternate C3 convertase - C3bBb \cite{pangburn1984alternative}.
-The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
-The classical and alternate C3 convertases recruit C3, Factor B and Factor D to form classical C5 convertase (C4bC2aC3b) and alternate C5 convertase (C3bBbc3B) respectively.
-The C5 convertases then cleave C5 to form C5a and C5b respectively.
-The cleavage of C5 is followed by a series of sequential cleavages of proteins C6, C7, C8 and C9 that combine with C5b to form the membrane attack complex (MAC) \cite{ricklin2010complement}.
-
-The activation of complement and formation of C5a and MAC  is regulated at different points through a number of plasma and host cell proteins.
-The initiation of the classical pathway through the attachment of C1 to an antibody is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
-C1-Inh irreversibly binds to and deactivates the active subunits of component C1 to prevent spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
-The serum and host-tissue regulation of the upstream elements of the complement system is also achieved through the binding of C4 binding protein (C4BP) to C4b and through the binding of factor H to C3b \cite{blom2001structural}.
-These proteins are also capable of binding their respective components in the convertase form.
-Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
-Delay accelerating factor (DAF or CD55) is able to recognize and dissociate both convertases \cite{lukacik2004complement}.
-MAC is inhibited by vitronectin and clusterin in the plasma and CD59 at the host surface.
-Carboxypeptidase-N, a regulator of inflammation is a plasma zinc metalloprotease that cleaves carboxyl-terminal arginines and lysines of complement complement components and  anaphylatoxins such as C3a, C4a, and C5a.
-\cite{liszewski1995control}.
-
-Research over the past decade has also shown that the role of complement extends beyond as a simple player in the immune system.
-Complement has been shown to contribute to homeostasis by inducing growth factors involved in tissue repair \cite{ricklin2010complement}.
-The malfunction of complement has been linked with several diseases including Alzheimers, glaucoma, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
-Complement has also been shown to play both a positive and negative role in certain cancers; attacking tumor cells with altered surface patterns in some cases potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
-Several networks maintain cross talk with complement including coagulation, autonomous nervous response and the ability to regulate inflammation \cite{ricklin2013complement}.
-The system's involvement in a variety of beneficial regulatory processes coupled with the role it can play in disease makes it important to understand complement in a more holistic perspective.
-
-Thus given the complexity and importance of complement, developing models of complement within an integrative framework (that includes other biochemical networks) are crucial to understanding its role.
-Traditionally, complement models have been formulated as linear or non-linear Ordinary Differential Equation (ODE) systems.
-Hirayama  et al.
-used a system of linear ODEs to model the classical pathway of complement \cite{hirayama1996linear}.
-Korotaevskiy and co-workers built a theoretical model of complement using a system of non-linear ODEs that included classical, lectin and alternate pathways \cite{korotaevskiy2009non}.
-However both these studies involve no validation studies with experimental data.
-Liu et al analyzed the formation of classical and lectin C3 convertases and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
-Recently, Zewde and co-workers built a detailed mechanistic model of alternative complement activation was built using 107 ODEs and 74 kinetic parameters \cite{zewde2016quantitative}.
-This model delineated the response of complement on a host cell and a foreign antigen.
-However, these previous models were largely based upon mechanistic knowledge.
-Given the complexity of complement and its interactions with other networks it is unfeasible and computationally expensive to build such large mechanistic models.
+%The system's involvement in a variety of beneficial regulatory processes coupled with the role it can play in disease makes it important to understand complement in a more holistic perspective.
+%Given the complexity of complement and its interactions with other networks it is unfeasible and computationally expensive to build such large mechanistic models.
 In addition is much more difficult to experimentally interrogate the response of various complement proteins under different conditions.
@@
 
-In this study, we present a hybrid modeling approach to build a reduced order model of complement.
-The key innovation of this approach is the use of simple equations to capture the behavior of a complex biochemical network.
-The hybrid approach combines ODEs with logical rules to model biochemical processes that are complex or for which a complete mechanistic understanding is missing.
-We used this framework to capture dynamics of C3a and C5a formation in the lectin and alternative pathways.
-The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
+Complement is a central part of innate immunity which plays a significant role in the inflammatory response.
+Complement was discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies [REF].
+However, research over the past decade has shown the importance of complement extends well beyond innate immunity.
+For example, complement contributes to tissue homeostasis by inducing growth factors involved in tissue repair \cite{ricklin2010complement}.
+Complement malfunctions have been linked with several diseases including Alzheimers, glaucoma, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+Complement can also play both a positive and negative role in certain cancers; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
+Several other important biochemical networks are integrated with complement including the coagulation cascade, the autonomous nervous response and the ability to regulate inflammation \cite{ricklin2013complement}.
+Thus, complement is an important system involved in a variety of both beneficial and potentially harmful functions in the body.
+
+Complement is mediated by over 30 soluble and cell surface proteins that are present as inactive forms in the circulation \cite{walport2001complement}.
+The central output of complement activation is the formation of the Membrane Attack Complex (MAC) and a key protein called C5a.
+The membrane attack complex forms transmembrane channels which disrupt the cell membrane of targeted cells, leading to cell lysis and death [REF].
+On the other hand, the C5a protein acts as a bridge between innate and adaptive immunity, and plays an important role in regulating inflammation and coagulation \cite{sarma2011complement}.
+Complement activation takes places through three pathways: the alternate, the classical and the lectin binding pathway.
+Each of these pathways involves a different initiator signal which leads to a cascade of downstream events in the complement system.
+The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
+A multimeric protein complex C1 binds to the antigen-antibody complex and undergoes a conformational change.
+This activated complex then cleaves complement proteins C4 and C2 into C4a, C4b, C2a and C2b respectively.
+The C4a and C2b fragments combine to form the C4bC2a protease, also known as the classical C3 convertase.
+The lectin binding pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
+This bound complex in turn cleaves C4 and C2, leading to the formation of C4bC2a.
+The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is hydrolyzed to form C3b.
+In presence of pathogens, the C3b fragment binds foreign surfaces and recruits the additional proteins, factor B and factor D, which lead to the formation of C3bBb, the alternate C3 convertase \cite{pangburn1984alternative}.
+The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
+The classical and alternate C3 convertases recruit C3, Factor B and Factor D to form the classical C5 convertase (C4bC2aC3b), and alternate C5 convertase (C3bBbc3B) respectively.
+The C5 convertases then cleave C5 to form the C5a and C5b fragments.
+The cleavage of C5 is followed by a series of sequential cleavage steps involving the C6, C7, C8 and C9 complement proteins
+which combine with C5b to form the membrane attack complex \cite{ricklin2010complement}.
+
+Complement activation is regulated by many plasma and host cell proteins.
+The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
+C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
+Regulation of the upstream elements of complement is also achieved through the interaction of the C4 binding protein (C4BP) with C4b, as well as
+through the interaction of factor H with C3b \cite{blom2001structural}.
+These regulatory proteins are also capable of binding their respective targets in the convertase form as well.
+Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
+Delay accelerating factor (DAF or CD55) is also able to recognize and dissociate both C3 and C5 convertases \cite{lukacik2004complement}.
+Carboxypeptidase-N, a well known inflammation regulator, cleaves carboxyl-terminal arginines and lysines of the complement proteins C3a, C4a, and C5a rendering them inactive \cite{liszewski1995control}.
+Lastly, the assembly of the MAC complex is inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface [REF].
+Thus, there are many points of control which influence complement activation across the three activation pathways.
+
+Developing quantitative mathematical models of complement could be crucial to understanding its role in the body.
+Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
+For example, Hirayama et al. modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
+while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways as a system of non-linear ODEs \cite{korotaevskiy2009non}.
+More recently, large mechanistic models of sections of complement have also been proposed.
+For example, Liu et al.
+analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
+Recently, Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated
+the complement response of the host and pathogen \cite{zewde2016quantitative}.
+However, these previous modeling studies involved little (if any) experimental validation.
+Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict experimentally validated complement dynamics.
+The central challenge is the estimation of model parameters from experimental data.
+Unlike other important cascades, such as coagulation for which there are well developed experimental tools and many publicly available data sets,
+the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of mechanistic complement models difficult.
+
+In this study, we analyzed an ensemble of reduced order complement models.
+Our reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
+We used this framework to capture the dynamics of C3a and C5a formation in the lectin and alternative pathways.
+The reduced order model consisted of only 18 differential equations with 28 parameters.
 Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
-We estimated the model parameters from in vitro time series data of C3a and C5a from Morad and coworkers \cite{morad2015time}.
-Subsequently we validated the model on unseen C3a and C5a experimental data that were not used for model training.
-After validation, we performed a sensitivity analysis on the model to estimate which parameters were critical to model performance under different experimental conditions.
-Given its small size, the hybrid approach produced a surprisingly predictive human complement model, similar to an earlier study on human coagulation using the same modeling framework \cite{sagar2015dynamic}.
-Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase elucidated new insight into the roles of parameters that govern the complement system.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a from Morad and coworkers \cite{morad2015time}.
+Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
+After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
+
+%Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
+%elucidated new insight into the roles of parameters that govern the complement system.
 A deeper understanding about how these parameters influence complement dynamics will greatly aid in the development of drugs for strategic therapeutic targets.
@@
 \ref{fig-prediction}).
-For the prediction of the $C3a$ at the three different inducer concentrations, our model was able to accurately capture the dynamic behaviors at a qualitative level, including the time the concentration of $C3a$ plateaus.
-For the prediction for $C5a$, our model was able to capture the behavior at a quantitative level for all three inducer concentrations.
+For the prediction of the C3a at the three different inducer concentrations, our model was able to accurately capture the dynamic behaviors at a qualitative level, including the time the concentration of C3a plateaus.
+For the prediction for C5a, our model was able to capture the behavior at a quantitative level for all three inducer concentrations.
 However, the our model was not able to capture the concave down curvature for the cases of $0.001~g$ and $0.01~g$ inducer.
-We believe some of the shortcomings of the prediction may be attributed by our reduced complexity in the modeling of $C4B$ interactions in our system.
-$C4BP$ interaction was modeled as irreversible mass action kinetics, in reality the binding of $C4BP$ onto other complement components are reversible.
-Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the $C4BP$ is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
+We believe some of the shortcomings of the prediction may be attributed by our reduced complexity in the modeling of C4B interactions in our system.
+C4BP interaction was modeled as irreversible mass action kinetics, in reality the binding of C4BP onto other complement components are reversible.
+Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the C4BP is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
 
@@
 The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin initiated pathway with 1g zymosan (\textbf{C}--\textbf{D}).
-The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of $C3a$ or $C5a$ for 50 independent particles.
-The shaded region denotes 99 \% confidence interval on the simulated mean concentration of $C3a$ or $C5a$ (uncertainty in the model simulation).
+The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
+The shaded region denotes 99 \% confidence interval on the simulated mean concentration of C3a or C5a (uncertainty in the model simulation).
 All initial concentrations of complement proteins are at human serum levels unless otherwise noted.}\label{fig-fit}
@@
 Simulations of C3a and C5a generated in the lectin pathway using different levels of zymosan ($0.1$, $0.01$, and $0.001$ grams of zymosan) were compared with the corresponding experimental data (\textbf{A}--\textbf{F}).
-The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of $C3a$ or $C5a$ for 50 independent particles.
-The shaded region denotes 99 \% confidence interval on the simulated mean concentration of $C3a$ or $C5a$ (uncertainty in the model simulation).
+The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
+The shaded region denotes 99 \% confidence interval on the simulated mean concentration of C3a or C5a (uncertainty in the model simulation).
 All initial concentrations of complement proteins are at human serum levels unless otherwise noted.

### 2016-05-26 Paper_v2.tex
Model ID paragraph

@@
 
-\subsection*{Formulation of a reduced order complement model}
-We developed a reduced order human complement network consisting of the most crucial steps of the human complement system (Fig.
-\ref{fig-schematic}).
-The core of our model was based upon the experimental measurements of Morad and coworker's earlier work \cite{morad2015time}, we only consider the activation of complement system through the alternate and the lectin pathways.
-In doing so we aim to capture a complex biological phenomenon using a few simple ordinary differential equations.
-A trigger event initiates the lectin pathway in the presence of zymosan, which activates the cleavage of C2 and C4 into C2a and C2b, and C4a and C4b respectively.
-Classical Pathway (CP) C3 convertase (C4aC2b) is a combination of C4a and C2b, which catalyzes the cleavage of C3 into C3a and C3b.
-Similarly, the activation of the alternative pathways happens through the spontaneous hydrolysis of C3 which facilitates the cleavage of C3.
-C3b then could combine with with C3 to form alternate pathway (AP) C3 convertase.
-Both C3 convertases catalyze the cleavage of C3 into C3a and C3b, and C3b can then combine with either CP or AP C3 convertase to form C5 convertase, CP or AP respectively that is responsible for the cleavage of C5 to C5a and C5b.
-Lectin pathway activation was approximated using a combination of saturation kinetics and Hill-like function control functions.
-These control coefficients then modified the rates of model processes at each time step.
-Hill-like transfer functions $0 \le \mathbf{f (Z)}  \le 1$ quantified the contribution of components upon a target process, in this study, $\mathbf{Z}$ represents the abundance of the initiator.
-Taken together, while the reduced order human complement model encodes significant biological complexity, it is highly compact (consisting of only 18 differential equations).
-Thus, it will serve as an excellent proof of principle example to study the reduction of a highly complex human subsystem.
+\subsection*{Formulation of the reduced order complement model.}
+The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
+A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
+Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
+On the other hand, activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
+The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
+Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
+A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
+The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
+Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions which modified the rates of model processes at each time step.
+These transfer functions allowed a significant reduction in the size of the model, while maintaining the biological performance.
+Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
+Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.
 
-\subsection*{An ensemble of complement models was estimated using dynamically dimensioned search.}
-A critical challenge for any dynamic model is the estimation of kinetic parameters. We estimated
-kinetic and control parameters in a hierarchical fashion using  two $in~vitro$ time-series human complement data sets with and without zymosan present.
-The residual between simulation and experimental measurements were minimized  using dynamically dimensioned search (DDS).
-An initial parameter set was initialized with randomized kinetic and control parameters and allowed to search for parameter vectors that minimized the residual.
-Knowing that the kinetic and control parameters of the lectin pathway does not affect the dynamics of the alternate pathway, we used a hierarchal approach that estimated the parameters for the alternative pathway and lectin pathway separately.
-For the alternative pathway, we utilized the time-course experimental measurements of Morad and coworkers \cite{morad2015time} of C3a and C5a in the absence of zymosan and only allowed the alternative parameters to vary (Fig.
+\subsection*{Estimating an ensemble of reduced order complement models.}
+A critical challenge for any dynamic model is the estimation of model parameters.
+We estimated the complement model parameters in a hierarchical fashion using two \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
+The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) approach,
+starting from an initial random parameter guess.
+Knowing the parameters of the lectin pathway did not affect the dynamics of the alternate pathway,
+we used a hierarchal approach that estimated the alternative and lectin pathway parameters separately.
+For the alternative pathway, we used the C3a and C5a measurements of Morad and coworkers \cite{morad2015time} in the absence of zymosan, and only allowed the alternative parameters to vary (Fig.
 \ref{fig-fit} A and B).
-The estimated alternate parameters was then fixed for the determination of lectin pathway parameters.
-The training for the lectin parameters, we used the experimental measurements of C3a and C5a in the presence of $1$ g of zymosan published by Morad et al \cite{morad2015time} (Fig.
-\ref{fig-fit} C and D).
-The reduced human complement model captured the behavior of the alternative and lectin pathways through the time-course abundance of C3a and C5a (Fig. \ref{fig-fit}).
-However we were not able to capture the curvature of the C5a alternative (Fig. \ref{fig-fit}).
-The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
-Taken together, the model identification results suggested that our reduced order approach could reproduce a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
+The putative alternate parameters were then fixed, and the lectin parameters were estimated.
+Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit} C and D).
+The reduced complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
+Taken together, the reduced order approach reproduced a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
 However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
+
+%However, we failed to fully capture the curvature of C5a alternative measureme (Fig. \ref{fig-fit}).
+%The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
 

### 2016-05-26 Paper_v2.tex
Going home ...

@@
 
-\subsection*{Formulation of the reduced order complement model.}
+\subsection*{Reduced order complement network.}
 The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
@@
 we used a hierarchal approach that estimated the alternative and lectin pathway parameters separately.
-For the alternative pathway, we used the C3a and C5a measurements of Morad and coworkers \cite{morad2015time} in the absence of zymosan, and only allowed the alternative parameters to vary (Fig.
-\ref{fig-fit} A and B).
+The reduced order complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
+For the alternative pathway, we used the C3a and C5a measurements in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit}A and B).
 The putative alternate parameters were then fixed, and the lectin parameters were estimated.
-Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit} C and D).
-The reduced complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
-Taken together, the reduced order approach reproduced a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
+Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit}C and D).
+Taken together, the reduced order model reproduced a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
 However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
+To address this question, we fixed the model parameters and simulated data not used for model training.
 
@@
 
-We tested the predictive power of the reduced order human complement model with validation data sets not used during model training.
+We tested the predictive power of the reduced order complement model with data sets not used during model training (Fig. \ref{fig-prediction}).
 Six validation data sets were used, three for C3a and C5a respectively at different zymosan concentrations.
-All kinetic and control parameters were
-fixed for the validation simulations.
-The reduced order model predicted the C3a and C5a time-course profiles at a qualitative level (Fig.
-\ref{fig-prediction}).
-For the prediction of the C3a at the three different inducer concentrations, our model was able to accurately capture the dynamic behaviors at a qualitative level, including the time the concentration of C3a plateaus.
-For the prediction for C5a, our model was able to capture the behavior at a quantitative level for all three inducer concentrations.
-However, the our model was not able to capture the concave down curvature for the cases of $0.001~g$ and $0.01~g$ inducer.
-We believe some of the shortcomings of the prediction may be attributed by our reduced complexity in the modeling of C4B interactions in our system.
-C4BP interaction was modeled as irreversible mass action kinetics, in reality the binding of C4BP onto other complement components are reversible.
+All model parameters were fixed for the validation simulations.
+The ensemble of reduced order models qualitatively predicted C3a and C5a dynamics as a function of inducer abundance.
+For the prediction of the C3a at the three different inducer concentrations, the model captured dynamics C3a formation.
+The model also captured the behavior of C5a formation at all three inducer concentrations.
+However, the model did not capture the concave down curvature for 0.001~g and 0.01~g zymosan.
+We believe some of the shortcomings of the prediction may be attributed to our reduced complexity in the modeling of C4B interactions in our system.
+C4BP interaction were modeled as irreversible mass action kinetics; however, the binding of C4BP with other complement components is likely reversible.
 Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the C4BP is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
 
-\subsection*{Global Sensitivity analysis of the reduced order complement model}
+\subsection*{Global sensitivity analysis of the reduced order complement model}
 We conducted a Sobol's sensitivity analysis to estimate which parameters controlled the performance of the reduced order model.

### 2016-05-26 Paper_v2.tex
Sensitvity section

@@
 %The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
+%Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the C4BP is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
 
-We tested the predictive power of the reduced order complement model with data sets not used during model training (Fig. \ref{fig-prediction}).
-Six validation data sets were used, three for C3a and C5a respectively at different zymosan concentrations.
+We tested the predictive power of the reduced order complement model with data not used during model training (Fig. \ref{fig-prediction}).
+Six validation cases were considered, three for C3a and C5a respectively at different zymosan concentrations.
 All model parameters were fixed for the validation simulations.
-The ensemble of reduced order models qualitatively predicted C3a and C5a dynamics as a function of inducer abundance.
-For the prediction of the C3a at the three different inducer concentrations, the model captured dynamics C3a formation.
-The model also captured the behavior of C5a formation at all three inducer concentrations.
-However, the model did not capture the concave down curvature for 0.001~g and 0.01~g zymosan.
-We believe some of the shortcomings of the prediction may be attributed to our reduced complexity in the modeling of C4B interactions in our system.
-C4BP interaction were modeled as irreversible mass action kinetics; however, the binding of C4BP with other complement components is likely reversible.
-Due to this discrepancy in between our model and the experimental results, we believe that the interaction and the reversibility of the C4BP is a key player in creating a threshold where C5a is resistant to minor changes in inducer concentration even though C3a is sensitive.
+The ensemble of reduced order models captured the qualitative dynamics of C3a formation (Fig.
+\ref{fig-prediction}, left column), and C5a formation (Fig.
+\ref{fig-prediction}, right column) at three inducer concentrations.
+However, there were shortcomings, especially for the C3a prediction.
+First, while the C3a dynamics were captured, and particularly the peak time, the overall level of C3a was under-predicted in all cases (Fig. \ref{fig-prediction}, inset left column).
+We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
+C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes;
+however, the binding of C4BP with complement proteins is likely reversible and convertases may have residual activity even in the bound form.
+Thus, we likely over-predicted the influence of C4BP.
+We also failed to capture the concave down curvature for the 0.001~g and 0.01~g zymosan cases in the C5a validation studies.
+The decreasing slope of the C5a measurements may indicate decreasing cofactors abundance, or missing biology which we have not explicitly accounted for in the reduced order approach.
+However, despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the dynamic time scale of C3a formation,
+and the correct order of magnitude for the concentration of C5a for three inducer levels.
+Next, we used global sensitivity analysis to determine which parameters controlled the performance of the complement model.
+
+%This gives us new insight in which of the parameters play a role in complement activation.
+Even though AP~C3~convertase is also responsible in the conversion of C3 and the production of C3a, the kinetic parameters that govern the equation was not sensitive at all.
+This elucidated that the activation of
+
+%
 
 \subsection*{Global sensitivity analysis of the reduced order complement model}
-We conducted a Sobol's sensitivity analysis to estimate which parameters controlled the performance of the reduced order model.
-We calculated the sensitivity of the change in C3a and C5a profiles using the residuals between simulation and experimentally measured data for the cases of 0 and 1g zymosan (Fig.
-\ref{fig-SA}).
-For the cases in absence of zymosan where only the alternative pathway is active, we observed that only a few variables are responsible for the system response.
-For C3a alternate, the sensitivity analysis found that $k_{c3b~basal}$ and $k_{degradationC3a}$ are the only sensitive parameters.
-This gives us new insight in which of the parameters play a role in complement activation.
-Even though AP~C3~convertase is also responsible in the conversion of C3 and the production of C3a, the kinetic parameters that govern the equation was not sensitive at all.
-This elucidated that the activation of alternative pathway is more heavily governed by the spontaneous hydrolysis of C3 rather than the activity of AP~C3~Convertase.
-Surprisingly, closely examining the sensitive parameters that control C5a, in addition to the expected kinetic and control parameters related to the formation of AP~C5~Convertase, we observed that $k_C3~Convertase2$, the was previously not sensitive to C3a, to be sensitive in the formation of C5a.
-The AP~C3~Convertase is a substrate required for the formation of AP~C5~Convertase and the formation of C3b.
-The change in activity of AP C3 Convertase  will not drastically change the C3a dynamics, but will effect AP~C5a~Convertase formation and C5a formation.
-The our reduced order human complement model in combination with Sobol's sensitivity analysis was able to unravel important indirect parameter interaction.
-Our sensitivity analysis yielded expected results for the lectin pathway analyzes (Fig.
-\ref{fig-SA} (C and D)).
-One key difference that was observed between the sensitivity of the parameters between C3a an C5a was their respective degradation terms.
-The degradation constant of C3a was sensitive between the two different cases of zymosan that was tested while the degradation constant of the C5a was not sensitive.
-We believe this different is attributed to the magnitude of the parameters and their respective concentrations.
+We conducted global sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
+We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}).
+In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (formation of C3b) and $k_{d,C3a}$ (degradation rate constant governing C3a)
+were largely responsible for the system response. Interestingly, $k_{c,C3}$ (the rate constant governing AP C3-convertase activity) was not sensitive in the absence of zymosan.
+Thus, the behavior of the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity.
+On the other hand, $k_{c,C3}$ was one of the parameters that controlled C5a formation, in addition to the expected parameters related to AP C5-convertase formation.
+The AP C3-convertase is required for AP C5-convertase formation, and the formation of the C3b fragment.
+Thus, changes in the activity of AP C3-convertase will not drastically change the C3a dynamics, but will effect AP C5-convertase activity and C5a formation.
+The sensitivity analysis yielded the expected results for the lectin pathway (Fig. \ref{fig-SA}C and D).
+One key difference observed between the sensitivity of C3a an C5a parameters, was their respective degradation constants.
+The rate constant governing C3a degradation was sensitive, while the degradation constant for C5a was not.
+We believe this difference was attributable to the magnitude of the parameters and the respective concentration of C3a and C5a.
+Taken together, the reduced order complement model in combination with global sensitivity analysis unravelled important indirect parameter interactions.
 

### 2016-05-27 Paper_v2.tex
Working on the discussion

@@
 The rate constant governing C3a degradation was sensitive, while the degradation constant for C5a was not.
-We believe this difference was attributable to the magnitude of the parameters and the respective concentration of C3a and C5a.
-Taken together, the reduced order complement model in combination with global sensitivity analysis unravelled important indirect parameter interactions.
+This difference was likely attributable to the magnitude of the degradation parameters and the respective concentration of C3a and C5a.
+Taken together, global sensitivity analysis of the reduced order complement model identified important indirect parameter interactions that could have therapeutic significance.
 
@@
 %As the dimensionality of  increases, the search region gets widened and thus the problem becomes more challenging.
-%
+%The influence of CRP, LF and the cross-talk can be captured through additional control functions that act upon the initiation pathways in a logical integration rule developed by Wayman and coworkers \cite{wayman2015dynamic}.
 
@@
 
-The discussion has three (sometimes four) paragraphs:
-\begin{enumerate}
-	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
+In this study, we analyzed an ensemble of reduced order complement models.
+The reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
+The reduced order model consisted of only 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
+We used this framework to simulate the dynamics of C3a and C5a formation in the lectin and alternative pathways.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a abundance.
+Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
+After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
+The global sensitivity analysis of the reduced order complement model unravelled important indirect parameter interactions that could have therapeutic significance.
 
-	In this study, we present a hybrid modeling approach to build a reduced order model of complement.
+The performance of the reduced order complement model was impressive given its limited size.
+However, there are several questions that should be explored further.
+A logical progression for this work would be to expand the network to include the classical pathway and the formation of the membrane attack complex (MAC).
+However, it is unclear whether the addition of the classical pathway will decrease the predictive quality of our existing model.
+Liu et al have shown cross-talk between the activation of the classical and lectin pathways that could influence model performance \cite{liu2011computational}.
+One potential approach to address such difficulties would be to incorporate C reactive proteins (CRP) and L-ficolin (LF) into the model,
+both of which are involved with the initiation of classical and lectin pathways.
+Time course measurements of MAC abundance (and MAC formation dynamics) are also scarce, making the inclusion of MAC challenging.
+Next, we should address the under-prediction of C3a. We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
+C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes;
+however, the binding of C4BP with complement proteins is likely reversible and C4BP-bound convertases may have residual activity.
+We should test alternative model structures which include reversible C4BP binding, and partially active convertases.
+Alternatively, we could also perform sensitivity analysis on the C3a prediction residual to determine which parameters controlled the C3a prediction.
+
+%that are implicitly lumped together with our effective kinetics and control parameters.
+Due to the reduction of parameters, the model cannot determine the dynamics or explicit impact of the omitted species on the system.
+However, we have created a hierarchy approach for parameter estimation that can be used to uncouple the kinetic parameter and contribution of any additional complement proteins and regulators.
+Using this simple and versatile modeling approach that we created, we took the first step in the development of a computation toolkit that can be readily used in a clinical setting.
+Our reduced order complement model is computationally inexpensive, and versatile so it could easily be incorporated into pre-existing or new pharmacokinetic models.
+Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
+
+% In this study, we present a hybrid modeling approach to build a reduced order model of complement.
 The key innovation of this approach is the use of simple equations to capture the behavior of a complex biochemical network.
@@
 The reduced order model consisted of only 18 differential equations with 28 kinetic and control parameters.
-Thus, the model was an order  of magnitude smaller and included more pathways than comparable ODE models in the literature.
 We estimated the model parameters from in vitro time series data of C3a and C5a from Morad and coworkers \cite{morad2015time}.
@@
 Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.
+
+The discussion has three (sometimes four) paragraphs:
+\begin{enumerate}
+	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
 
@@
 In the diseases driven by dysfunctional alternative pathway, the activation of C5 appears to be the source that fueled the cycle by causing endothelial damage.
-
-The performance of the reduced order complement model was impressive given its limited size.
-However, there are several critical questions that should be explored following this study.
-A logical progression for this work would include expanding the network to include the classical pathway and the formation of the membrane attack complex (MAC).
-It is unclear whether the addition of the classical pathway will decrease the prediction of our existing model due to the cross-talk between the classical and lectin activation shown by Liu et al \cite{liu2011computational}.
-One potential approach in addressing such difficulties would be the incorporation of additional species such as C reactive proteins (CRP) and L-ficolin (LF) that involved in complement initiation of classical and lectin pathways.
-The influence of CRP, LF and the cross-talk can be captured through additional control functions that act upon the initiation pathways in a logical integration rule developed by Wayman and coworkers \cite{wayman2015dynamic}.
-Another issue with our reduced order model involve the omitted species that are implicitly lumped together with our effective kinetics and control parameters.
-Due to the reduction of parameters, the model cannot determine the dynamics or explicit impact of the omitted species on the system.
-However, we have created a hierarchy approach for parameter estimation that can be used to uncouple the kinetic parameter and contribution of any additional complement proteins and regulators.
-Using this simple and versatile modeling approach that we created, we took the first step in the development of a computation toolkit that can be readily used in a clinical setting.
-Our reduced order complement model is computationally inexpensive, and versatile so it could easily be incorporated into pre-existing or new pharmacokinetic models.
-Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
 

### 2016-05-27 Paper_v2.tex
One more discussion paragraph to go

@@
 This difference was likely attributable to the magnitude of the degradation parameters and the respective concentration of C3a and C5a.
-Taken together, global sensitivity analysis of the reduced order complement model identified important indirect parameter interactions that could have therapeutic significance.
+Taken together, global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
 
@@
 Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
-Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
+Given its small size, the hybrid approach produced a surprisingly predictive complement model.
 After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
-The global sensitivity analysis of the reduced order complement model unravelled important indirect parameter interactions that could have therapeutic significance.
+The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+Using a simple and versatile modeling approach, we developed a reduced order complement model is computationally inexpensive, and versatile so it could easily be incorporated into pre-existing or new pharmacokinetic models.
+Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
+
+Despite its importance, there has been a paucity of validated mathematical models of complement pathway activation.
+To our knowledge, this study is one of the first complement models that combined multiple initiation pathways with experimental validation of important complement products like C5a.
+However, there have been several theoretical models of components of the cascade in the literature.
+Liu and co-workers modeled the formation of C3a through the classical pathway using 45 non-linear ODEs  \cite{liu2011computational}.
+In contrast, in this study we modeled lectin mediated C3a formation using only five ODEs.
+Though we did not model all the initiation interactions in detail, especially the cross-talk between the lectin and classical pathways,
+we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
+The model also captured the dynamics of C3a and C5a formed from the alternate pathway using only seven ODEs.
+The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al which involved over 100 ODEs \cite{zewde2016quantitative}.
+However, we found that the quantity of C3a produced in the alternate pathway was nearly 1000 times the quantity of C5a produced.
+Though this was in agreement with the experimental data \cite{morad2015time},
+it differed from the theoretical predictions made by Zewde et al. who showed C3a was ${10^8}$ times the C5a concentration \cite{zewde2016quantitative}.
+In our model, the time profile of C5a generation from the lectin pathway changed with respect to the quantity of zymosan (the lectin pathway initiator).
+The lag phase for generation was inversely proportional to the initiator concentration.
+Korotaevskiy et al. showed a similar trend using a theoretical model of complement, albeit for much shorter time scales \cite{korotaevskiy2009non}.
+Taken together, the reduced order complement model performed similarly to existing large mechanistic models, despite being significantly smaller.
+
+% complement in immune response has been well known since long, there
+% The lag phase in the initiation of C5a is followed by an acceleration in the production of C5a.
+%  Similarly, our model was able to capture C3a formation from the alternate pathway that showed the same qualitative trends as Zewde et al \cite{zewde2016quantitative}.
+% It is possible that the experimental system has trace amounts of C5a which leads to this difference.
+% Zewde et al. \cite{zewde2016quantitative} show a similar time profile for C5a generated in the alternate pathway as well.
 
@@
 however, the binding of C4BP with complement proteins is likely reversible and C4BP-bound convertases may have residual activity.
+We also did not capture the maximum concentration of C3a at low initiator levels.
+One possible reasons for this could the C2-by pass pathway, which was not included in the model.
+This pathway further accelerates C3a production without the involvement of a C3 convertase.
+Currently the C3a in the model is generated only through the activity of a C3 convertase.
+Incorporating this additional step within the reduced order modeling framework would be a future direction that we need to consider.
 We should test alternative model structures which include reversible C4BP binding, and partially active convertases.
@@
 	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
-
-Though the role of complement in immune response has been well known since long, there has been a paucity of mathematical models of complement.
-To our knowledge this is the first model of complement that combines different pathways of initiation and validates the dynamics of downstream proteins like C5a using experimental data.
-Liu and co-workers modeled formation of C3a through the classical pathway using 45 non-linear ODEs  \cite{liu2011computational}.
-The hybrid modeling framework, however, allowed us to model the lectin mediated formation of C3a using only 5 ODEs.
-Though we do not model all the interactions of initiation in detail, especially the cross-talk between lectin and classical pathways like Liu et al.
-\cite{liu2011computational} we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
-The model was also surprisingly accurate in capturing the quantitative dynamics of C3a and C5a formed from the alternate pathway with only 7 ODEs.
-The lag phase in the initiation of C5a is followed by an acceleration in the production of C5a.
-This is also qualitatively similar to the predicted C5a time profiles using a theoretical model of complement with 107 equations \cite{zewde2016quantitative}.
-Similarly, our model was able to capture C3a formation from the alternate pathway that showed the same qualitative trends as Zewde et al \cite{zewde2016quantitative}.
-We also observe in our model simulation that the quantity of C3a produced in the alternate pathway is nearly 1000 times the quantity of  C5a produced.
-Though this is in agreement with the experimental data  \cite{morad2015time}, it differs from the theoretical predictions by Zewde et al.\cite{zewde2016quantitative} who show that C3a is ${10^8}$ times the C5a concentration.
-It is possible that the experimental system has trace amounts of C5a which leads to this difference.
-In our model the time profile of C5a generation from the lectin pathway changes with respect to the quantity of zymosan (the lectin pathway initiator).
-We see that the lag phase for generation progressively decreases with increasing concentration of the initiator.
-Korotaevskiy et al.
-\cite{korotaevskiy2009non} show a similar lag phase followed by an accelerated production of C5a using a theoretical model of complement for much smaller time scales.
-Zewde et al.
-\cite{zewde2016quantitative} show a similar time profile for C5a generated in the alternate pathway as well.
-We do observe a similar trend in our model.
-However given the different time scale in our study we do not observe a very prominent lag phase.
-Our model also captures the maximum concentrations of C5a with different initiators and in the alternate pathway.
-However we do not capture the maximum concentration of C3a with low initiator levels though we capture the overall trend which includes the initiation and amplification phases.
-One of the possible reasons for this could be the non-inclusion of C2-by pass pathway which further accelerates production of C3a without the involvement of a C3 convertase.
-Currently the C3a in the model is generated only through the activity of a C3 convertase.
-Incorporating this additional step within the reduced order modeling framework would be a future direction that we need to consider.
-Taken together we surprisingly do very well in capturing the dynamics of key complement proteins and observe similar trends as in large mechanistic models.
 

### 2016-05-27 Paper_v2.tex
Draft discussion

@@
 Korotaevskiy et al. showed a similar trend using a theoretical model of complement, albeit for much shorter time scales \cite{korotaevskiy2009non}.
-Taken together, the reduced order complement model performed similarly to existing large mechanistic models, despite being significantly smaller.
+Thus, the reduced order complement model performed similarly to existing large mechanistic models, despite being significantly smaller.
+
+Global sensitivity analysis of the complement model estimated potential important therapeutic targets.
+Complement malfunctions are implicated in a number of diseases, however the
+development of complement specific therapeutics has been challenging \cite{ricklin2007complement}.
+Previously, we have shown that mathematical modeling and sensitivity analysis can be useful tools to estimate therapeutically important mechanisms in biochemical networks \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
+In this study, we analyzed a validated ensemble of reduced order complement models to estimate therapeutically important mechanisms.
+In presence of an initiator, C5a formation was primarily sensitive to the lectin initiation parameters,
+and parameters governing the conversion of C5 to C5a and C5b.
+This result agrees well with the current protease inhibitors targeting initiating complexes, including mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{morgan2015complement}.
+The most commonly used anti-complement drug eculizumab \cite{morgan2015complement}, targets the C5 protein which is cleaved to form C5a.
+Our sensitivity analysis showed that kinetic parameters governing C5 conversion were sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
+The formation of basal C3b was also a sensitive parameter in the formation of C3a through the alternate pathway.
+Thus, this mechanism can act as a target for both C3a and C5a inhibitors.
+Lectin initiated C3a formation showed a number of sensitive parameters.
+This included the lectin initiation parameters that controlled C5a formation, C3 convertase inhibition by C4BP, and parameters governing C3 convertase activity.
+All these mechanisms are potential drug targets.
+For example, both C4BP and C3 convertase inhibitors are currently being developed as next generation complement inhibitors [REF ORIGINAL STUDIES NOT REVIEW].
+Thus, analysis of the ensemble of complement models identified potentially important therapeutic targets.
+
+%Without initiator, C5a formation was the most sensitive to the formation of basal C3b through the tick over mechanism.
+%On the other hand,
+
+%. Currently there are only four approved drugs targeting complement \cite{morgan2015complement}. Eculizumab targets C5, Berinert, Ruconest and Cinryze target C1.
 
@@
 One potential approach to address such difficulties would be to incorporate C reactive proteins (CRP) and L-ficolin (LF) into the model,
-both of which are involved with the initiation of classical and lectin pathways.
+both of which are involved with the initiation of classical and lectin pathways [HOW DOES THIS HELP THE CROSSTALK ISSUE?].
 Time course measurements of MAC abundance (and MAC formation dynamics) are also scarce, making the inclusion of MAC challenging.
@@
 
-The discussion has three (sometimes four) paragraphs:
-\begin{enumerate}
-	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
-
-	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
-
-%%% We talk about complement therapeutics here in relation with sensitivity analysis
-
-The global sensitivity analysis of our reduced order model provides interesting insights into the regulation of C3a and C5a which are important therapeutic targets.
-We see that the formation of C5a in the absence of an initiator is the most sensitive to the formation of basal C3b through the tick over mechanism.
-This mechanism could be a potential target for C5a inhibitors especially in cases of autoimmune disorders where there is an absence of initiators of the lectin or the classical pathway.
-In presence of an initiator we see that C5a formation is primarily sensitive to the lectin initiation parameters and the kinetic parameters that govern the conversion of C5 to C5a and C5b.
-This result agrees well with the current protease inhibitors targeting initiating complexes that includes mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{morgan2015complement}.
-The most commonly used anti-complement drug eculizumab \cite{morgan2015complement}, targets C5 protein which is cleaved to form C5a.
-Our sensitivity analysis shows that kinetic parameters governing C5 conversion are sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
-The formation of basal C3b is also a sensitive parameter in the formation of C3a through the alternate pathway.
-Thus this mechanism can act as a target for both C3a and C5a inhibitors.
-The lectin initiated formation of C3a shows a number of sensitive parameters.
-This includes the lectin initiation parameters that control the formation of C5a along with C3 convertase inhibition using C4BP and the parameter governing C3 convertase.
-All these mechanisms involve potential drug targets and literature shows that C4BP, C3 convertase inhibitors are the being developed as the next generation inhibitors of complement \cite{morgan2015complement}.
-Taken together it is surprising to see that reduced order model has the ability to identify mechanisms for potential drug targets that are in agreement with existing literature.
-%Though it is known complement malfunction causes a number of diseases \cite{ricklin2007complement} developing complement specific therapeutics has been challenging\cite{ricklin2007complement}.
-Currently there are only four approved drugs targeting complement \cite{morgan2015complement}.
-Eculizumab targets C5, Berinert, Ruconest and Cinryze target C1.
-
-	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
-	In this case, we will have a scaling issue if we extend to genome scale. We should extend to dynamic cases, and we need to experimentally validate the findings. }
-\end{enumerate}
-%This suggests that our reduced order model's prediction for the alternate pathway have great potential application for the study of glomerular diseases that show strong correlation with disturbed alternative pathway activity (Pickering2011).
+% The discussion has three (sometimes four) paragraphs:
+% \begin{enumerate}
+% 	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
+%
+%
+%
+% 	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
+%
+%
+% %%% We talk about complement therapeutics here in relation with sensitivity analysis
+%
+%
+% 	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
+% 	In this case, we will have a scaling issue if we extend to genome scale. We should extend to dynamic cases, and we need to experimentally validate the findings. }
+% \end{enumerate}
+% %This suggests that our reduced order model's prediction for the alternate pathway have great potential application for the study of glomerular diseases that show strong correlation with disturbed alternative pathway activity (Pickering2011).
 In the diseases driven by dysfunctional alternative pathway, the activation of C5 appears to be the source that fueled the cycle by causing endothelial damage.

### 2016-05-27 Paper_v2.tex
Updated the equations

@@
 
+%The reaction rates controlling  such that $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
+
 \section*{Materials and Methods}
-We used ordinary differential equations (ODEs) to model the time evolution of proteins ($x_{i}$) in our reduced order complement model:
+We used ordinary differential equations (ODEs) to model the time evolution of complement proteins ($x_{i}$) in the reduced order model:
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{R}$ denotes the number of reactions, $\mathcal{M}$ denotes the number of protein species in the model.
+where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of protein species in the model.
 The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
-Typically, reaction $j$ is a non-linear function of biochemical species abundance, as well as unknown kinetic parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
+Typically, reaction $j$ is a non-linear function of biochemical and enzyme species abundance, as well as unknown model parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
 The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$.
@@
 
-The reaction rates controlling formation C4a, C4b, C2a and C2b were written as a product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$) such that $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
-The kinetic term for these rates was modeled using saturation kinetics.
-The control term $0\leq v_{j}\leq 1$ for these reaction rates was modeled using regulatory transfer functions which took the form:
-\begin{equation}\label{eqn:control-factor}
-	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right) = k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
-\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
-
-We used saturation kinetics to model the lectin pathway activation and C3 and C5 convertase activity $\bar{r}_{j}$:
+Rate processes were written as the product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$) in the complement model.
+The kinetic term for the formation of C4a, C4b, C2a and C2b, lectin pathway activation, and C3 and C5 convertase activity was given by:
 \begin{equation}\label{eqn:rate-saturation}
@@
 \end{equation}
-where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$, $\eta$ denotes a cooperativity parameter (similar to a Hill coefficient), and $K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
-On the other hand,
-we used mass action kinetics to model the protein conversion reactions within the network $\bar{r}_{j}$:
+where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the abundance of the enzyme catalyzing reaction $j$,
+$\eta$ denotes a cooperativity parameter, and $K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
+We used mass action kinetics to model protein-protein binding interactions within the network:
 \begin{equation}\label{eqn:rate-action}
-	\bar{r}_{j} = k_{j}^{max}\prod_{s\in{m_{j}^{-}}}x_{s}
+	\bar{r}_{j} = k_{j}^{max}\prod_{s\in{m_{j}^{-}}}x_{s}^{-\sigma_{sj}}
 \end{equation}
-where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the enzyme abundance which catalyzes reaction $j$.
-The product in Eqn \eqref{eqn:rate-action} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).
+where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\sigma_{sj}$ denotes the stoichiometric coefficient for species $s$ in reaction $j$,
+and $s\in{m_{j}}$ denotes the set of \textit{reactants} for reaction $j$.
+The control terms $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
+For each rate, we used a rule-based approach to select from competing control factors.
+If rate j was influenced by $1,\dots,m$ factors, we modeled this relationship as
+$v_{j}=\mathcal{I}_{j}\left(f_{1j}\left(\cdot\right),\hdots,f_{mj}\left(\cdot\right)\right)$
+where $0\leq f_{ij}\left(\cdot\right)\leq 1$ denotes a regulatory transfer function quantifying the influence of factor $i$ on rate $j$.
+The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
+variable. Each regulatory transfer function took the form:
+\begin{equation}\label{eqn:control-factor}
+	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right)=k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
+\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance of factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
+In this study, we used $\mathcal{I}_{j}\in\left\{mean\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
+The model equations implemented in MATLAB and solved using the \texttt{ODE15s} routine (The Mathworks, Natick MA).
+The complement model code and parameter ensemble can be downloaded from http://www.varnerlab.org.
 
@@
 \end{equation}
-
 where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.

### 2016-05-27 Paper_v2.tex
Materials and methods done

@@
 \end{equation}where $\mathcal{Z}_{i}$ denotes the abundance of factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
-In this study, we used $\mathcal{I}_{j}\in\left\{mean\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
-The model equations implemented in MATLAB and solved using the \texttt{ODE15s} routine (The Mathworks, Natick MA).
+In this study, we used $\mathcal{I}_{j}\in\left\{min,max\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
+The model equations were implemented in MATLAB and solved using the \texttt{ODE15s} routine (The Mathworks, Natick MA).
 The complement model code and parameter ensemble can be downloaded from http://www.varnerlab.org.
@@
 \subsection*{Estimation of an ensemble of model parameters.}
-Model parameters were estimated by minimizing the difference between simulations and experimental C3a and C5a measurements (squared residual):
+We minimized the residual between simulations and experimental C3a and C5a measurements using Dynamic Optimization with Particle Swarms (DOPS). DOPS minimized the objective:
 \begin{equation}\label{eqn:objective-function}
@@
 time, while the inner summation is with respect to state.
-
-We minimized the model residual using Dynamic Optimization with Particle Swarms (DOPS).
 DOPS is a novel metaheuristic that combines multi swarm particle swarm optimization (PSO) with a greedy global optimization algorithm called dynamically dimensioned search (DDS).
-DOPS is much faster than conventional global optimizers and has the ability to find near optimal solutions for high dimensional systems within a relatively few function evaluations.
-It uses an adaptive switching strategy based on error convergence rates to switch from swarms search to DDS search.
-This enables it to find quickly, globally optimal or close to globally optimal solutions even in the presence of many local minima.
+DOPS is faster than conventional global optimizers and has the ability to find near optimal solutions for high dimensional systems within a relatively few function evaluations.
+It uses an adaptive switching strategy based on error convergence rates to switch from the particle swarm to DDS search phases.
+This enables DOPS to quickly estimate globally optimal or near optimal solutions even in the presence of many local minima.
 In the swarm search, for each iteration the particles compute error within each sub-swarm by evaluating the model equations using their specific parameter vector realization.
@@
 The quality of parameter estimates was measured using goodness of fit (model residual).
-The DOPS routine was implemented in the MATLAB programming language.
+The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org.
 
@@
 
-\subsection*{Global sensitivity analysis of model performance}
-We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \cite{SOBOL_METHOD}.
-We computed the total sensitivity index of each parameter relative to four performance objectives, each objective was based on the sum of squared errors between model and experimental data for C3a alternate, C5a alternate, C3a lectin, and C5a lectin simulations.
-We established the sampling bounds for each parameter from the minimum and maximum value of that parameter in the parameter set ensemble.
-We used the sampling method of Saltelli \textit{et al.} \cite{saltelli2010variance} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
-where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from
-11,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
+\subsection*{Global sensitivity analysis of model performance.}
+We conducted a global sensitivity analysis to estimate which parameters controlled the performance of the reduced order model using the Sobol method \cite{SOBOL_METHOD}.
+We computed the total sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases.
+We established the sampling bounds for each parameter from the minimum and maximum value for that parameter in the parameter ensemble.
+We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
+where $N$ was the number of trials, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
+In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from 11,600 model evaluations.
+The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
 

### 2016-06-02 Paper_v2.tex
Changes to discussion figures and references

@@
 Complement is a central part of innate immunity which plays a significant role in the inflammatory response.
-Complement was discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies [REF].
+Complement was discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{sarma2011complement}.
 However, research over the past decade has shown the importance of complement extends well beyond innate immunity.
@@
 Complement can also play both a positive and negative role in certain cancers; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
-Several other important biochemical networks are integrated with complement including the coagulation cascade, the autonomous nervous response and the ability to regulate inflammation \cite{ricklin2013complement}.
+Several other important biochemical networks are integrated with complement including the coagulation cascade, the autonomous nervous system and the ability to regulate inflammation \cite{ricklin2013complement}.
 Thus, complement is an important system involved in a variety of both beneficial and potentially harmful functions in the body.
@@
 The central output of complement activation is the formation of the Membrane Attack Complex (MAC) and a key protein called C5a.
-The membrane attack complex forms transmembrane channels which disrupt the cell membrane of targeted cells, leading to cell lysis and death [REF].
-On the other hand, the C5a protein acts as a bridge between innate and adaptive immunity, and plays an important role in regulating inflammation and coagulation \cite{sarma2011complement}.
+The membrane attack complex forms transmembrane channels which disrupt the cell membrane of targeted cells, leading to cell lysis and death.
+The C5a protein acts as a bridge between innate and adaptive immunity, and plays an important role in regulating inflammation and coagulation \cite{sarma2011complement}.
 Complement activation takes places through three pathways: the alternate, the classical and the lectin binding pathway.
@@
 The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is hydrolyzed to form C3b.
-In presence of pathogens, the C3b fragment binds foreign surfaces and recruits the additional proteins, factor B and factor D, which lead to the formation of C3bBb, the alternate C3 convertase \cite{pangburn1984alternative}.
+In the presence of pathogens, the C3b fragment binds foreign surfaces and recruits the additional proteins, factor B and factor D, which lead to the formation of C3bBb, the alternate C3 convertase \cite{pangburn1984alternative}.
 The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
@@
 through the interaction of factor H with C3b \cite{blom2001structural}.
-These regulatory proteins are also capable of binding their respective targets in the convertase form as well.
+These regulatory proteins are also capable of binding their respective targets while they are bound in convertase complexes.
 Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
@@
 Carboxypeptidase-N, a well known inflammation regulator, cleaves carboxyl-terminal arginines and lysines of the complement proteins C3a, C4a, and C5a rendering them inactive \cite{liszewski1995control}.
-Lastly, the assembly of the MAC complex is inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface [REF].
+Lastly, the assembly of the MAC complex is inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
 Thus, there are many points of control which influence complement activation across the three activation pathways.
@@
 the complement response of the host and pathogen \cite{zewde2016quantitative}.
-However, these previous modeling studies involved little (if any) experimental validation.
+However, these previous modeling studies involved little experimental validation.
 Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict experimentally validated complement dynamics.
@@
 We used this framework to capture the dynamics of C3a and C5a formation in the lectin and alternative pathways.
-The reduced order model consisted of only 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
+The reduced order model consisted of 18 differential equations with 28 parameters.
+The model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
 We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a from Morad and coworkers \cite{morad2015time}.
-Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Subsequently, we validated the model by predicting C3a and C5a measurements that were not used for model training.
 Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
-After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
-
+After validation, we performed a global analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions and to identify potential therapeutic targets.
 %Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
@@
 
+%Paragraph Edited by Mason
 \subsection*{Reduced order complement network.}
@@
 Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
-On the other hand, activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
+Activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
 The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
@@
 The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
-Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions which modified the rates of model processes at each time step.
-These transfer functions allowed a significant reduction in the size of the model, while maintaining the biological performance.
+Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
 Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
@@
 starting from an initial random parameter guess.
-Knowing the parameters of the lectin pathway did not affect the dynamics of the alternate pathway,
-we used a hierarchal approach that estimated the alternative and lectin pathway parameters separately.
+A hierarchical approach was taken to determine model parameters in which the alternate pathway parameters were first estimated and then fixed during the estimation of the lectin pathway parameters.
 The reduced order complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
 For the alternative pathway, we used the C3a and C5a measurements in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit}A and B).
-The putative alternate parameters were then fixed, and the lectin parameters were estimated.
 Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit}C and D).
@@
 However, there were shortcomings, especially for the C3a prediction.
-First, while the C3a dynamics were captured, and particularly the peak time, the overall level of C3a was under-predicted in all cases (Fig. \ref{fig-prediction}, inset left column).
+First, while the C3a dynamics and concentration peak times were captured, the overall level of C3a was under-predicted in all cases (Fig. \ref{fig-prediction}, inset left column).
 We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
@@
 however, the binding of C4BP with complement proteins is likely reversible and convertases may have residual activity even in the bound form.
-Thus, we likely over-predicted the influence of C4BP.
+Thus, the model may over-predict the influence of C4BP.
 We also failed to capture the concave down curvature for the 0.001~g and 0.01~g zymosan cases in the C5a validation studies.
@@
 %
-
-\subsection*{Global sensitivity analysis of the reduced order complement model}
-We conducted global sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
-We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}).
+\subsection*{Global analysis of the reduced order complement model}
+We conducted sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
+We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA} (A-D)).
 In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (formation of C3b) and $k_{d,C3a}$ (degradation rate constant governing C3a)
@@
 Thus, changes in the activity of AP C3-convertase will not drastically change the C3a dynamics, but will effect AP C5-convertase activity and C5a formation.
-The sensitivity analysis yielded the expected results for the lectin pathway (Fig. \ref{fig-SA}C and D).
-One key difference observed between the sensitivity of C3a an C5a parameters, was their respective degradation constants.
+The sensitivity analysis yielded the expected results for the lectin pathway that included parameters sensitive to pathway initiation (Fig. \ref{fig-SA}C and D).
+One key difference observed between the sensitivity of C3a and C5a parameters, was their respective degradation constants.
 The rate constant governing C3a degradation was sensitive, while the degradation constant for C5a was not.
-This difference was likely attributable to the magnitude of the degradation parameters and the respective concentration of C3a and C5a.
-Taken together, global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+This difference was likely attributable to the magnitude of the degradation parameters and the respective concentrations of C3a and C5a.
+Taken together, sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+
+We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-ROBUSTNESS}).
+We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors.
+We perturbed the initial concentration of C3, C5 and a combination of both by decreasing their concentrations by 50 \% and 90\%.
+We calculated corresponding robustness indices for C3a and C5a with no initiator (alternate pathway activation) and maximum initiator of lectin pathway (1g of zymosan).
+A robustness index of one indicated no effect due to the perturbation whereas an index of zero indicated a strong effect of the perturbation.
+In both cases (50 \% and 90\% knockdown) we see that C3a and C5a levels are strongly influenced by initial concentrations of C3 and C5 in the presence of lectin pathway initiator.
+The C3a generated from the alternate pathway also follows the same pattern where it is influenced by the initial concentration of C3.
+However we observe an interesting trend with respect to C5a generated in the alternate pathway.
+The amount of C5a decreases with decrease in C3 concentration.
+This suggests that the C3b that is generated in the alternate pathway is primarily involved in the formation of AP C5-convertase (and is a rate limiting step).
+This is in agreement with our sensitivity analysis where were we see that the AP C3-convertase primarily can affect the dynamics of C5a formation.
+At the same time, in the lectin initiated pathway, the robustness of C5a levels to initial C3 concentration in both cases (robustness index of 1 or almost 1 in the 90\%) indicates that small concentrations of C3 and C5 convertases can catalyze the downstream formation of C5a.
+Thus the formation of C5a in presence of an initiator remains unaffected until C3 reaches a very low concentration.
+The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.
+We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}
+
+%Could be in discussion - the below text
+%Using the sensitive parameters that was determined from the analysis, we constructed a visual representing possible pathways that are attractive possible therapeutic targets (Fig.
+\ref{fig-SA} (E)).
+The sensitivity provided insightful information about the how system output changes due to perturbations kinetics parameters which is extremely useful in strategically targeting the possible enzymes and proteases in the network, for instance C3 convertase.
+However, the sensitivity analysis fails to provide any information about how the concentration of each species influences the dynamics of the system.
+Since many therapeutic dugs are antibody based and targets small molecules for inactivation  and decrease the total active concentration(Fig \ref{fig-SA}(E)), a computational method for analyzing the system influence contributed by active species concentration is of need.
+%To overcome those shortcomings, we conducted a robustness analysis to interrogate the effects of decreasing C3 and C5 concentration in our simulation to simulate the impact of their respective antibody based drugs (Fig \ref{fig-ROBUSTNESS}).
+
+%To overcome those shortcomings, we conducted a robustness analysis to interrogate the effects of decreasing C3 and C5 concentration in our simulation to simulate the impact of their respective antibody based drugs (Fig \ref{fig-ROBUSTNESS}).
+We calculated the robustness index for C3a and C5a using the ratio of area under the curve of perturbed and unperturbed initial C3 and/or C5 concentration (see Material and Methods for details) for the ensemble parameter set of 50.
+We perturbed our initial concentrations of C3, C5, and a combination of both by decreasing the concentration by 50 \% and 90\% to investigate how the perturbation effects the model predictions.
+As expected, most of the results responded proportionally to the decrease in C3 or C5 concentrations.
+However one unexpected observation was the C5a alternate response to decrease in C3 concentration while the C5a lectin is completely robust to the change in C3 concentration (Fig \ref{fig-ROBUSTNESS} (A and B)).
+This suggest that in the absence of zymosan, the production of C3b for the formation of C5 convertase of the alternate pathway may be rate limiting and the C3b.
+Interesting the robustness of C5a to C3 concentration in the presence of zymosan is close to unity meaning that changes in initial C3 concentration does not effect C5a formation.
+This is due to the excess free floating C3b compared to the C3 convertase and C5 convertase.
+With a decreased concentration of C3, the formation of C5a in the presence of zymosan is unaffected until it reaches a critical point where C3 concentration starts to play a role in C5a formation as shown with a 90 \% decrease in C3 initial concentration (Fig \ref{fig-ROBUSTNESS}(B)).
+%Add Comment
 
@@
 The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
-Using a simple and versatile modeling approach, we developed a reduced order complement model is computationally inexpensive, and versatile so it could easily be incorporated into pre-existing or new pharmacokinetic models.
+Subsequently we performed a robustness analysis where we analyzed the effect of perturbing initial C3 and C5 levels on the total amount of C3a and C5a generated.
+Using a simple and versatile modeling approach, we developed a reduced order complement model that is computationally inexpensive, and one that could easily be incorporated into pre-existing or new pharmacokinetic models.
 Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
@@
 
-Global sensitivity analysis of the complement model estimated potential important therapeutic targets.
+Global analysis of the complement model estimated potential important therapeutic targets.
 Complement malfunctions are implicated in a number of diseases, however the
-development of complement specific therapeutics has been challenging \cite{ricklin2007complement}.
+development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
 Previously, we have shown that mathematical modeling and sensitivity analysis can be useful tools to estimate therapeutically important mechanisms in biochemical networks \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
@@
 and parameters governing the conversion of C5 to C5a and C5b.
-This result agrees well with the current protease inhibitors targeting initiating complexes, including mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{morgan2015complement}.
+This result agrees well with the current protease inhibitors targeting initiating complexes, including mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{heja2012monospecific}.
 The most commonly used anti-complement drug eculizumab \cite{morgan2015complement}, targets the C5 protein which is cleaved to form C5a.
@@
 All these mechanisms are potential drug targets.
-For example, both C4BP and C3 convertase inhibitors are currently being developed as next generation complement inhibitors [REF ORIGINAL STUDIES NOT REVIEW].
-Thus, analysis of the ensemble of complement models identified potentially important therapeutic targets.
+
+To further validate these results from sensitivity analysis about potential drug targets we did a robustness analysis.
+We knocked down C3 and C5 levels and studied their impact on the generation of C3a and C5a.
+The C3a and C5a levels in the lectin pathway were strongly influenced by initial levels of C3 and C5.
+Thus direct inhibition of C3 and C5, or targeting complexes (MASP complex, C3 and C5 convertases) that act on C3 and C5 have a direct impact on production of C3a and C5a.
+This is also in agreement with sensitivity analysis that C5 is a good drug target.
+A number of drugs targeting C5 are being developed.
+For example LFG316 by Novartis is being used to target C5 in cases of Age-Related Macular Degeneration \cite{roguska2014generation}, Mubodina is an antibody that targets C5 in the treatment of Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}, Coversin is a small molecule targeting C5 \cite{weston2014clinical}, Zimura is an aptamer targeting C5 \cite{epstein2007complement}, small peptides and RNAi are also being used to inhibit C5 \cite{borodovsky2014aln}.
+Another important conclusion that can be drawn together from sensitivity and robustness analysis is that C3 and C5 convertases can be important therapeutic targets.
+Though knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate \cite{sissons1977metabolism, swaak1982determination} of these proteins make them difficult targets.
+Thus targeting C3 and C5 directly will require high dosage of drugs.
+It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
+The issue of high dosage can be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
+Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
+Formation of convertases is inhibited by targeting upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
+For example, Omeros is a protease inhibitor that targets MASP-2 complex and thereby inhibits formation of downstream convertases \cite{schwaeble2011methods}.
+Lampalizumab (an immunoglobulin) and Bikaciomab (an antibody fragment) target Factor B and Factor D respectively.
+Factor B and Factor D are crucial to formation alternate pathway convertases \cite{katschke2012inhibiting, hu2013therapeutic}.
+Novelmed Therapeutics recently developed antibody,NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
+Cobra Venom Factor (CVF), an analogue of C3b has been used to bind to Factor B to regulate alternate convertases \cite{vogel2004recombinant}.
+Thus, analysis of the ensemble of complement models identified potentially important therapeutic targets that are consistent with therapeutic strategies that are under development.
+
+%Indeed a number of next generation drugs are targeting convertases directly \cite{}, inhibiting their assembly or formation \cite{} or targeting the individual convertase components \cite{}
 
@@
 One potential approach to address such difficulties would be to incorporate C reactive proteins (CRP) and L-ficolin (LF) into the model,
-both of which are involved with the initiation of classical and lectin pathways [HOW DOES THIS HELP THE CROSSTALK ISSUE?].
+both of which are involved with the initiation of classical and lectin pathways.
+Liu et al.
+showed that under inflammation conditions interactions between lectin and classical pathways was mediated through CRP and LF  \cite{liu2011computational}.
+Thus incorporating these two proteins would help us in modeling cross talk.
 Time course measurements of MAC abundance (and MAC formation dynamics) are also scarce, making the inclusion of MAC challenging.
@@
 At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ for our model which we use for plotting best fits against the experimental data.
-We perturb this parameter vector to generate an ensemble of parameter vectors.
-The quality of parameter estimates was measured using goodness of fit (model residual).
 The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org.
+
+We generated an ensemble of parameter vectors by perturbing the optimal parameter vector obtained at the end of DDS phase.
+Through repeated perturbations of this vector we roughly identified sensitive and robust parameters within the parameter vector.
+This enabled us to approximately determine the bounds on the perturbation of each parameter.
+Thereafter the optimal parameter vector was perturbed within these bounds for 100,000 iterations.
+Within each iteration the quality of perturbed vector was measured using goodness of fit (model residual).
+If the residual was too high or generated a numerical error the vector was rejected.
+In this fashion we generated an ensemble of 50 parameter vectors
 
@@
 The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
-
 \clearpage
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fit.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final.pdf}
 \caption{Reduced order complement model training simulations.
@@
 The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
-The shaded region denotes 99 \% confidence interval on the simulated mean concentration of C3a or C5a (uncertainty in the model simulation).
+The  dark shaded region denotes 99 \% confidence interval of the simulated mean concentrations of C3a or C5a , while the light shaded region is the 99 \% confidence interval of the best prediction.
 All initial concentrations of complement proteins are at human serum levels unless otherwise noted.}\label{fig-fit}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure3_Predictions_2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final.pdf}
 \caption{Reduced order complement model predictions vs experimental data for C3a and C5a generated in the lectin pathway.
@@
 The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
-The shaded region denotes 99 \% confidence interval on the simulated mean concentration of C3a or C5a (uncertainty in the model simulation).
+The shaded region denotes 99 \% confidence interval of the simulated mean concentrations of C3a or C5a, while the light shaded region is the 99 \% confidence interval of the best prediction.
 All initial concentrations of complement proteins are at human serum levels unless otherwise noted.
@@
 
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_may25.pdf}
+%\caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
+Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0 zymosan, (B) C5a 0 zymosan, (C) C3a 1 g zymosan, and (D) C5a $1$ g zymosan.
+The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions.
+The error bars are the 95 percent confidence interval.
+$k$ represents association rate, $km$ denote Michaelis-Menten saturation constants, and $alpha$ and $n$ refers to the exponentials of the control functions.}\label{fig-SA}
+%\end{figure}
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_Sensitivity_Analysis_may25.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
@@
 The error bars are the 95 percent confidence interval.
-$k$ represents association rate, $km$ denote Michaelis-Menten saturation constants, and $alpha$ and $n$ refers to the exponentials of the control functions.}\label{fig-SA}
+Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.}\label{fig-SA}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis.pdf}
+\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations using 50 parameter sets.
+Robustness analysis was conducted on the four cases we used to train our model, C3a alternate (0 zymosan), C5a alternate (0 zymosan),  C3a lectin (1 g zymosan), and C5a lectin (1 g zymosan), by reducing the initial concentration of C3 and/or C5 by (A) 50 \% and (B) 90 \%.
+The bars denote robustness index which a measure of system changes from the perturbation of initial concentration that defined by the ratio of the area under the concentration curve of perturbed case and that of the unperturbed case.
+The error bars represent one standard deviation.
+At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-ROBUSTNESS}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./figs/Figure6_target.pdf}
+\caption{ The figure graphically illustrates some of the current complement therapeutics and their targets within our network.
+A vast majority of the drugs target the terminal pathway by targeting C5 or C5a.
+Targeting C5 would reduce the formation of C5a and C5b, however targeting C5a directly would reduce the influence of the anaphylatoxin but still produce precursors for the MAC formation.
+A number of therapeutics also target C3 and the formation of AP C3 convertase through the inhibition of Factors B and D.
+Finally, very few drugs target complement initiation, this may be due to a need for balance between immunity and inflammation or other complement related diseases.
+}\label{fig-targets}
 \end{figure}

### 2016-06-02 Paper_v4.tex
A few more changes

--- 2016-06-02 Paper_v2.tex
@@
 
+%Paragraph Edited by Mason
 We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-ROBUSTNESS}).
@@
 Thus the formation of C5a in presence of an initiator remains unaffected until C3 reaches a very low concentration.
+These results agree with the current push for complement therapeutics targeting C3 and C5.
+Conversely, the significance of the convertases identified by the sensitivity analysis indicate that they may also be effective targets for therapeutic intervention, however, these results require experimental validation before any definitive conclusions can be reached.
 The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.
@@
 After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
-The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance including convertase formation and activity as well as C3 and C5 inhibition.
 Subsequently we performed a robustness analysis where we analyzed the effect of perturbing initial C3 and C5 levels on the total amount of C3a and C5a generated.
@@
 It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
-The issue of high dosage can be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
+The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
 Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
@@
 The error bars represent one standard deviation.
-At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-ROBUSTNESS}
+At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-SA}
 \end{figure}

### 2016-06-03 Paper_v4.tex
A few more changes
Some initial changes on final draft

@@
 
-%Paragraph Edited by Mason
 \subsection*{Reduced order complement network.}
@@
 
-%Paragraph Edited by Mason
-We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-ROBUSTNESS}).
+We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-robustness-analysis}).
 We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors.
@@
 The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.
-We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}
+
+%We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}
 
@@
 
-\subsection*{Global sensitivity analysis of model performance.}
+\subsection*{Sensitivity and robustness analysis of model performance.}
 We conducted a global sensitivity analysis to estimate which parameters controlled the performance of the reduced order model using the Sobol method \cite{SOBOL_METHOD}.
@@
 The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
+
+Robustness coefficients were calculated as shown previously \cite{Tasseff:2011aa}.
+Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture.
+Robustness coefficients denoted by $\alpha\left(i,j,t_{o},t_{f}\right)$ are defined as:
+\begin{equation}
+\alpha\left(i,j,t_{o},t_{f}\right)=\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)
+\end{equation}
+Here $t_o$ and $t_f$ denote the initial and final simulation time respectively, while $i$ and $j$ denote the indices for the marker and the perturbation respectively.
+A value of $\alpha\left(i,j,t_{o},t_{f}\right)>1$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<1$ indicates decreased marker abundance following perturbation $j$.
+If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{1}$ the $j$th perturbation does not influence the abundance of marker $i$.
+
 \clearpage
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN_v2.pdf}
 \caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
-Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0 zymosan, (B) C5a 0 zymosan, (C) C3a 1 g zymosan, and (D) C5a $1$ g zymosan.
+Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0g zymosan, (B) C5a 0g zymosan, (C) C3a 1g zymosan, and (D) C5a 1g zymosan.
 The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions.
 The error bars are the 95 percent confidence interval.
-Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.}\label{fig-SA}
+Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.
+Red indicates current complement therapeutics.
+}\label{fig-SA}
 \end{figure}
@@
 The error bars represent one standard deviation.
-At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-SA}
+At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-robustness-analysis}
 \end{figure}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure6_target.pdf}
-\caption{ The figure graphically illustrates some of the current complement therapeutics and their targets within our network.
+% \begin{figure}[h]
+% \centering
+% \includegraphics[width=1.0\textwidth]{./figs/Figure6_target.pdf}
+% \caption{ The figure graphically illustrates some of the current complement therapeutics and their targets within our network.
 A vast majority of the drugs target the terminal pathway by targeting C5 or C5a.
@@
 Finally, very few drugs target complement initiation, this may be due to a need for balance between immunity and inflammation or other complement related diseases.
-}\label{fig-targets}
-\end{figure}
+% }\label{fig-targets}
+% \end{figure}
 

### 2016-06-03 Paper_v4.tex
Some more updates

@@
 and the correct order of magnitude for the concentration of C5a for three inducer levels.
-Next, we used global sensitivity analysis to determine which parameters controlled the performance of the complement model.
+Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.
 
@@
 We conducted sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
-We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA} (A-D)).
+We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}A - D).
 In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (formation of C3b) and $k_{d,C3a}$ (degradation rate constant governing C3a)
@@
 This difference was likely attributable to the magnitude of the degradation parameters and the respective concentrations of C3a and C5a.
-Taken together, sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+Thus, sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
+However, sensitivity coefficients are a local measure of how small changes in a parameter value effects a performance objective, for example
+the abundance of C5a. To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody we performed
+robustness analysis. Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture,
+in this case a decrease in the initial abundance of C3 and C5.
 
-We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-robustness-analysis}).
+We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig. \ref{fig-robustness-analysis}).
 We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors.
@@
 
+%Need to talk more about biochemical benefits and importance for of biochemical problems
+
 \clearpage
 
-%Need to talk more about biochemical benefits and importance for of biochemical problems
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Author's contributions}
+J.V directed the study. A.S developed the reduced order complement model.
+W.D and M.M analyzed model simulations and generated figures for the manuscript.
+The manuscript was prepared and edited for publication by A.S., W.D. M.M and J.V.
 
 \section*{Acknowledgements}
-This study was supported by an award from [FILL ME IN].
+We gratefully acknowledge the suggestions from the anonymous reviewers to improve this manuscript.
+
+\section*{Funding}
+This study was supported by an award from the US Army and Systems Biology of Trauma Induced Coagulopathy (W911NF-10-1-0376) to J.V. for the support of A.S.
+
 \clearpage
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v2.pdf}
 \caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations using 50 parameter sets.

### 2016-06-04 Paper_v4.tex
Working on robustness results

@@
 This elucidated that the activation of
+% We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a
+%
 
-%
 \subsection*{Global analysis of the reduced order complement model}
@@
 However, sensitivity coefficients are a local measure of how small changes in a parameter value effects a performance objective, for example
-the abundance of C5a. To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody we performed
-robustness analysis. Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture,
-in this case a decrease in the initial abundance of C3 and C5.
+the abundance of C5a. To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody, we performed
+robustness analysis. Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to the network architecture.
+In this case, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and C5.
 
-We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig. \ref{fig-robustness-analysis}).
-We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors.
-We perturbed the initial concentration of C3, C5 and a combination of both by decreasing their concentrations by 50 \% and 90\%.
-We calculated corresponding robustness indices for C3a and C5a with no initiator (alternate pathway activation) and maximum initiator of lectin pathway (1g of zymosan).
-A robustness index of one indicated no effect due to the perturbation whereas an index of zero indicated a strong effect of the perturbation.
-In both cases (50 \% and 90\% knockdown) we see that C3a and C5a levels are strongly influenced by initial concentrations of C3 and C5 in the presence of lectin pathway initiator.
-The C3a generated from the alternate pathway also follows the same pattern where it is influenced by the initial concentration of C3.
-However we observe an interesting trend with respect to C5a generated in the alternate pathway.
-The amount of C5a decreases with decrease in C3 concentration.
-This suggests that the C3b that is generated in the alternate pathway is primarily involved in the formation of AP C5-convertase (and is a rate limiting step).
-This is in agreement with our sensitivity analysis where were we see that the AP C3-convertase primarily can affect the dynamics of C5a formation.
-At the same time, in the lectin initiated pathway, the robustness of C5a levels to initial C3 concentration in both cases (robustness index of 1 or almost 1 in the 90\%) indicates that small concentrations of C3 and C5 convertases can catalyze the downstream formation of C5a.
-Thus the formation of C5a in presence of an initiator remains unaffected until C3 reaches a very low concentration.
-These results agree with the current push for complement therapeutics targeting C3 and C5.
-Conversely, the significance of the convertases identified by the sensitivity analysis indicate that they may also be effective targets for therapeutic intervention, however, these results require experimental validation before any definitive conclusions can be reached.
-The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.
+Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
+We calculated robustness indices for C3a and C5a for the 50 parameter sets in the ensemble with and without the lectin pathway initiator.
+We simulated the addition of different doses of anti-complement antibody cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
+A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
+As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
+(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
+However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
+This agreed with the sensitivity results; changes in AP C3-convertase formation or activity affected the downstream dynamics of C5a formation.
+Thus, these results suggested that C3 alone would be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig.
+\ref{fig-robustness-analysis}A or B, lane 2).
+Yet, in lectin initiated complement activation, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4).
+Thus, above some limiting threshold, even small concentrations of C3 and C5 convertases catalyzed the downstream formation of C5a.
+The only reliable intervention that consistently reduced C5a formation for all cases was a dual-knockdown.
+For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4).
 
+% These results agree with the current push for complement therapeutics targeting both C3 and C5.
+% Conversely, the significance of the convertases identified by the sensitivity analysis indicate that they may also be effective targets for therapeutic intervention, however, these results require experimental validation before any definitive conclusions can be reached.
+% The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.
 %We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}
+%For both a 50\% and 90\% knockdown, C3a and C5a levels were strongly influenced by initial concentrations of C3 and C5 in the presence of lectin pathway initiator.
+%The C3a generated from the alternate pathway also followed the same pattern where it was influenced by the initial concentration of C3.
+%Thus, the formation of C5a in presence of an initiator remained unaffected until C3 reached a limiting threshold concentration.
 

### 2016-06-04 Paper_v4.tex
Working on robustness results
Going to bed ...

@@
 DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
-At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ for our model which we use for plotting best fits against the experimental data.
+At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ which we use for plotting best fits against the experimental data, and for generating a parameter ensemble.
+
+An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by perturbing each parameter and measuring the increase in the residual.
+Thereafter, the optimal parameter vector was perturbed within these bounds for approximately 100,000 iterations.
+Within each iteration the quality of perturbed vector was measured using goodness of fit (model residual).
+If the residual was too high or the perturbed vector generated a numerical error, the vector was rejected.
+We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
 The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org.
-
-We generated an ensemble of parameter vectors by perturbing the optimal parameter vector obtained at the end of DDS phase.
-Through repeated perturbations of this vector we roughly identified sensitive and robust parameters within the parameter vector.
-This enabled us to approximately determine the bounds on the perturbation of each parameter.
-Thereafter the optimal parameter vector was perturbed within these bounds for 100,000 iterations.
-Within each iteration the quality of perturbed vector was measured using goodness of fit (model residual).
-If the residual was too high or generated a numerical error the vector was rejected.
-In this fashion we generated an ensemble of 50 parameter vectors
 
@@
 If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{1}$ the $j$th perturbation does not influence the abundance of marker $i$.
+In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\% or 90\% and measured the AUC of C3a or C5a with and without lectin initiator.
+Robustness coefficients were calculated for every member of the ensemble, the mean $\pm$ 1$\times$standard-deviation are reported.
 
@@
 \section*{Author's contributions}
-J.V directed the study. A.S developed the reduced order complement model.
+J.V directed the study. A.S developed the reduced order complement model and the parameter ensemble.
 W.D and M.M analyzed model simulations and generated figures for the manuscript.
-The manuscript was prepared and edited for publication by A.S., W.D. M.M and J.V.
+The manuscript was prepared and edited for publication by A.S, W.D, M.M and J.V.
 

### 2016-06-04 Paper_v4.tex
Small changes

@@
 Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
-After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
+After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
+were critical to model performance.
+These analyses suggested complement was robust to any single therapeutic intervention.
+The only reliable intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
 
@@
 
-The performance of the reduced order complement model was impressive given its limited size.
+The performance of the complement model was impressive given its limited size.
 However, there are several questions that should be explored further.
@@
 In this study, we used $\mathcal{I}_{j}\in\left\{min,max\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
-The model equations were implemented in MATLAB and solved using the \texttt{ODE15s} routine (The Mathworks, Natick MA).
-The complement model code and parameter ensemble can be downloaded from http://www.varnerlab.org.
+The model equations were implemented in MATLAB and solved using the \texttt{ODE23s} routine (The Mathworks, Natick MA).
+The complement model code and parameter ensemble is freely available under an MIT software license and can be downloaded from http://www.varnerlab.org.
 
@@
 \subsection*{Sensitivity and robustness analysis of model performance.}
-We conducted a global sensitivity analysis to estimate which parameters controlled the performance of the reduced order model using the Sobol method \cite{SOBOL_METHOD}.
-We computed the total sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases.
-We established the sampling bounds for each parameter from the minimum and maximum value for that parameter in the parameter ensemble.
+We conducted global sensitivity and robustness analysis to estimate which parameters and species controlled the performance of the reduced order model.
+We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases using the Sobol method \cite{SOBOL_METHOD}.
+The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
 We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
-where $N$ was the number of trials, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
-In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from 11,600 model evaluations.
+where $N$ was the number of trials per parameters, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
+In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed using 11,600 model evaluations.
 The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
 
+Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture.
 Robustness coefficients were calculated as shown previously \cite{Tasseff:2011aa}.
-Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture.
 Robustness coefficients denoted by $\alpha\left(i,j,t_{o},t_{f}\right)$ are defined as:
 \begin{equation}
-\alpha\left(i,j,t_{o},t_{f}\right)=\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)
+\alpha\left(i,j,t_{o},t_{f}\right)=\log_{10}\left[\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)\right]
 \end{equation}
 Here $t_o$ and $t_f$ denote the initial and final simulation time respectively, while $i$ and $j$ denote the indices for the marker and the perturbation respectively.
-A value of $\alpha\left(i,j,t_{o},t_{f}\right)>1$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<1$ indicates decreased marker abundance following perturbation $j$.
-If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{1}$ the $j$th perturbation does not influence the abundance of marker $i$.
-In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\% or 90\% and measured the AUC of C3a or C5a with and without lectin initiator.
-Robustness coefficients were calculated for every member of the ensemble, the mean $\pm$ 1$\times$standard-deviation are reported.
+A value of $\alpha\left(i,j,t_{o},t_{f}\right)>0$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<0$ indicates decreased marker abundance following perturbation $j$.
+If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{0}$ the $j$th perturbation did not influence the abundance of marker $i$.
+In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\% or 90\% and measured the area under the curve (AUC)
+of C3a or C5a with and without lectin initiator.
+Robustness coefficients were calculated for every member of the ensemble, where the mean $\pm$ 1$\times$standard-deviation are reported.
 

### 2016-06-07 Paper_v4.tex
Final draft

@@
 %\usepackage[authoryear,round,comma,sort&compress]{natbib}
-\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
+\usepackage[round,sort,comma,numbers,sort&compress]{natbib}
 %\usepackage[authoryear,round]{natbib}
@@
 \section*{Abstract}
-Complement is a central part of innate immunity which plays a significant role in the inflammatory response, and many other disease processes.
-In this study, we analyzed an ensemble of experimentally validated reduced order complement models.
-Our reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
-We used this framework to capture the dynamics of C3a and C5a formation in the lectin and alternative pathways.
-The reduced order model consisted of only 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
+Complement is an important pathway of innate immunity which plays a significant role in inflammation, and many other disease processes.
+However, despite its importance, there has been a paucity of validated mathematical models of complement activation.
+In this study, we developed an ensemble of experimentally validated reduced order complement models.
+The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
+The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
 We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
 Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
-Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
+Despite its small size, the model was surprisingly predictive.
 After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
-were critical to model performance.
-These analyses suggested complement was robust to any single therapeutic intervention.
-The only reliable intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
+The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+Taken together, we developed a reduced order complement model that was computationally inexpensive,
+and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
+The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
 
+% We used this framework to capture the dynamics of C3a and C5a formation from the lectin and alternative pathways.
 %and includes more pathways than comparable purely ODE models in the literature.
@@
 \vspace{0.1in}
-{\noindent \textbf{Keywords:}~Biochemical engineering, systems biology, reduced order models, complement system}
+{\noindent \textbf{Keywords:}~Complement system, systems biology, reduced order models, biochemical engineering}
 
@@
 
-In this study, we analyzed an ensemble of reduced order complement models.
-Our reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
-We used this framework to capture the dynamics of C3a and C5a formation in the lectin and alternative pathways.
-The reduced order model consisted of 18 differential equations with 28 parameters.
-The model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
-We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a from Morad and coworkers \cite{morad2015time}.
-Subsequently, we validated the model by predicting C3a and C5a measurements that were not used for model training.
-Given its small size, the hybrid approach produced a surprisingly predictive human complement model.
-After validation, we performed a global analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions and to identify potential therapeutic targets.
+In this study, we developed an ensemble of experimentally validated reduced order complement models.
+The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
+The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
+Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Despite its small size, the model was surprisingly predictive.
+After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
+controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
+The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+Taken together, we developed a reduced order complement model that was computationally inexpensive,
+and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
+The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
 %Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
@@
 %The influence of CRP, LF and the cross-talk can be captured through additional control functions that act upon the initiation pathways in a logical integration rule developed by Wayman and coworkers \cite{wayman2015dynamic}.
+% Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
 
@@
 
-In this study, we analyzed an ensemble of reduced order complement models.
-The reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
-The reduced order model consisted of only 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
-We used this framework to simulate the dynamics of C3a and C5a formation in the lectin and alternative pathways.
-We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a abundance.
+% In this study, we analyzed an ensemble of reduced order complement models.
+% The reduced order modeling approach combined ODEs with logical rules to produce a predictive model with a limited number of equations and parameters.
+% The reduced order model consisted of only 18 differential equations with 28 parameters.
+% Thus, the model was an order of magnitude smaller and included more pathways than comparable ODE models in the literature.
+% We used this framework to simulate the dynamics of C3a and C5a formation in the lectin and alternative pathways.
+% We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of C3a and C5a abundance.
+% Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+% Given its small size, the hybrid approach produced a surprisingly predictive complement model.
+% After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
+% The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance including convertase formation and activity as well as C3 and C5 inhibition.
+% Subsequently we performed a robustness analysis where we analyzed the effect of perturbing initial C3 and C5 levels on the total amount of C3a and C5a generated.
+
+%Given its small size, the hybrid approach produced a surprisingly predictive model.
+
+In this study, we developed an ensemble of experimentally validated reduced order complement models.
+The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
+The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
 Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
-Given its small size, the hybrid approach produced a surprisingly predictive complement model.
-After validation, we performed a global sensitivity analysis on the model ensemble to estimate which parameters were critical to model performance under different experimental conditions.
-The global sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance including convertase formation and activity as well as C3 and C5 inhibition.
-Subsequently we performed a robustness analysis where we analyzed the effect of perturbing initial C3 and C5 levels on the total amount of C3a and C5a generated.
-Using a simple and versatile modeling approach, we developed a reduced order complement model that is computationally inexpensive, and one that could easily be incorporated into pre-existing or new pharmacokinetic models.
-Furthermore this approach model has the potential to create individualized treatment plans for patients with complement deficiency.
+Despite its small size, the model was surprisingly predictive.
+After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
+controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
+The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+Taken together, we developed a reduced order complement model that was computationally inexpensive,
+and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
+The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
 

### 2016-06-07 Paper_v4.tex
Starting final round of changes

@@
 The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable models in the literature.
 We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
@@
 and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
-The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
+The model described experimental data, and predicted the need for multiple points of therapeutic intervention to disrupt complement activation.
 
@@
 We estimated the model parameters from in vitro human complement time-course experiments of C3a and C5a from Morad and coworkers, in the presence and absence of zymosan, and without the classical pathway.
-
 %We then compared the model predictions with C3a and C5a data sets, for alternative and lectin pathway dynamics, that were not used for model training.
@@
 
-Complement is a central part of innate immunity which plays a significant role in the inflammatory response.
-Complement was discovered in the 1890s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{sarma2011complement}.
+Complement is an important pathway in innate immunity. It plays a significant role in inflammation, host defense as well as many disease processes.
+Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
 However, research over the past decade has shown the importance of complement extends well beyond innate immunity.

### 2016-06-08 Paper_v4.tex
Introduction

@@
 \section*{Abstract}
-Complement is an important pathway of innate immunity which plays a significant role in inflammation, and many other disease processes.
+Complement is an important pathway of innate immunity which plays a significant role in inflammation, and many disease processes.
 However, despite its importance, there has been a paucity of validated mathematical models of complement activation.
@@
 Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
-However, research over the past decade has shown the importance of complement extends well beyond innate immunity.
-For example, complement contributes to tissue homeostasis by inducing growth factors involved in tissue repair \cite{ricklin2010complement}.
-Complement malfunctions have been linked with several diseases including Alzheimers, glaucoma, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
-Complement can also play both a positive and negative role in certain cancers; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
-Several other important biochemical networks are integrated with complement including the coagulation cascade, the autonomous nervous system and the ability to regulate inflammation \cite{ricklin2013complement}.
-Thus, complement is an important system involved in a variety of both beneficial and potentially harmful functions in the body.
+However, research over the past decade has shown that the importance of complement extends well beyond innate immunity.
+For example, complement contributes to tissue homeostasis by inducing tissue repair \cite{ricklin2010complement}.
+Complement has also been linked with several diseases including Alzheimers, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+Complement plays both positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
+Lastly, several other important biochemical subsystems are integrated with complement including the coagulation cascade, the autonomous nervous system and inflammation \cite{ricklin2013complement}.
+Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
 
-Complement is mediated by over 30 soluble and cell surface proteins that are present as inactive forms in the circulation \cite{walport2001complement}.
-The central output of complement activation is the formation of the Membrane Attack Complex (MAC) and a key protein called C5a.
-The membrane attack complex forms transmembrane channels which disrupt the cell membrane of targeted cells, leading to cell lysis and death.
-The C5a protein acts as a bridge between innate and adaptive immunity, and plays an important role in regulating inflammation and coagulation \cite{sarma2011complement}.
+Complement is mediated by over 30 soluble and cell surface proteins and regulators.
+The molecular connectivity of complement is complex, and has been extensively reviewed by Walport \cite{Walport:2001ab,Walport:2001aa}.
+The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediators C3a and C5a.
+The membrane attack complex forms transmembrane channels which disrupt the membrane integrity of targeted cells, leading to cell lysis and death.
+On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation and coagulation in the case of C5a \cite{sarma2011complement}.
 Complement activation takes places through three pathways: the alternate, the classical and the lectin binding pathway.
 Each of these pathways involves a different initiator signal which leads to a cascade of downstream events in the complement system.
-The classical pathway is triggered when antibodies form complexes with foreign antigens or other pathogens.
-A multimeric protein complex C1 binds to the antigen-antibody complex and undergoes a conformational change.
-This activated complex then cleaves complement proteins C4 and C2 into C4a, C4b, C2a and C2b respectively.
-The C4a and C2b fragments combine to form the C4bC2a protease, also known as the classical C3 convertase.
+The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
+A multimeric protein complex C1 binds to antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
+This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b respectively.
+The C4a and C2b fragments combine to form the C4bC2a protease, which is also known as the classical C3 convertase.
 The lectin binding pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
-This bound complex in turn cleaves C4 and C2, leading to the formation of C4bC2a.
-The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is hydrolyzed to form C3b.
-In the presence of pathogens, the C3b fragment binds foreign surfaces and recruits the additional proteins, factor B and factor D, which lead to the formation of C3bBb, the alternate C3 convertase \cite{pangburn1984alternative}.
-The formation of classical and alternate C3 convertases on bacterial surfaces is followed by the formation of proteases called C5 convertases.
-The classical and alternate C3 convertases recruit C3, Factor B and Factor D to form the classical C5 convertase (C4bC2aC3b), and alternate C5 convertase (C3bBbc3B) respectively.
-The C5 convertases then cleave C5 to form the C5a and C5b fragments.
-The cleavage of C5 is followed by a series of sequential cleavage steps involving the C6, C7, C8 and C9 complement proteins
-which combine with C5b to form the membrane attack complex \cite{ricklin2010complement}.
+These complexes also cleave C4 and C2, leading the formation of additional classical C3 convertase.
+Thus, the classical and lectin pathways converge at the classical C3 convertase, and are initiated by the recognition of a foreign surface.
+However, the final pathway, known as the alternate pathway, works differently.
+The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
+C3w recruits factor B and factor D, leading to the formation of C3wBb which cleaves C3 into C3a and C3b.
+The C3b fragment can further recruit additional factor B and factor D to form C3bBbC3b, otherwise known as the alternate C3 convertase \cite{pangburn1984alternative}.
+The role of classical and alternative C3 convertases is varied.
+First, the C3 convertases encode an amplification loop by cleaving C3 into C3a and C3b, the latter is then free to form additional alternate C3 convertase.
+Next, they link complement initiation with the terminal response of the cascade through the formation of C5 convertases.
+The classical and alternate C3 convertases recruit C3b, form the classical C5 convertase (C4bC2aC3b), and alternate C5 convertase (C3bBbC3b) respectively.
+The C5 convertases cleave C5 to form the C5a and C5b fragments.
+The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
+On the other hand, both C3a and C5a are important inflammatory signals involved in several responses.
 

### 2016-06-08 Paper_v4.tex
Updates to the introduction and materials and methods

@@
 
-Complement is mediated by over 30 soluble and cell surface proteins and regulators.
-The molecular connectivity of complement is complex, and has been extensively reviewed by Walport \cite{Walport:2001ab,Walport:2001aa}.
-The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediators C3a and C5a.
-The membrane attack complex forms transmembrane channels which disrupt the membrane integrity of targeted cells, leading to cell lysis and death.
-On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation and coagulation in the case of C5a \cite{sarma2011complement}.
-Complement activation takes places through three pathways: the alternate, the classical and the lectin binding pathway.
-Each of these pathways involves a different initiator signal which leads to a cascade of downstream events in the complement system.
+The complement cascade involves over 30 soluble and cell surface proteins, receptors and regulators.
+The molecular connectivity of complement is complex, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
+The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
+The membrane attack complex, generated during the terminal phase of the response,
+forms transmembrane channels which disrupt the membrane integrity of targeted cells, leading to cell lysis and death.
+On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation \cite{sarma2011complement}.
+Complement activation takes places through three pathways: the classical, the lectin binding and the alternate pathways.
+Each of these pathways involves a different initiator signal which triggers downstream events in the complement system.
 The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
 A multimeric protein complex C1 binds to antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
-This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b respectively.
-The C4a and C2b fragments combine to form the C4bC2a protease, which is also known as the classical C3 convertase.
-The lectin binding pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
-These complexes also cleave C4 and C2, leading the formation of additional classical C3 convertase.
-Thus, the classical and lectin pathways converge at the classical C3 convertase, and are initiated by the recognition of a foreign surface.
-However, the final pathway, known as the alternate pathway, works differently.
+This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
+The C4a and C2b fragments bind to form the C4bC2a protease, which is also known as the classical C3 convertase.
+The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
+These complexes, in combination with the associated mannose-associated serine proteases
+1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional classical C3 convertase.
+Thus, the classical and lectin pathways, initiated by the recognition of a foreign surface,  converge at the classical C3 convertase.
+However, the alternate pathway works differently.
 The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
-C3w recruits factor B and factor D, leading to the formation of C3wBb which cleaves C3 into C3a and C3b.
-The C3b fragment can further recruit additional factor B and factor D to form C3bBbC3b, otherwise known as the alternate C3 convertase \cite{pangburn1984alternative}.
-The role of classical and alternative C3 convertases is varied.
-First, the C3 convertases encode an amplification loop by cleaving C3 into C3a and C3b, the latter is then free to form additional alternate C3 convertase.
-Next, they link complement initiation with the terminal response of the cascade through the formation of C5 convertases.
-The classical and alternate C3 convertases recruit C3b, form the classical C5 convertase (C4bC2aC3b), and alternate C5 convertase (C3bBbC3b) respectively.
-The C5 convertases cleave C5 to form the C5a and C5b fragments.
+C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb can cleave C3 into C3a and C3b, where
+the C3b fragment can further recruit additional factor B and factor D to form C3bBbC3b, which is also known as the alternate C3 convertase \cite{pangburn1984alternative}.
+The role of classical and alternate C3 convertases is varied.
+First, C3 convertases encode an amplification loop by cleaving C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
+thereby forming a positive feedback loop.
+Next, C3 convertase activity links complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
+Both classical and alternate C3 convertases can recruit an additional C3b subunit to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
+C5 convertases cleave C5 into the C5a and C5b fragments.
 The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
@@
 
-Complement activation is regulated by many plasma and host cell proteins.
+%Complement activation is largely irreversible, involving positive feedback driven proteolytic cleavage.
+%Thus,
+Activation of the complement cascade is strongly regulated by many plasma and host cell proteins.
 The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
@@
 Factor B and Factor D are crucial to formation alternate pathway convertases \cite{katschke2012inhibiting, hu2013therapeutic}.
-Novelmed Therapeutics recently developed antibody,NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
+Novelmed Therapeutics recently developed antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
 Cobra Venom Factor (CVF), an analogue of C3b has been used to bind to Factor B to regulate alternate convertases \cite{vogel2004recombinant}.
@@
 where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\sigma_{sj}$ denotes the stoichiometric coefficient for species $s$ in reaction $j$,
-and $s\in{m_{j}}$ denotes the set of \textit{reactants} for reaction $j$.
+and $s\in{m_{j}}$ denotes the set of \textit{reactants} for reaction $j$. We assumed all binding interactions were irreversible.
 The control terms $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
@@
 
-\subsection*{Estimation of an ensemble of model parameters.}
+\subsection*{Estimation of an ensemble of complement model parameters.}
 We minimized the residual between simulations and experimental C3a and C5a measurements using Dynamic Optimization with Particle Swarms (DOPS). DOPS minimized the objective:
@@
 We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
-The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org.
+The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
 
@@
 
-\subsection*{Sensitivity and robustness analysis of model performance.}
+\subsection*{Sensitivity and robustness analysis of complement model performance.}
 We conducted global sensitivity and robustness analysis to estimate which parameters and species controlled the performance of the reduced order model.
@@
 Robustness coefficients were calculated as shown previously \cite{Tasseff:2011aa}.
-Robustness coefficients denoted by $\alpha\left(i,j,t_{o},t_{f}\right)$ are defined as:
+Log-transformed robustness coefficients denoted by $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)$ are defined as:
 \begin{equation}
-\alpha\left(i,j,t_{o},t_{f}\right)=\log_{10}\left[\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)\right]
+\hat{\alpha}\left(i,j,t_{o},t_{f}\right)=\log_{10}\left[\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)\right]
 \end{equation}
-Here $t_o$ and $t_f$ denote the initial and final simulation time respectively, while $i$ and $j$ denote the indices for the marker and the perturbation respectively.
-A value of $\alpha\left(i,j,t_{o},t_{f}\right)>0$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<0$ indicates decreased marker abundance following perturbation $j$.
-If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{0}$ the $j$th perturbation did not influence the abundance of marker $i$.
+Here $t_o$ and $t_f$ denote the initial and final simulation time, while $i$ and $j$ denote the indices for the marker and the perturbation, respectively.
+A value of $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)>0$, indicates increased marker abundance, while $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)<0$ indicates decreased marker abundance following perturbation $j$.
+If $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)\sim{0}$, perturbation $j$ did not influence the abundance of marker $i$.
 In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\% or 90\% and measured the area under the curve (AUC)
 of C3a or C5a with and without lectin initiator.
-Robustness coefficients were calculated for every member of the ensemble, where the mean $\pm$ 1$\times$standard-deviation are reported.
+Log-transformed robustness coefficients were calculated for every member of the ensemble, where the mean $\pm$ 1$\times$standard-deviation are reported.
 
@@
 J.V directed the study. A.S developed the reduced order complement model and the parameter ensemble.
-W.D and M.M analyzed model simulations and generated figures for the manuscript.
+A.S, W.D and M.M analyzed the model ensemble, and generated figures for the manuscript.
 The manuscript was prepared and edited for publication by A.S, W.D, M.M and J.V.

### 2016-06-08 Paper_v4.tex
First p of Results

@@
 \subsection*{Reduced order complement network.}
-The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
+The complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
 A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
 Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
-Activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
-The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
+The alternate pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b (not C3w).
+The C3b fragment generated by hydrolysis (or by CP C3 convertase) could then form the alternate pathway (AP) C3 convertase (C3bBb).
+We did not consider C3w, nor the formation of the initial alternate C3 convertase (C3wBb).
+Rather, we assumed C3w was equivalent to C3b and only modeled the formation of the main AP C3 convertase.
 Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
-A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
-The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
-Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
-Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
+A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP or AP C5 convertase (C4bC2aC3b or C3bBbC3b).
+Both C5 convertases catalyzed the cleavage of C5 into the C5a and C5b fragments.
+In this initial study, we simplified the model by assuming both Factor B and Factor D were in excess.
+However, we did explicitly account for two control proteins, Factor H and C4BP.
+Lastly, we did not consider MAC formation, instead we stopped at C5a and C5b.
+Lectin pathway activation, and C3/C5 convertase activity was modeled using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
+Binding interactions were modeled using mass-action kinetics, where we assumed all binding was irreversible.
+Thus, while the reduced order complement model encoded significant biology, it was highly compact consisting of only 18 differential equations and 28 model parameters.
 Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.
@@
 \section*{Materials and Methods}
+
+\subsection*{Formulation and solution of the complement model equations.}
 We used ordinary differential equations (ODEs) to model the time evolution of complement proteins ($x_{i}$) in the reduced order model:

### 2016-06-08 Paper_v4.tex
Ensemble p

@@
 \subsection*{Estimating an ensemble of reduced order complement models.}
-A critical challenge for any dynamic model is the estimation of model parameters.
-We estimated the complement model parameters in a hierarchical fashion using two \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
-The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) approach,
-starting from an initial random parameter guess.
-A hierarchical approach was taken to determine model parameters in which the alternate pathway parameters were first estimated and then fixed during the estimation of the lectin pathway parameters.
-The reduced order complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
-For the alternative pathway, we used the C3a and C5a measurements in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit}A and B).
-Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit}C and D).
-Taken together, the reduced order model reproduced a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
+A critical challenge for the development of any dynamic model is the estimation of model parameters.
+We estimated an ensemble of complement model parameters in a hierarchical fashion using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
+The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) routine,
+starting from an initial random parameter guess. Unless otherwise specified, all initial conditions were assumed to be their mean physiological values.
+A hierarchical approach was taken in which the alternate pathway parameters were estimated first and then fixed during the estimation of the lectin pathway parameters.
+While we had significant training data, the parameter estimation problem was underdetermined (we were not able to uniquely determine model parameters).
+Thus, instead of using a best-fit yet uncertain parameter set, we estimated an ensemble of probable parameter sets (N = 50, see materials and methods).
+The reduced order complement model ensemble captured the behavior of both the alternative and lectin pathways (Fig. \ref{fig-fit}).
+For the alternative pathway, we used C3a and C5a measurements in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit}A and B).
+On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1g zymosan with alternate pathway parameters fixed (Fig. \ref{fig-fit}C and D).
+The reduced order model reproduced a panel of alternate and lectin pathway data sets in the neighborhood of physiological factor and inhibitor concentrations.
 However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
-To address this question, we fixed the model parameters and simulated data not used for model training.
+To address this question, we fixed the model parameters and simulated data sets not used for model training.
 

### 2016-06-08 Paper_v4.tex
Captions

@@
 We tested the predictive power of the reduced order complement model with data not used during model training (Fig. \ref{fig-prediction}).
-Six validation cases were considered, three for C3a and C5a respectively at different zymosan concentrations.
-All model parameters were fixed for the validation simulations.
-The ensemble of reduced order models captured the qualitative dynamics of C3a formation (Fig.
-\ref{fig-prediction}, left column), and C5a formation (Fig.
-\ref{fig-prediction}, right column) at three inducer concentrations.
-However, there were shortcomings, especially for the C3a prediction.
-First, while the C3a dynamics and concentration peak times were captured, the overall level of C3a was under-predicted in all cases (Fig. \ref{fig-prediction}, inset left column).
-We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
-C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes;
-however, the binding of C4BP with complement proteins is likely reversible and convertases may have residual activity even in the bound form.
-Thus, the model may over-predict the influence of C4BP.
-We also failed to capture the concave down curvature for the 0.001~g and 0.01~g zymosan cases in the C5a validation studies.
-The decreasing slope of the C5a measurements may indicate decreasing cofactors abundance, or missing biology which we have not explicitly accounted for in the reduced order approach.
-However, despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the dynamic time scale of C3a formation,
+Six validation cases were considered, three for C3a and C5a, respectively.
+All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan).
+The ensemble of reduced order models predicted the qualitative dynamics of C3a formation (Fig. \ref{fig-prediction}, left column),
+and C5a formation (Fig. \ref{fig-prediction}, right column) at three inducer concentrations.
+However, there were shortcomings with model performance, especially for the C3a prediction.
+First, while the C3a trend was captured, the overall level of C3a was consistently under-predicted in all cases (Fig. \ref{fig-prediction}, left column).
+We believe the C3a under-prediction was related to how we modeled C4BP interactions;
+C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes.
+However, the binding of C4BP with complement proteins is likely reversible, and CP C3 convertase may have residual activity even in the bound form.
+Thus, the model likely over-predicted the influence of C4BP.
+Second, while the C5a measurements were within the 95\% confidence estimate, we failed to capture the concave down curvature for the 0.001~g and 0.01~g zymosan cases.
+The decreasing slope of the C5a may indicate cofactor limitation, or missing biology which we have not accounted for in the reduced order approach.
+Despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the dynamic time scale of C3a formation,
 and the correct order of magnitude for the concentration of C5a for three inducer levels.
@@
 \includegraphics[width=1.0\textwidth]{./figs/Fig1_Schematic_v2.pdf}
-\caption{Simplified schematic of the human complement system.
-The complement cascade is activated through any one, or more, of the three pathways:  the classical, the lectin, and the alternate pathways.
-The classical pathway is activated by the binding of C1 complex through the C1q subunit to the IgG or IgM immune complex.
-This binding leads to conformational changes in the C1 complex that leads to the activation of C1r and C1s subunits.
-Activated C1-antibody complex cleaves C4 and C2 to form the classical C3 convertase.
-The lectin pathway is initiated by the binding mannose-binding lectins (MBL) and ficolins to carbohydrate moieties on the pathogen surfaces.
-This results in the formation mannose-binding lectin-associated serine proteases (MASPs).
-The MBL-MASP complex cleaves C4 and C2 to form the lectin C3 convertase.
-The alternative pathway is activated through a spontaneous tick-over mechanism by the hydrolysis of C3 to form fluid phase C3 convertase.
-The C3 convertases  cleaves C3 into C3a, and C3b.
-C3b combines with C4b and C2a to form classical C5 convertase (C4bC3aC3b).
-The C3b binds with Factor B to form the alternate C5 convertase (C3bBbC3b) .
-The C5 convertases cleave C5 into C5a, and C5b that undergoes a series of reactions to form the membrane attack complex (MAC).}\label{fig-schematic}
+\caption{Simplified schematic of the human complement system. The complement cascade is activated through three pathways:  the classical, the lectin, and the alternate pathways.
+Complement initiation results in the formation of classical or alternative C3 convertases, which amplify the initial complement response and signal to the adaptive immune system by cleaving C3 into C3a and C3b.
+C3 convertases further react to form C5 convertases which catalyze the cleavage of the C5 complement protein to C5a and C5b.
+C5b is critical to the formation of the membrane attack complex (MAC), while C5a recruits an adaptive immune response.}\label{fig-schematic}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final.pdf}
-\caption{Reduced order complement model training simulations.
-Reduced order complement model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS).
-The model was trained against experimental data from Shaw and co-workers \cite{morad2015time} in the presence and absence of zymosan.
-The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin initiated pathway with 1g zymosan (\textbf{C}--\textbf{D}).
-The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
-The  dark shaded region denotes 99 \% confidence interval of the simulated mean concentrations of C3a or C5a , while the light shaded region is the 99 \% confidence interval of the best prediction.
-All initial concentrations of complement proteins are at human serum levels unless otherwise noted.}\label{fig-fit}
+\caption{Reduced order complement model training.
+Model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS) from C3a and C5a measurements with and without zymosan \cite{morad2015time}.
+The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin pathway initiated with 1g zymosan (\textbf{C}--\textbf{D}).
+The solid red line shows the simulation with the best-fit parameter set, the solid black lines show the simulated mean value of C3a or C5a for the ensemble (N = 50).
+The dark shaded region denotes the 99\% confidence interval of the simulated mean concentrations, while the light shaded region denotes the 99\%
+confidence interval of the best-fit simulation for C3a and C5a. All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.}\label{fig-fit}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final.pdf}
-\caption{Reduced order complement model predictions vs experimental data for C3a and C5a generated in the lectin pathway.
-The reduced order coagulation model parameter estimates were tested against data not used during model training.
-Simulations of C3a and C5a generated in the lectin pathway using different levels of zymosan ($0.1$, $0.01$, and $0.001$ grams of zymosan) were compared with the corresponding experimental data (\textbf{A}--\textbf{F}).
-The solid red line shows the simulation with the best-fit parameter, the solid black lines show the simulated mean value of C3a or C5a for 50 independent particles.
-The shaded region denotes 99 \% confidence interval of the simulated mean concentrations of C3a or C5a, while the light shaded region is the 99 \% confidence interval of the best prediction.
-All initial concentrations of complement proteins are at human serum levels unless otherwise noted.
+\caption{Reduced order complement model predictions. The reduced order coagulation model parameter estimates were tested against data not used during model training \cite{morad2015time}.
+Simulations of C3a and C5a generated in the lectin pathway using 0.1g, 0.01g, and 0.001g zymosan were compared with the corresponding experimental measurement.
+The solid red line shows the simulation with the best-fit parameter set, the solid black lines show the simulated mean value of C3a or C5a for the ensemble (N = 50).
+The dark shaded region denotes the 99\% confidence interval of the simulated mean concentrations, while the light shaded region denotes the 99\%
+confidence interval of the best-fit simulation for C3a and C5a. All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
 }\label{fig-prediction}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN_v2.pdf}
-\caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.
-Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0g zymosan, (B) C5a 0g zymosan, (C) C3a 1g zymosan, and (D) C5a 1g zymosan.
-The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions.
-The error bars are the 95 percent confidence interval.
-Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.
-Red indicates current complement therapeutics.
-}\label{fig-SA}
+\caption{Global sensitivity analysis of the reduced order complement model.
+Sensitivity analysis was conducted on the four cases used for model training.
+\textbf{A}: Sensitivity of the C3a residual at 0g zymosan,
+\textbf{B}: Sensitivity of the C5a residual at 0g zymosan,
+\textbf{C}: Sensitivity of the C3a residual at 1g zymosan, and
+\textbf{D}: Sensitivity of the C5a residual at 1g zymosan.
+The bars denote the mean total sensitivity index for each parameter, while the error bars denote the 95\% confidence interval.
+\textbf{E}: Pathways controlled by the sensitivity parameters.
+Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.
+Red indicates current complement therapeutics. }\label{fig-SA}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v2.pdf}
-\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations using 50 parameter sets.
-Robustness analysis was conducted on the four cases we used to train our model, C3a alternate (0 zymosan), C5a alternate (0 zymosan),  C3a lectin (1 g zymosan), and C5a lectin (1 g zymosan), by reducing the initial concentration of C3 and/or C5 by (A) 50 \% and (B) 90 \%.
-The bars denote robustness index which a measure of system changes from the perturbation of initial concentration that defined by the ratio of the area under the concentration curve of perturbed case and that of the unperturbed case.
-The error bars represent one standard deviation.
-At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-robustness-analysis}
+\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations.
+Robustness analysis was conducted on the four cases used for model training:
+C3a alternate (0g zymosan), C5a alternate (0g zymosan),  C3a lectin (1g zymosan), and C5a lectin (1g zymosan), by reducing the initial concentration of C3 and/or C5
+by \textbf{A} 50\% and \textbf{B} 90 \%.
+The bars denote the log-transformated robustness index while error bars denote one standard deviation. At zero, the perturbed initial concentration has no impact on the measured output.
+A log-transformed robustness index less than zero indicates a negative relation between the perturbed initial concentration and the measured output.}\label{fig-robustness-analysis}
 \end{figure}

### 2016-06-09 Paper_v4.tex
Captions
Small changes

@@
 At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ which we use for plotting best fits against the experimental data, and for generating a parameter ensemble.
+The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
 
-An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by perturbing each parameter and measuring the increase in the residual.
+An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by analyzing the results of repeated DOPS runs.
 Thereafter, the optimal parameter vector was perturbed within these bounds for approximately 100,000 iterations.
 Within each iteration the quality of perturbed vector was measured using goodness of fit (model residual).
-If the residual was too high or the perturbed vector generated a numerical error, the vector was rejected.
+If the residual was too high or the perturbed vector generated a numerical error, the vector was rejected, otherwise it was accepted.
 We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
-The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
+The best fit parameter set, and parameter ensemble is available in tab-delimited plain-text format from the Complement model GitHub repository.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN_v2.pdf}
+\includegraphics[width=1.1\textwidth]{./figs/Figure4_SEN_v2.pdf}
 \caption{Global sensitivity analysis of the reduced order complement model.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v2.pdf}
-\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations.
+\includegraphics[width=0.9\textwidth]{./figs/Figure5_RobustnessAnalysis_v2.pdf}
+\caption{Robustness analysis of the complement model with respect to C3 and C5 initial conditions.
 Robustness analysis was conducted on the four cases used for model training:

### 2016-06-09 Paper_v4.tex
Updated the sensitvity

@@
 \vspace{0.1in}
-{\par \centering \textbf{To be submitted:}~\emph{PLoS~ONE}}
+{\par \centering \textbf{To be submitted:}~\emph{Scientific~Reports}}
 \vspace{0.5in}
@@
 \section*{Abstract}
-Complement is an important pathway of innate immunity which plays a significant role in inflammation, and many disease processes.
-However, despite its importance, there has been a paucity of validated mathematical models of complement activation.
+Complement is an important pathway in innate immunity, inflammation, and many disease processes.
+However, despite its importance, there have been few validated mathematical models of complement activation.
 In this study, we developed an ensemble of experimentally validated reduced order complement models.
-The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
-The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable models in the literature.
-We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
-Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+We combined ordinary differential equations with logical rules to produce a compact yet predictive complement model.
+The model,  which described the lectin and alternative pathways, was an order of magnitude smaller than comparable models in the literature.
+We estimated an ensemble of model parameters from \textit{in~vitro} dynamic measurements of the C3a and C5a complement proteins.
+Subsequently, we validated the model on unseen C3a and C5a measurements not used for model training.
 Despite its small size, the model was surprisingly predictive.
-After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
-controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
-The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
+Only dual dual-knockdown of both C3 and C5 consistently reduced C5a formation when all pathways were activated.
 Taken together, we developed a reduced order complement model that was computationally inexpensive,
@@
 The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
-On the other hand, both C3a and C5a are important inflammatory signals involved in several responses.
+On the other hand, both C3a and C5a are important inflammatory signals involved in several responses \cite{Walport:2001ab,Walport:2001aa}.
 
@@
 However, there were shortcomings with model performance, especially for the C3a prediction.
-First, while the C3a trend was captured, the overall level of C3a was consistently under-predicted in all cases (Fig. \ref{fig-prediction}, left column).
-We believe the C3a under-prediction was related to how we modeled C4BP interactions;
-C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes.
+First, while the overall C3a trend was captured, the level of C3a was consistently under-predicted (outside of our 99\% confidence interval) in all cases (Fig. \ref{fig-prediction}, left column).
+We believe the C3a under-prediction was related to our choice of training data, and how we modeled C4BP interactions.
+We trained the model using 1g zymosan, but predicted cases with much less initiator.
+Next, the C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes.
 However, the binding of C4BP with complement proteins is likely reversible, and CP C3 convertase may have residual activity even in the bound form.
 Thus, the model likely over-predicted the influence of C4BP.
-Second, while the C5a measurements were within the 95\% confidence estimate, we failed to capture the concave down curvature for the 0.001~g and 0.01~g zymosan cases.
+Second, while the C5a measurements were within the 99\% confidence estimate, we failed to capture the concave down curvature for the 0.001g and 0.01g zymosan cases.
 The decreasing slope of the C5a may indicate cofactor limitation, or missing biology which we have not accounted for in the reduced order approach.
@@
 
-\subsection*{Global analysis of the reduced order complement model}
+\subsection*{Global analysis of the reduced order complement model.}
 We conducted sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
-We calculated the sensitivity of the C3a and C5a residuals with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}A - D).
-In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (formation of C3b) and $k_{d,C3a}$ (degradation rate constant governing C3a)
-were largely responsible for the system response. Interestingly, $k_{c,C3}$ (the rate constant governing AP C3-convertase activity) was not sensitive in the absence of zymosan.
-Thus, the behavior of the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity.
-On the other hand, $k_{c,C3}$ was one of the parameters that controlled C5a formation, in addition to the expected parameters related to AP C5-convertase formation.
-The AP C3-convertase is required for AP C5-convertase formation, and the formation of the C3b fragment.
-Thus, changes in the activity of AP C3-convertase will not drastically change the C3a dynamics, but will effect AP C5-convertase activity and C5a formation.
-The sensitivity analysis yielded the expected results for the lectin pathway that included parameters sensitive to pathway initiation (Fig. \ref{fig-SA}C and D).
-One key difference observed between the sensitivity of C3a and C5a parameters, was their respective degradation constants.
-The rate constant governing C3a degradation was sensitive, while the degradation constant for C5a was not.
-This difference was likely attributable to the magnitude of the degradation parameters and the respective concentrations of C3a and C5a.
-Thus, sensitivity analysis identified important indirect parameter interactions that could have therapeutic significance.
-However, sensitivity coefficients are a local measure of how small changes in a parameter value effects a performance objective, for example
-the abundance of C5a. To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody, we performed
+We calculated the sensitivity of the C3a and C5a residuals to changes in model parameters with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}A - D).
+In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (tickover formation of C3b) and $k_{d,C3a}$ (rate constant governing C3a degradation)
+controlled the C3a fit (Fig. \ref{fig-SA}A). Interestingly, $k_{c,C3}$ (the rate constant governing AP C3-convertase activity) was not sensitive in the absence of zymosan.
+Thus, C3a formation in the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity, in the absence of zymosan.
+On the other hand, the C5a residual was sensitive to tickover parameters e.g., $k_{f,C3b}$, and $k_{f,C3}$ (the rate constant governing AP C3 convertase formation) as well as
+parameters controlling AP C5 convertase activity in the absence of zymosan (Fig. \ref{fig-SA}B).
+Thus, the AP C3-convertase did not strongly influence C3a dynamics, but did effect C5a formation in the absence of zymosan.
+In the presence of zymosan, the C3a residual was controlled by the formation and activity of the CP C3 convertase, as well as tickover and degradation parameters (Fig. \ref{fig-SA}C).
+Further, the C5a residual was influenced by lectin initiation parameters, the formation and activity of CP C5 convertase, and the inhibition by C4BP in the presence of zymosan (Fig. \ref{fig-SA}D).
+Globally, sensitivity analysis suggested that lectin pathway parameters dominated system performance in the presence of zymosan, and that C5a formation was sensitive to both
+C3 and C5 convertase activity.
+In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to C3 and C5 convertase activity.
+However, sensitivity coefficients are a local measure of how small changes in a parameter value effects model performance.
+To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody, we performed
 robustness analysis. Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to the network architecture.
@@
 Thus direct inhibition of C3 and C5, or targeting complexes (MASP complex, C3 and C5 convertases) that act on C3 and C5 have a direct impact on production of C3a and C5a.
-This is also in agreement with sensitivity analysis that C5 is a good drug target.
-A number of drugs targeting C5 are being developed.
-For example LFG316 by Novartis is being used to target C5 in cases of Age-Related Macular Degeneration \cite{roguska2014generation}, Mubodina is an antibody that targets C5 in the treatment of Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}, Coversin is a small molecule targeting C5 \cite{weston2014clinical}, Zimura is an aptamer targeting C5 \cite{epstein2007complement}, small peptides and RNAi are also being used to inhibit C5 \cite{borodovsky2014aln}.
+This is also in agreement with sensitivity analysis that C5 is a good drug target. A number of drugs targeting C5 are being developed.
+For example LFG316 by Novartis is being used to target C5 in cases of Age-Related Macular Degeneration \cite{roguska2014generation},
+Mubodina is an antibody that targets C5 in the treatment of Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement},
+Coversin is a small molecule targeting C5 \cite{weston2014clinical}, Zimura is an aptamer targeting C5 \cite{epstein2007complement},
+small peptides and RNAi are also being used to inhibit C5 \cite{borodovsky2014aln}.
 Another important conclusion that can be drawn together from sensitivity and robustness analysis is that C3 and C5 convertases can be important therapeutic targets.
 Though knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate \cite{sissons1977metabolism, swaak1982determination} of these proteins make them difficult targets.
-Thus targeting C3 and C5 directly will require high dosage of drugs.
-It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
+Thus targeting C3 and C5 directly will require high dosage of drugs. It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS),
+a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
 The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
-Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
-Formation of convertases is inhibited by targeting upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
+Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a. Formation of convertases is inhibited by targeting
+upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
 For example, Omeros is a protease inhibitor that targets MASP-2 complex and thereby inhibits formation of downstream convertases \cite{schwaeble2011methods}.

### 2016-06-09 Paper_v4.tex
More edits on sensitivity

@@
 robustness analysis. Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to the network architecture.
-In this case, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and C5.
+In this case, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
 
 Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
-We calculated robustness indices for C3a and C5a for the 50 parameter sets in the ensemble with and without the lectin pathway initiator.
+We calculated robustness indices for C3a and C5a for the parameter ensemble (N = 50) with and without the lectin pathway initiator.
 We simulated the addition of different doses of anti-complement antibody cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.

### 2016-06-09 Paper_v4.tex
Updates

@@
 In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to C3 and C5 convertase activity.
-However, sensitivity coefficients are a local measure of how small changes in a parameter value effects model performance.
-To more closely simulate a clinical intervention e.g., administration of an anti-complement antibody, we performed
-robustness analysis. Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to the network architecture.
-In this case, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
+Pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}E).
+However, sensitivity coefficients are only a local measure of how small changes in a parameter value affects model performance.
+To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
+robustness analysis.
 
+Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to a biochemical network.
+Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
 Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
 We calculated robustness indices for C3a and C5a for the parameter ensemble (N = 50) with and without the lectin pathway initiator.
-We simulated the addition of different doses of anti-complement antibody cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
+We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
+This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}E).
 A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
+The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
 As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
@@
 This agreed with the sensitivity results; changes in AP C3-convertase formation or activity affected the downstream dynamics of C5a formation.
-Thus, these results suggested that C3 alone would be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig.
+Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig.
 \ref{fig-robustness-analysis}A or B, lane 2).
-Yet, in lectin initiated complement activation, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4).
-Thus, above some limiting threshold, even small concentrations of C3 and C5 convertases catalyzed the downstream formation of C5a.
-The only reliable intervention that consistently reduced C5a formation for all cases was a dual-knockdown.
-For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4).
+Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
+C5a formation was catalyzed by CP C3 and C5 convertases.
+Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a dual-knockdown of C3 and C5.
+For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
+while C3a was reduced
 
@@
 This along with the particle best within the sub-swarm $\mathcal{S}_{k}$ is used to update the parameter estimate for each particle using the following rules:
-
 \begin{equation}\label{eqn:update-rule}
@@
 Sensitivity analysis was conducted on the four cases used for model training.
-\textbf{A}: Sensitivity of the C3a residual at 0g zymosan,
-\textbf{B}: Sensitivity of the C5a residual at 0g zymosan,
+\textbf{A}: Sensitivity of the C3a residual at w/o zymosan,
+\textbf{B}: Sensitivity of the C5a residual at w/o zymosan,
 \textbf{C}: Sensitivity of the C3a residual at 1g zymosan, and
@@
 \textbf{E}: Pathways controlled by the sensitivity parameters.
-Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.
-Red indicates current complement therapeutics. }\label{fig-SA}
+Bold black lines indicate the pathway involves one or more sensitive parameters, while the red lines show current therapeutics targets.
+Current complement therapeutics were taken from the review of Morgan and Harris \cite{morgan2015complement}. }\label{fig-SA}
 \end{figure}
@@
 Robustness analysis was conducted on the four cases used for model training:
-C3a alternate (0g zymosan), C5a alternate (0g zymosan),  C3a lectin (1g zymosan), and C5a lectin (1g zymosan), by reducing the initial concentration of C3 and/or C5
-by \textbf{A} 50\% and \textbf{B} 90 \%.
+C3a alternate (w/o zymosan), C5a alternate (w/o zymosan),  C3a lectin (1g zymosan), and C5a lectin (1g zymosan), by reducing the initial concentration of C3 and/or C5
+by 50\% or 90 \%.
+\textbf{A}: Robustness results for a 50\% decrease in the C3, C5, or C3 and C5 initial condition.
+\textbf{B}: Robustness results for a 90\% decrease in the C3, C5, or C3 and C5 initial condition.
 The bars denote the log-transformated robustness index while error bars denote one standard deviation. At zero, the perturbed initial concentration has no impact on the measured output.
@@
 
-\section*{Supplemental materials.}
-\subsection*{Model equations.}
-The reduced-order complement model consisted of 18 ordinary differential equations, 12 rate equations, and two control equations:
-\begin{eqnarray}
-	\frac{dx_{1}}{dt} & =& - r_{1}f_{1} \\ 								%C4
-	\frac{dx_{2}}{dt} &=& - r_{2}f_{2} \\  								%C2
-	\frac{dx_{3}}{dt} &=&  r_{1}f_{1} \\ 								%C4a
-	\frac{dx_{4}}{dt} &=& r_{1}f_{1} - r_{6} \\ 					%C4b
-	\frac{dx_{5}}{dt} & = & r_{2}f_{2} - r_{6} \\ 					%C2a
-	\frac{dx_{6}}{dt} &=& r_{2}f_{2} \\ 								%C2b
-	\frac{dx_{7}}{dt} &=& r_{3} - r_{4} - r_{5}\\ 				%C3
-	\frac{dx_{8}}{dt} &=& r_{3}  + r_{4} + r_{5}  - k_{deg,c3a}*C3a \\		%C3a
-	\frac{dx_{9}}{dt} &=& r_{3}  + r_{4} + r_{5}  -  r_{7} \\		%C3b
-	\frac{dx_{10}}{dt} &=& r_{6} - r_{10} -  r_{8} \\			%CP C3C
-	\frac{dx_{11}}{dt} &=& r_{7} - r_{11} -  r_{9} \\			%AP C3C
-	\frac{dx_{12}}{dt} &=& r_{10} - r_{14} \\							%CP C5C
-	\frac{dx_{13}}{dt} &=& r_{10}  \\											%AP C5C
-	\frac{dx_{14}}{dt} &=& - r_{12} - r_{13} \\								%C5
-	\frac{dx_{15}}{dt} &=&  r_{12} + r_{13} - k_{deg,c5a} \\					%C5a
-	\frac{dx_{16}}{dt} &=& r_{12} + r_{13} \\								%C5b
-	\frac{dx_{17}}{dt} &=& - r_{8} - r_{14} \\  				         %C4BP
-	\frac{dx_{18}}{dt} &=& - r_{9} \\									%Factor H
-\end{eqnarray}where the rate equations are given by:
-\begin{eqnarray}
-	r_{1} &=& \frac{k_{i1}(C4)}{(K_{1s} + C4)} \\
-	r_{2} &=& \frac{k_{2}(C2)}{(K_{2s} + C2)} \\
-	f_{1} &=& \frac{Zymo^{\eta_{1}}}{(Zymo^{\eta_{1}} + \alpha_{1}^{\eta_{1}})} \\
-	f_{2} &=& \frac{Zymo^{\eta_{2}}}{(Zymo^{\eta_{2}} + \alpha_{2}^{\eta_{2}})} \\
-	r_{3} &=& k_{3}(C3) \\
-	r_{4} &=& \frac{k_{4}(C3C_{L})(C3^{\eta_{3})}}{(K_{4s}^{\eta_{3}} + C3^{\eta_{3}})}  \\
-	r_{5} &=& \frac{k_{5}(C3C_{A})(C3)}{(K_{5s} + C3)} \\
-	r_{6} &=& k_{6}(C4b)(C2a) \\
-	r_{7} &=& k_{7}(C4b)(C2a) \\
-	r_{8} &=& k_{8}(C3C_{L})(C4b)(C4BP) \\
-	r_{9} &=& k_{9}(C3C_{A})(FactorH) \\
-	r_{10} &=& k_{10}(C3C_{L})(C3b) \\
-	r_{11} &=& k_{11}(C3C_{A})(C3b) \\
-	r_{12} &=& \frac{k_{12}(C5C_{L})(C5^{\eta_{4}})}{(K_{12s}^{\eta_{4}} + C5^{\eta_{4}})} \\
-	r_{13} &=& \frac{k_{13}(C5C_{A})(C5)}{(K_{13s} + C5)} \\
-	r_{14} &=& k_{14}(C5C_{L})(C4BP)
-\end{eqnarray}
+% \section*{Supplemental materials.}
+% \subsection*{Model equations.}
+% The reduced-order complement model consisted of 18 ordinary differential equations, 12 rate equations, and two control equations:
+% \begin{eqnarray}
+% 	\frac{dx_{1}}{dt} & =& - r_{1}f_{1} \\ 								%C4
+% 	\frac{dx_{2}}{dt} &=& - r_{2}f_{2} \\  								%C2
+% 	\frac{dx_{3}}{dt} &=&  r_{1}f_{1} \\ 								%C4a
+% 	\frac{dx_{4}}{dt} &=& r_{1}f_{1} - r_{6} \\ 					%C4b
+% 	\frac{dx_{5}}{dt} & = & r_{2}f_{2} - r_{6} \\ 					%C2a
+% 	\frac{dx_{6}}{dt} &=& r_{2}f_{2} \\ 								%C2b
+% 	\frac{dx_{7}}{dt} &=& r_{3} - r_{4} - r_{5}\\ 				%C3
+% 	\frac{dx_{8}}{dt} &=& r_{3}  + r_{4} + r_{5}  - k_{d,c3a}x_{8} \\		%C3a
+% 	\frac{dx_{9}}{dt} &=& r_{3}  + r_{4} + r_{5}  -  r_{7} \\		%C3b
+% 	\frac{dx_{10}}{dt} &=& r_{6} - r_{10} -  r_{8} \\			%CP C3C
+% 	\frac{dx_{11}}{dt} &=& r_{7} - r_{11} -  r_{9} \\			%AP C3C
+% 	\frac{dx_{12}}{dt} &=& r_{10} - r_{14} \\							%CP C5C
+% 	\frac{dx_{13}}{dt} &=& r_{10}  \\											%AP C5C
+% 	\frac{dx_{14}}{dt} &=& - r_{12} - r_{13} \\								%C5
+% 	\frac{dx_{15}}{dt} &=&  r_{12} + r_{13} - k_{d,c5a}x_{15} \\					%C5a
+% 	\frac{dx_{16}}{dt} &=& r_{12} + r_{13} \\								%C5b
+% 	\frac{dx_{17}}{dt} &=& - r_{8} - r_{14} \\  				         %C4BP
+% 	\frac{dx_{18}}{dt} &=& - r_{9} \\									%Factor H
+% \end{eqnarray}where the rate equations are given by:
+% \begin{eqnarray}
+% 	r_{1} &=& \frac{k_{i1}(C4)}{(K_{1s} + C4)} \\
+% 	r_{2} &=& \frac{k_{2}(C2)}{(K_{2s} + C2)} \\
+% 	f_{1} &=& \frac{Zymo^{\eta_{1}}}{(Zymo^{\eta_{1}} + \alpha_{1}^{\eta_{1}})} \\
+% 	f_{2} &=& \frac{Zymo^{\eta_{2}}}{(Zymo^{\eta_{2}} + \alpha_{2}^{\eta_{2}})} \\
+% 	r_{3} &=& k_{3}(C3) \\
+% 	r_{4} &=& \frac{k_{4}(C3C_{L})(C3^{\eta_{3})}}{(K_{4s}^{\eta_{3}} + C3^{\eta_{3}})}  \\
+% 	r_{5} &=& \frac{k_{5}(C3C_{A})(C3)}{(K_{5s} + C3)} \\
+% 	r_{6} &=& k_{6}(C4b)(C2a) \\
+% 	r_{7} &=& k_{7}(C4b)(C2a) \\
+% 	r_{8} &=& k_{8}(C3C_{L})(C4b)(C4BP) \\
+% 	r_{9} &=& k_{9}(C3C_{A})(FactorH) \\
+% 	r_{10} &=& k_{10}(C3C_{L})(C3b) \\
+% 	r_{11} &=& k_{11}(C3C_{A})(C3b) \\
+% 	r_{12} &=& \frac{k_{12}(C5C_{L})(C5^{\eta_{4}})}{(K_{12s}^{\eta_{4}} + C5^{\eta_{4}})} \\
+% 	r_{13} &=& \frac{k_{13}(C5C_{A})(C5)}{(K_{13s} + C5)} \\
+% 	r_{14} &=& k_{14}(C5C_{L})(C4BP)
+% \end{eqnarray}
 

### 2016-06-10 Paper_v4.tex
Small changes

@@
 In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (tickover formation of C3b) and $k_{d,C3a}$ (rate constant governing C3a degradation)
-controlled the C3a fit (Fig. \ref{fig-SA}A). Interestingly, $k_{c,C3}$ (the rate constant governing AP C3-convertase activity) was not sensitive in the absence of zymosan.
+controlled the C3a fit (Fig.
+\ref{fig-SA}A).
+Interestingly, neither $k_{c,C3C,AP}$ (the rate constant governing AP C3-convertase activity) nor $k_{f,C3C,AP}$ (the rate constant governing AP C3-convertase formation) were sensitive in the absence of zymosan.
 Thus, C3a formation in the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity, in the absence of zymosan.
-On the other hand, the C5a residual was sensitive to tickover parameters e.g., $k_{f,C3b}$, and $k_{f,C3}$ (the rate constant governing AP C3 convertase formation) as well as
+On the other hand, the C5a residual was strongly sensitive to tickover parameters e.g., $k_{f,C3b}$, and to a lesser extent $k_{f,C3C,AP}$ as well as
 parameters controlling AP C5 convertase activity in the absence of zymosan (Fig. \ref{fig-SA}B).
-Thus, the AP C3-convertase did not strongly influence C3a dynamics, but did effect C5a formation in the absence of zymosan.
+Thus, AP C3-convertase did not influence C3a dynamics, but strongly effected C5a formation in the absence of zymosan.
 In the presence of zymosan, the C3a residual was controlled by the formation and activity of the CP C3 convertase, as well as tickover and degradation parameters (Fig. \ref{fig-SA}C).
@@
 Globally, sensitivity analysis suggested that lectin pathway parameters dominated system performance in the presence of zymosan, and that C5a formation was sensitive to both
-C3 and C5 convertase activity.
-In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to C3 and C5 convertase activity.
-Pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}E).
+C3 and C5 convertase activity, while C3a was robust to the AP C3 convertase.
+In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to both AP C3 and C5 convertase activity.
+We compared the sensitivity results to current therapeutic approaches; pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}E).
+Thus, there was a qualitative overlap between sensitivity and the potential of biochemical efficacy.
 However, sensitivity coefficients are only a local measure of how small changes in a parameter value affects model performance.
 To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
-robustness analysis.
+robustness analysis on the model ensemble.
 
@@
 For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
-while C3a was reduced
+while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).
 
@@
 and $s\in{m_{j}}$ denotes the set of \textit{reactants} for reaction $j$. We assumed all binding interactions were irreversible.
+
 The control terms $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
@@
 At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ which we use for plotting best fits against the experimental data, and for generating a parameter ensemble.
+
+An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by analyzing repeated DOPS runs.
+Parameters were selected  using goodness of fit (model residual) and numerical stability criteria.
+We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
+The best fit parameter set, and parameter ensemble is available in a tab delimited plain text format from the Complement model GitHub repository.
 The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
-
-An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by analyzing the results of repeated DOPS runs.
-Thereafter, the optimal parameter vector was perturbed within these bounds for approximately 100,000 iterations.
-Within each iteration the quality of perturbed vector was measured using goodness of fit (model residual).
-If the residual was too high or the perturbed vector generated a numerical error, the vector was rejected, otherwise it was accepted.
-We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
-The best fit parameter set, and parameter ensemble is available in tab-delimited plain-text format from the Complement model GitHub repository.
 

### 2016-06-10 Paper_v4.tex
Updates

@@
 The decreasing slope of the C5a may indicate cofactor limitation, or missing biology which we have not accounted for in the reduced order approach.
-Despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the dynamic time scale of C3a formation,
-and the correct order of magnitude for the concentration of C5a for three inducer levels.
+Despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the time scale of C3a formation,
+and the correct order of magnitude for C5a for three inducer levels.
 Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.
@@
 C3 and C5 convertase activity, while C3a was robust to the AP C3 convertase.
-In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to both AP C3 and C5 convertase activity.
+In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to both AP C3 and C5 convertase formation.
 We compared the sensitivity results to current therapeutic approaches; pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}E).
 Thus, there was a qualitative overlap between sensitivity and the potential of biochemical efficacy.
-However, sensitivity coefficients are only a local measure of how small changes in a parameter value affects model performance.
+However, sensitivity coefficients are only a local measure of how small changes in a parameters affect model performance.
 To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
-robustness analysis on the model ensemble.
+robustness analysis.
 

### 2016-06-12 Paper_v4.tex
Updated robustness figure
Introduction is done

@@
 The complement cascade involves over 30 soluble and cell surface proteins, receptors and regulators.
-The molecular connectivity of complement is complex, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
+The molecular connectivity of complement involves both fluid phase and cell surface events, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
 The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
@@
 Complement activation takes places through three pathways: the classical, the lectin binding and the alternate pathways.
-Each of these pathways involves a different initiator signal which triggers downstream events in the complement system.
 The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
@@
 This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
-The C4a and C2b fragments bind to form the C4bC2a protease, which is also known as the classical C3 convertase.
+The C4a and C2b fragments bind to form the C4bC2a protease, also known as the classical C3 convertase.
 The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
@@
 1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional classical C3 convertase.
-Thus, the classical and lectin pathways, initiated by the recognition of a foreign surface,  converge at the classical C3 convertase.
-However, the alternate pathway works differently.
-The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
-C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb can cleave C3 into C3a and C3b, where
-the C3b fragment can further recruit additional factor B and factor D to form C3bBbC3b, which is also known as the alternate C3 convertase \cite{pangburn1984alternative}.
+Thus, the classical and lectin pathways, initiated by the recognition of different cues on foreign surfaces,  converge at the classical C3 convertase.
+However, the alternate pathway works differently. It is activated by a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
+C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb cleaves C3 into C3a and C3b, where
+the C3b fragment further recruits additional factor B and factor D to form C3bBbC3b, the alternate C3 convertase \cite{pangburn1984alternative}.
 The role of classical and alternate C3 convertases is varied.
-First, C3 convertases encode an amplification loop by cleaving C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
+First, C3 convertases mediate signal amplification.
+C3 convertases cleave C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
 thereby forming a positive feedback loop.
-Next, C3 convertase activity links complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
-Both classical and alternate C3 convertases can recruit an additional C3b subunit to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
+Next, C3 convertases link complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
+Both classical and alternate C3 convertases can recruit C3b subunits to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
 C5 convertases cleave C5 into the C5a and C5b fragments.
@@
 On the other hand, both C3a and C5a are important inflammatory signals involved in several responses \cite{Walport:2001ab,Walport:2001aa}.
+Thus, the complement cascade directly attacks the invading pathogen, and signals for help from the immune system.
 
@@
 %Thus,
-Activation of the complement cascade is strongly regulated by many plasma and host cell proteins.
-The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
-C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
-Regulation of the upstream elements of complement is also achieved through the interaction of the C4 binding protein (C4BP) with C4b, as well as
-through the interaction of factor H with C3b \cite{blom2001structural}.
-These regulatory proteins are also capable of binding their respective targets while they are bound in convertase complexes.
-Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
-Delay accelerating factor (DAF or CD55) is also able to recognize and dissociate both C3 and C5 convertases \cite{lukacik2004complement}.
-Carboxypeptidase-N, a well known inflammation regulator, cleaves carboxyl-terminal arginines and lysines of the complement proteins C3a, C4a, and C5a rendering them inactive \cite{liszewski1995control}.
-Lastly, the assembly of the MAC complex is inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
-Thus, there are many points of control which influence complement activation across the three activation pathways.
+The complement cascade is regulated by plasma and host cell proteins which balance host cell safety with effectiveness.
+The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh).
+C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing chronic complement activation \cite{walker1995complement}.
+Regulation of upstream processes in the lectin and alternate pathways also occurs through the interaction of the C4 binding protein (C4BP) with C4b, and factor H with C3b \cite{blom2001structural}.
+Interestingly,
+both factor H and C4BP are capable of binding their respective targets while in convertase complexes as well.
+At the host cell surface, membrane cofactor protein (MCP or CD46) can interact with C4b and C3b, which protects the host cell from complement self-activation \cite{riley2004cd46}.
+Delay accelerating factor (DAF or CD55) also recognizes and dissociates both C3 and C5 convertases on host cell surfaces \cite{lukacik2004complement}.
+More generally, Carboxypeptidase-N, a well known inflammation regulator, has broad activity against the complement proteins C3a, C4a, and C5a, rendering them inactive by cleavage of carboxyl-terminal arginine and lysine residues \cite{liszewski1995control}.
+Although Carboxypeptidase-N does not directly influence complement activation, it silences the important inflammatory signals produced by complement.
+Lastly, the assembly of the MAC complex itself can be inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
+Thus, there are many points of control which influence complement across the three activation pathways.
 
-Developing quantitative mathematical models of complement could be crucial to understanding its role in the body.
+Developing quantitative mathematical models of complement will be crucial to fully understanding its role in the body.
 Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
 For example, Hirayama et al. modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
-while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways as a system of non-linear ODEs \cite{korotaevskiy2009non}.
+while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways (up to C3a) as a system of non-linear ODEs \cite{korotaevskiy2009non}.
 More recently, large mechanistic models of sections of complement have also been proposed.
@@
 analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
-Recently, Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated
-the complement response of the host and pathogen \cite{zewde2016quantitative}.
+Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated between
+the host cell and pathogen \cite{zewde2016quantitative}.
 However, these previous modeling studies involved little experimental validation.
-Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict experimentally validated complement dynamics.
-The central challenge is the estimation of model parameters from experimental data.
-Unlike other important cascades, such as coagulation for which there are well developed experimental tools and many publicly available data sets,
-the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of mechanistic complement models difficult.
+Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict complement measurements.
+The central challenge of complement model identification is the estimation of model parameters from experimental measurements.
+Unlike other important cascades, such as coagulation where there are well developed experimental tools and publicly available data sets,
+the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of large mechanistic complement models difficult.
+Thus, reduced order modeling approaches, which describe the key biology with a limited number of parameters, will be important for developing validated models of complement, and to fully understandings its role in the body.
 
-In this study, we developed an ensemble of experimentally validated reduced order complement models.
-The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
-The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
-Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
-We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
-Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
-Despite its small size, the model was surprisingly predictive.
-After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
-controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
-The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
-Taken together, we developed a reduced order complement model that was computationally inexpensive,
-and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
-The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
 %Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
@@
 \section*{Results}
+
+In this study, we developed an ensemble of experimentally validated reduced order complement models.
+The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
+The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
+Thus, the model was an order of magnitude smaller and included more pathways than comparable models in the literature.
+We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
+Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
+Despite its small size, the model was surprisingly predictive.
+After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
+controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention;
+only the knockdown of both C3 and C5 consistently reduced C3a and C5a formation for all cases.
+Taken together, we developed a reduced order complement model that was computationally inexpensive,
+and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
+The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
 
@@
 However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
-This agreed with the sensitivity results; changes in AP C3-convertase formation or activity affected the downstream dynamics of C5a formation.
+This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
 Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig.

### 2016-06-13 Paper_v4.tex
New figs

@@
 \begin{equation}
-	\frac{dx_{i}}{dt}  =  \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
+	\frac{1}{\tau_{i}}\frac{dx_{i}}{dt}  =  \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
 \end{equation}
 where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of protein species in the model.
+The quantity $\tau_{i}$ denotes a timescale parameter for species $i$ which captures unmodeled effects; by default $\tau_{i}$ = 1, for all species unless otherwise specified.
 The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final_v2.pdf}
 \caption{Reduced order complement model training.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final_v2.pdf}
 \caption{Reduced order complement model predictions. The reduced order coagulation model parameter estimates were tested against data not used during model training \cite{morad2015time}.
@@
 \centering
-\includegraphics[width=1.1\textwidth]{./figs/Figure4_SEN_v2.pdf}
+\includegraphics[width=1.1\textwidth]{./figs/Figure4_Sensitivity_v3.pdf}
 \caption{Global sensitivity analysis of the reduced order complement model.
-Sensitivity analysis was conducted on the four cases used for model training.
-\textbf{A}: Sensitivity of the C3a residual at w/o zymosan,
-\textbf{B}: Sensitivity of the C5a residual at w/o zymosan,
-\textbf{C}: Sensitivity of the C3a residual at 1g zymosan, and
-\textbf{D}: Sensitivity of the C5a residual at 1g zymosan.
+Sensitivity analysis was conducted on the two objectives used for model training.
+\textbf{A}: Sensitivity of the C3a and C5a residual w/o zymosan,
+\textbf{B}: Sensitivity of the C3a and C5a residual with 1mg/ml zymosan.
 The bars denote the mean total sensitivity index for each parameter, while the error bars denote the 95\% confidence interval.
-\textbf{E}: Pathways controlled by the sensitivity parameters.
+\textbf{C}: Pathways controlled by the sensitivity parameters.
 Bold black lines indicate the pathway involves one or more sensitive parameters, while the red lines show current therapeutics targets.

### 2016-06-13 Paper_v4.tex
Man, we are close ...

@@
 Six validation cases were considered, three for C3a and C5a, respectively.
-All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan).
-The ensemble of reduced order models predicted the qualitative dynamics of C3a formation (Fig. \ref{fig-prediction}, left column),
-and C5a formation (Fig. \ref{fig-prediction}, right column) at three inducer concentrations.
-However, there were shortcomings with model performance, especially for the C3a prediction.
-First, while the overall C3a trend was captured, the level of C3a was consistently under-predicted (outside of our 99\% confidence interval) in all cases (Fig. \ref{fig-prediction}, left column).
-We believe the C3a under-prediction was related to our choice of training data, and how we modeled C4BP interactions.
-We trained the model using 1g zymosan, but predicted cases with much less initiator.
-Next, the C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes.
-However, the binding of C4BP with complement proteins is likely reversible, and CP C3 convertase may have residual activity even in the bound form.
-Thus, the model likely over-predicted the influence of C4BP.
-Second, while the C5a measurements were within the 99\% confidence estimate, we failed to capture the concave down curvature for the 0.001g and 0.01g zymosan cases.
-The decreasing slope of the C5a may indicate cofactor limitation, or missing biology which we have not accounted for in the reduced order approach.
-Despite these shortcomings, we qualitatively predicted unseen experimental data, including correctly capturing the time scale of C3a formation,
-and the correct order of magnitude for C5a for three inducer levels.
+All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan, and other experimentally mandated changes).
+The ensemble of reduced order models predicted the qualitative dynamics of C3a formation (Fig. \ref{fig-prediction}, top),
+and C5a formation (Fig. \ref{fig-prediction}, bottom) at three inducer concentrations.
+The rate of C3a formation and C3a peak time were directly proportional to initiator dose.
+Similarly, the C5a plateau and rate of formation were also directly proportional to initiator dose, with the lag time being indirectly proportional to initiator exposure.
+However, there were shortcomings with model performance.
+First, while the overall C3a trend was captured (within the 99\% confidence interval), the C3a dynamics were too fast with the exception of the low dose case.
+We believe the C3a timescale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
+We trained the model using either no or 1 mg/ml zymosan, but predicted cases in a different initiator range;
+comparing the training to prediction, the model performance e.g., the shape of the C3a trajectory was biased towards either high or very low initiator doses.
+Next, tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
+Thus, we skipped an important step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
+We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
+The C5a predictions followed a similar trend as C3a; we captured the long-time C5a behavior but over predicted the time scale of C5 cleavage.
+However, because the C5a time scale depends strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
+Despite these shortcomings, we qualitatively predicted unseen experimental measurements within the 99\% confidence of the ensemble, for three inducer levels.
 Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.
@@
 We conducted sensitivity analysis to estimate which parameters controlled the performance of the reduced order complement model.
-We calculated the sensitivity of the C3a and C5a residuals to changes in model parameters with and without zymosan for the ensemble of parameter sets (Fig. \ref{fig-SA}A - D).
-In the absence of zymosan (where only the alternative pathway is active), $k_{f,C3b}$ (tickover formation of C3b) and $k_{d,C3a}$ (rate constant governing C3a degradation)
-controlled the C3a fit (Fig.
-\ref{fig-SA}A).
-Interestingly, neither $k_{c,C3C,AP}$ (the rate constant governing AP C3-convertase activity) nor $k_{f,C3C,AP}$ (the rate constant governing AP C3-convertase formation) were sensitive in the absence of zymosan.
+We calculated the total sensitivity of the C3a and C5a residual to changes in model parameters with and without zymosan (Fig. \ref{fig-SA}).
+In the absence of zymosan (where only the alternative pathway is active), the most sensitive parameter was the rate constant governing the assembly of the AP C3 convertase, as
+well as the rate constant controlling basal C3b formation. The C5a trajectory was sensitive to the AP C5 convertase kinetic parameters (Fig. \ref{fig-SA}A).
+Interestingly, neither the rate nor the saturation constants governing AP C3-convertase activity were sensitive in the absence of zymosan.
 Thus, C3a formation in the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity, in the absence of zymosan.
-On the other hand, the C5a residual was strongly sensitive to tickover parameters e.g., $k_{f,C3b}$, and to a lesser extent $k_{f,C3C,AP}$ as well as
-parameters controlling AP C5 convertase activity in the absence of zymosan (Fig. \ref{fig-SA}B).
-Thus, AP C3-convertase did not influence C3a dynamics, but strongly effected C5a formation in the absence of zymosan.
-In the presence of zymosan, the C3a residual was controlled by the formation and activity of the CP C3 convertase, as well as tickover and degradation parameters (Fig. \ref{fig-SA}C).
-Further, the C5a residual was influenced by lectin initiation parameters, the formation and activity of CP C5 convertase, and the inhibition by C4BP in the presence of zymosan (Fig. \ref{fig-SA}D).
-Globally, sensitivity analysis suggested that lectin pathway parameters dominated system performance in the presence of zymosan, and that C5a formation was sensitive to both
-C3 and C5 convertase activity, while C3a was robust to the AP C3 convertase.
-In the absence of zymosan, the tickover parameters were important to both C3a and C5a formation, but C5a was also sensitive to both AP C3 and C5 convertase formation.
-We compared the sensitivity results to current therapeutic approaches; pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}E).
-Thus, there was a qualitative overlap between sensitivity and the potential of biochemical efficacy.
-However, sensitivity coefficients are only a local measure of how small changes in a parameters affect model performance.
+In the presence of zymosan, the C3a residual was controlled by the formation and activity of the CP C3 convertase, as well as tickover and degradation parameters.
+On the other hand, the C5a residual was controlled by the formation and activity of CP C5 convertase, and tickover C3b formation in the presence of zymosan (Fig. \ref{fig-SA}B).
+The lectin initiation parameters were sensitive, but to a lesser extent than CP convertase kinetic parameters and tickover C3b formation.
+Thus, sensitivity analysis suggested that CP C3/C5 convertase formation and activity dominated in the presence of zymosan, but tickover parameters and AP C5 convertase were more important
+without initiator. Next, we compared the sensitivity results to current therapeutic approaches;
+pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}C).
+In particular, the sensitivity analysis suggested AP/CP C5 convertase inhibitors, or interventions aimed at attenuating C3 or C5 would most strongly influence complement performance.
+Thus, there was at least a qualitative overlap between sensitivity and the potential of biochemical efficacy.
+However, sensitivity coefficients are only a local measure of how small changes in parameters affect model performance.
 To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
@@
 Though knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate \cite{sissons1977metabolism, swaak1982determination} of these proteins make them difficult targets.
-Thus targeting C3 and C5 directly will require high dosage of drugs. It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS),
-a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
+Thus targeting C3 and C5 directly will require high dosage of drugs.
+It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
 The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
-Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a. Formation of convertases is inhibited by targeting
-upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
+Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
+Formation of convertases is inhibited by targeting upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
 For example, Omeros is a protease inhibitor that targets MASP-2 complex and thereby inhibits formation of downstream convertases \cite{schwaeble2011methods}.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final_v3.pdf}
 \caption{Reduced order complement model predictions. The reduced order coagulation model parameter estimates were tested against data not used during model training \cite{morad2015time}.

### 2016-06-13 Paper_v4.tex
working on updated MM

@@
 In this study, we used $\mathcal{I}_{j}\in\left\{min,max\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
-The model equations were implemented in MATLAB and solved using the \texttt{ODE23s} routine (The Mathworks, Natick MA).
+The model equations were implemented in Julia and solved using the \texttt{SUNDIALS} routine [REFs].
 The complement model code and parameter ensemble is freely available under an MIT software license and can be downloaded from http://www.varnerlab.org.
@@
 \subsection*{Estimation of an ensemble of complement model parameters.}
-We minimized the residual between simulations and experimental C3a and C5a measurements using Dynamic Optimization with Particle Swarms (DOPS). DOPS minimized the objective:
-\begin{equation}\label{eqn:objective-function}
-	\min_{\mathbf{k}} \sum_{\tau=1}^{\mathcal{T}}\sum_{j=1}^{\mathcal{S}}\left(\frac{\hat{x}_{j}\left(\tau\right) - x_{j}\left(\tau,\mathbf{k}\right)}{\omega_{j}\left(\tau\right)}\right)^{2}
-\end{equation}where $\hat{x}_{j}\left(\tau\right)$ denotes the measured value of species $j$ at time $\tau$, $x_{j}\left(\tau,\mathbf{k}\right)$ denotes the simulated
-value for species $j$ at time $\tau$, and $\omega_{j}\left(\tau\right)$ denotes the experimental measurement variance for species $j$ at time $\tau$. The outer summation is with respect to
-time, while the inner summation is with respect to state.
+We estimated a single initial parameter set using the using Dynamic Optimization with Particle Swarms (DOPS) technique [REF].
+DOPS minimized the squared residual between simulated and experimental C3a and C5a measurements with and without zymosan as a single objective formulation.
+This initial best fit set then served as the starting point for multiobjective ensemble generation using Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs) [REF].
+JuPOETs is a multiobjective approach which integrates simulated annealing with Pareto optimality to estimate model ensembles on or near the optimal tradeoff surface between competing training objectives.
+JuPOETs minimized training objective functions of the form:
+\begin{equation}\label{eqn_cost2}
+	E_{j}(\mathbf{k}) = \sum_{i=1}^{\mathcal{T}_{j}}\biggl(\hat{\mathcal{M}}_{ij}-\hat{y}_{ij}(\mathbf{k})\biggr)^2 + \left(\frac{{\mathcal{M}^{\prime}_{ij}}-\max{y_{ij}}}{{\mathcal{M}^{\prime}_{ij}}}\right)^{2}
+\end{equation}
+The symbol $\hat{\mathcal{M}}_{ij}$ denotes an experimental observation (from training set $j$) while the symbol $\hat{y}_{ij}$ denotes the simulation output (from training objective $j$).
+The quantity $i$ denotes the sampled time-index and $\mathcal{T}_{j}$ denotes the number of time points for experiment $j$.
+In this study, we minimized two training objectives, the sum of the C3a and C5a residual w/o zymosan (objective )
+
+% \subsubsection*{Dynamic Optimization with Particle Swarms (DOPS).}
+% DOPS minimized the residual between simulations and experimental C3a and C5a measurements using Dynamic Optimization with Particle Swarms (DOPS).
+%
+
+% DOPS minimized the objective:
+% \begin{equation}\label{eqn:objective-function}
+% 	\min_{\mathbf{k}} \sum_{\tau=1}^{\mathcal{T}}\sum_{j=1}^{\mathcal{S}}\left(\frac{\hat{x}_{j}\left(\tau\right) - x_{j}\left(\tau,\mathbf{k}\right)}{\omega_{j}\left(\tau\right)}\right)^{2}
+% \end{equation}where $\hat{x}_{j}\left(\tau\right)$ denotes the measured value of species $j$ at time $\tau$, $x_{j}\left(\tau,\mathbf{k}\right)$ denotes the simulated
+% value for species $j$ at time $\tau$, and $\omega_{j}\left(\tau\right)$ denotes the experimental measurement variance for species $j$ at time $\tau$. The outer summation is with respect to
+% time, while the inner summation is with respect to state.
 DOPS is a novel metaheuristic that combines multi swarm particle swarm optimization (PSO) with a greedy global optimization algorithm called dynamically dimensioned search (DDS).
-DOPS is faster than conventional global optimizers and has the ability to find near optimal solutions for high dimensional systems within a relatively few function evaluations.
-It uses an adaptive switching strategy based on error convergence rates to switch from the particle swarm to DDS search phases.
-This enables DOPS to quickly estimate globally optimal or near optimal solutions even in the presence of many local minima.
-In the swarm search, for each iteration the particles compute error within each sub-swarm by evaluating the model equations using their specific parameter vector realization.
-From each of these points within a sub-swarm a local best is identified.
+% DOPS is faster than conventional global optimizers and has the ability to find near optimal solutions for high dimensional systems within a relatively few function evaluations.
+% It uses an adaptive switching strategy based on error convergence rates to switch from the particle swarm to DDS search phases.
+% This enables DOPS to quickly estimate globally optimal or near optimal solutions even in the presence of many local minima.
+% In the swarm search, for each iteration the particles compute error within each sub-swarm by evaluating the model equations using their specific parameter vector realization.
+% From each of these points within a sub-swarm a local best is identified.
 This along with the particle best within the sub-swarm $\mathcal{S}_{k}$ is used to update the parameter estimate for each particle using the following rules:
-\begin{equation}\label{eqn:update-rule}
-	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
-	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
-\end{equation}
-where ${z}_{i,j}$ is the parameter vector, $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm
-$\mathcal{S}_{k}$ for function evaluations $1\rightarrow j-1$, and
-$\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
-The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
-At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was used to initialize the DDS phase.
-For the DDS phase, the best parameter estimate was updated using the rule:
-\begin{equation}
-  \mathcal{G}_{new}({J})=\begin{cases}
-    \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
-    \mathcal{G}(\mathbf{J}), & \text{otherwise}.
-  \end{cases}
-\end{equation}
-where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
-and $\sigma$ denotes the perturbation amplitude:
-\begin{equation}
-	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
-\end{equation}
-where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
+% \begin{equation}\label{eqn:update-rule}
+% 	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+% 	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
+% \end{equation}
+% where ${z}_{i,j}$ is the parameter vector, $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm
+% $\mathcal{S}_{k}$ for function evaluations $1\rightarrow j-1$, and
+% $\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
+% The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
+% At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was used to initialize the DDS phase.
+% For the DDS phase, the best parameter estimate was updated using the rule:
+% \begin{equation}
+%   \mathcal{G}_{new}({J})=\begin{cases}
+%     \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
+%     \mathcal{G}(\mathbf{J}), & \text{otherwise}.
+%   \end{cases}
+% \end{equation}
+% where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
+% and $\sigma$ denotes the perturbation amplitude:
+% \begin{equation}
+% 	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
+% \end{equation}
+% where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
 The set $\mathbf{J}$ was constructed using a monotonically decreasing probability function $\mathcal{P}_{i}$ that represents a threshold for determining whether a specific dimension $j$ was perturbed or not.
 DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
-At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ which we use for plotting best fits against the experimental data, and for generating a parameter ensemble.
+% At the end of DDS phase we obtain the optimal vector  $\mathcal{G}$ which we use for plotting best fits against the experimental data, and for generating a parameter ensemble.
 
-An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by analyzing repeated DOPS runs.
-Parameters were selected  using goodness of fit (model residual) and numerical stability criteria.
-We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
-The best fit parameter set, and parameter ensemble is available in a tab delimited plain text format from the Complement model GitHub repository.
-The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
+% \subsubsection*{Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).}
+% An ensemble of parameter sets was estimated using JuPOETs using .
+%
+%
+% Two objective functions were used to train the
+%
+% An ensemble of parameters was obtained by randomly perturbing the optimal parameter set within bounds established by analyzing repeated DOPS runs.
+% Parameters were selected  using goodness of fit (model residual) and numerical stability criteria.
+% We selected an ensemble of N = 50 parameter sets for this study using this sampling procedure.
+% The best fit parameter set, and parameter ensemble is available in a tab delimited plain text format from the Complement model GitHub repository.
+% The DOPS routine was implemented in MATLAB (The Mathworks, Natick MA) and can be downloaded from http://www.varnerlab.org under an MIT software license.
 
@@
 We conducted global sensitivity and robustness analysis to estimate which parameters and species controlled the performance of the reduced order model.
-We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases using the Sobol method \cite{SOBOL_METHOD}.
-The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
+We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a/C5a alternate and C3a/C5a lectin objectives
+using the Sobol method \cite{SOBOL_METHOD}. The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
 We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
 where $N$ was the number of trials per parameters, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
-In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed using 11,600 model evaluations.
+In our case, $N$ = 400 and $d$ = 28, so the total sensitivity indices were computed using 23,200 model evaluations.
 The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.

### 2016-06-14 Paper_v4.tex
Condensed treatment paragraph in discussion

@@
 The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al which involved over 100 ODEs \cite{zewde2016quantitative}.
-However, we found that the quantity of C3a produced in the alternate pathway was nearly 1000 times the quantity of C5a produced.
-Though this was in agreement with the experimental data \cite{morad2015time},
-it differed from the theoretical predictions made by Zewde et al. who showed C3a was ${10^8}$ times the C5a concentration \cite{zewde2016quantitative}.
-In our model, the time profile of C5a generation from the lectin pathway changed with respect to the quantity of zymosan (the lectin pathway initiator).
-The lag phase for generation was inversely proportional to the initiator concentration.
+However, we found that the quantity of C3a produced in the alternate pathway was nearly 1000 times the amount of C5a generated.
+While this was in agreement with the experimental data \cite{morad2015time},
+it differed from the theoretical predictions made by Zewde et al. who showed C3a was ${10^8}$ times greater than the C5a concentration \cite{zewde2016quantitative}.
+In our model, the time profile of both C3a and C5a generated changed with respect to the quantity of zymosan (the lectin pathway initiator).
+In particular, the C3a peak time was directly proportional to initiator, while the lag phase for generation was inversely proportional to the initiator concentration.
 Korotaevskiy et al. showed a similar trend using a theoretical model of complement, albeit for much shorter time scales \cite{korotaevskiy2009non}.
-Thus, the reduced order complement model performed similarly to existing large mechanistic models, despite being significantly smaller.
+Thus, the reduced order complement model performed at least as well as existing larger mechanistic models, despite being significantly smaller.
 
-Global analysis of the complement model estimated potential important therapeutic targets.
-Complement malfunctions are implicated in a number of diseases, however the
+Global analysis of the complement model suggested potentially important therapeutic targets.
+Complement malfunctions are implicated in a spectrum of diseases, however the
 development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
@@
 In this study, we analyzed a validated ensemble of reduced order complement models to estimate therapeutically important mechanisms.
-In presence of an initiator, C5a formation was primarily sensitive to the lectin initiation parameters,
-and parameters governing the conversion of C5 to C5a and C5b.
-This result agrees well with the current protease inhibitors targeting initiating complexes, including mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{heja2012monospecific}.
-The most commonly used anti-complement drug eculizumab \cite{morgan2015complement}, targets the C5 protein which is cleaved to form C5a.
-Our sensitivity analysis showed that kinetic parameters governing C5 conversion were sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
-The formation of basal C3b was also a sensitive parameter in the formation of C3a through the alternate pathway.
-Thus, this mechanism can act as a target for both C3a and C5a inhibitors.
-Lectin initiated C3a formation showed a number of sensitive parameters.
-This included the lectin initiation parameters that controlled C5a formation, C3 convertase inhibition by C4BP, and parameters governing C3 convertase activity.
-All these mechanisms are potential drug targets.
-
-To further validate these results from sensitivity analysis about potential drug targets we did a robustness analysis.
-We knocked down C3 and C5 levels and studied their impact on the generation of C3a and C5a.
-The C3a and C5a levels in the lectin pathway were strongly influenced by initial levels of C3 and C5.
-Thus direct inhibition of C3 and C5, or targeting complexes (MASP complex, C3 and C5 convertases) that act on C3 and C5 have a direct impact on production of C3a and C5a.
-This is also in agreement with sensitivity analysis that C5 is a good drug target. A number of drugs targeting C5 are being developed.
-For example LFG316 by Novartis is being used to target C5 in cases of Age-Related Macular Degeneration \cite{roguska2014generation},
-Mubodina is an antibody that targets C5 in the treatment of Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement},
-Coversin is a small molecule targeting C5 \cite{weston2014clinical}, Zimura is an aptamer targeting C5 \cite{epstein2007complement},
-small peptides and RNAi are also being used to inhibit C5 \cite{borodovsky2014aln}.
-Another important conclusion that can be drawn together from sensitivity and robustness analysis is that C3 and C5 convertases can be important therapeutic targets.
-Though knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate \cite{sissons1977metabolism, swaak1982determination} of these proteins make them difficult targets.
-Thus targeting C3 and C5 directly will require high dosage of drugs.
+In the presence of an initiator, C3a and C5a formation was primarily sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
+Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
+For example, Omeros, a protease inhibitor that targets the MASP-2 complex, has been shown to inhibit the formation of downstream convertases \cite{schwaeble2011methods}.
+Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
+Removing supporting molecules could destabilize the convertases.
+For example, Novelmed Therapeutics recently developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
+On the other hand, robustness analysis suggested C3a and C5a generation could be significantly attenuated by modulating the free levels of C3 and C5.
+The most commonly used anti-complement drug eculizumab, targets the C5 protein which is cleaved to form C5a \cite{morgan2015complement}.
+A number of other drugs targeting C5 are also being developed. For example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
+while Mubodina, an antibody that targets C5, is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
+Other agents such as the small molecule Coversin \cite{weston2014clinical}, the aptamer Zimura \cite{epstein2007complement},
+as well as small peptides and RNAi could also be used to inhibit C5 \cite{borodovsky2014aln}.
+However, knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
+Thus, targeting C3 and C5 directly will require high dosage.
 It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
 The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
-Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
-Formation of convertases is inhibited by targeting upstream protease complexes like MASP-1,2 from lectin pathway (or C1r, C1s from classical pathway).
-For example, Omeros is a protease inhibitor that targets MASP-2 complex and thereby inhibits formation of downstream convertases \cite{schwaeble2011methods}.
-Lampalizumab (an immunoglobulin) and Bikaciomab (an antibody fragment) target Factor B and Factor D respectively.
-Factor B and Factor D are crucial to formation alternate pathway convertases \cite{katschke2012inhibiting, hu2013therapeutic}.
-Novelmed Therapeutics recently developed antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
-Cobra Venom Factor (CVF), an analogue of C3b has been used to bind to Factor B to regulate alternate convertases \cite{vogel2004recombinant}.
-Thus, analysis of the ensemble of complement models identified potentially important therapeutic targets that are consistent with therapeutic strategies that are under development.
 
+% and parameters governing the conversion of C5 to C5a and C5b.
+% This result agrees well with the current protease inhibitors targeting initiating complexes, including mannose-associated serine proteases 1 and  2 (MASP-1,2) \cite{heja2012monospecific}.
+% Our sensitivity analysis showed that kinetic parameters governing C5 conversion were sensitive in both lectin initiated and alternate pathways, thus agreeing with targeting C5 protein.
+% The formation of basal C3b was also a sensitive parameter in the formation of C3a through the alternate pathway.
+% Thus, this mechanism can act as a target for both C3a and C5a inhibitors.
+% Lectin initiated C3a formation showed a number of sensitive parameters.
+% This included the lectin initiation parameters that controlled C5a formation, C3 convertase inhibition by C4BP, and parameters governing C3 convertase activity.
+% All these mechanisms are potential drug targets.
+% To further validate these results from sensitivity analysis about potential drug targets we did a robustness analysis.
+% We knocked down C3 and C5 levels and studied their impact on the generation of C3a and C5a.
+% Thus direct inhibition of C3 and C5, or targeting complexes (MASP complex, C3 and C5 convertases) that act on C3 and C5 have a direct impact on production of C3a and C5a.
+% This is also in agreement with sensitivity analysis that C5 is a good drug target.
+% Another important conclusion that can be drawn together from sensitivity and robustness analysis is that C3 and C5 convertases can be important therapeutic targets.
+% Though Our analysis shows that formation and assembly of these convertases are sensitive mechanisms that strongly impact downstream proteins like C5a.
+% Thus, analysis of the ensemble of complement models identified potentially important therapeutic targets that are consistent with therapeutic strategies that are under development.
 %Indeed a number of next generation drugs are targeting convertases directly \cite{}, inhibiting their assembly or formation \cite{} or targeting the individual convertase components \cite{}
-
 %Without initiator, C5a formation was the most sensitive to the formation of basal C3b through the tick over mechanism.
@@
 %. Currently there are only four approved drugs targeting complement \cite{morgan2015complement}. Eculizumab targets C5, Berinert, Ruconest and Cinryze target C1.
-
 % complement in immune response has been well known since long, there
@@
 In this study, we used $\mathcal{I}_{j}\in\left\{min,max\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
-The model equations were implemented in Julia and solved using the \texttt{SUNDIALS} routine [REFs].
-The complement model code and parameter ensemble is freely available under an MIT software license and can be downloaded from http://www.varnerlab.org.
+The model equations were implemented in Julia and solved using the \texttt{CVODE} routine of the Sundials package \cite{Julia,Sundials}.
+The model code and parameter ensemble is freely available under an MIT software license and can be downloaded from http://www.varnerlab.org.
 
-\subsection*{Estimation of an ensemble of complement model parameters.}
-We estimated a single initial parameter set using the using Dynamic Optimization with Particle Swarms (DOPS) technique [REF].
-DOPS minimized the squared residual between simulated and experimental C3a and C5a measurements with and without zymosan as a single objective formulation.
-This initial best fit set then served as the starting point for multiobjective ensemble generation using Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs) [REF].
+\subsection*{Estimating an ensemble of complement model parameters.}
+We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique [REF].
+DOPS is a novel hybrid meta-heuristic which combines a multi-swarm particle swarm method with the dynamically dimensioned search approach of Shoemaker and colleagues \cite{Tolson2007}.
+DOPS minimized the squared residual between simulated and C3a and C5a measurements with and without zymosan as a single objective.
+The best fit set estimated by DOPS served as the starting point for multiobjective ensemble generation using Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs) \cite{JuPOETs-BioArXiv}.
 JuPOETs is a multiobjective approach which integrates simulated annealing with Pareto optimality to estimate model ensembles on or near the optimal tradeoff surface between competing training objectives.
-JuPOETs minimized training objective functions of the form:
+JuPOETs minimized training objectives of the form:
 \begin{equation}\label{eqn_cost2}
-	E_{j}(\mathbf{k}) = \sum_{i=1}^{\mathcal{T}_{j}}\biggl(\hat{\mathcal{M}}_{ij}-\hat{y}_{ij}(\mathbf{k})\biggr)^2 + \left(\frac{{\mathcal{M}^{\prime}_{ij}}-\max{y_{ij}}}{{\mathcal{M}^{\prime}_{ij}}}\right)^{2}
+	O_{j}(\mathbf{k}) = \sum_{i=1}^{\mathcal{T}_{j}}\biggl(\hat{\mathcal{M}}_{ij}-\hat{y}_{ij}(\mathbf{k})\biggr)^2 + \left(\frac{{\mathcal{M}^{\prime}_{ij}}-\max{y_{ij}}}{{\mathcal{M}^{\prime}_{ij}}}\right)^{2}
+\end{equation}subject to the model equations, initial conditions and parameter bounds $\mathcal{L}\leq\mathbf{k}\leq\mathcal{U}$.
+The first term in the objective function measured the shape difference between the simulations and measurements.
+The symbol $\hat{\mathcal{M}}_{ij}$ denotes a scaled experimental observation (from training set $j$) while the symbol $\hat{y}_{ij}$ denotes the scaled simulation output (from training set $j$).
+The quantity $i$ denotes the sampled time-index and $\mathcal{T}_{j}$ denotes the number of time points for experiment $j$.
+The scaled measurement is given by:
+\begin{equation}\label{norm_exp_data}
+\hat{\mathcal{M}}_{ij} = \frac{\mathcal{M}_{ij} - \min_{i}\mathcal{M}_{ij}}{\max_{i}{\mathcal{M}_{ij}}-\min_{i}{\mathcal{M}_{ij}}}
 \end{equation}
-The symbol $\hat{\mathcal{M}}_{ij}$ denotes an experimental observation (from training set $j$) while the symbol $\hat{y}_{ij}$ denotes the simulation output (from training objective $j$).
-The quantity $i$ denotes the sampled time-index and $\mathcal{T}_{j}$ denotes the number of time points for experiment $j$.
-In this study, we minimized two training objectives, the sum of the C3a and C5a residual w/o zymosan (objective )
+Under this scaling, the lowest measured concentration become zero while the highest equaled one, where a similar scaling was defined for the simulation output.
+The second-term in the objective function quantified the absolute error in the estimated concentration scale, where the absolute measured
+concentration (denoted by $\mathcal{M}^{\prime}_{ij}$) was compared with the largest simulated value.
+In this study, we minimized two training objectives, the total C3a and C5a residual w/o zymosan (O$_{1}$) and the total C3a and C5a residual for 1 mg/ml zymosan (O$_{2}$).
+JuPOETs identified an ensemble of N $\simeq$ 2100 parameter sets which were used for model simulations and uncertainty quantification subsequently.
+JuPOETs is open source, available under an MIT software license.
+The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+The objective functions used in this study are available in the GitHub model repository available from http://varnerlab.org.
 

### 2016-06-14 Paper_v4.tex
Updates to discussion

@@
 development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
-Previously, we have shown that mathematical modeling and sensitivity analysis can be useful tools to estimate therapeutically important mechanisms in biochemical networks \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
-In this study, we analyzed a validated ensemble of reduced order complement models to estimate therapeutically important mechanisms.
-In the presence of an initiator, C3a and C5a formation was primarily sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
+Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
+In this study, we analyzed a validated ensemble of reduced order complement models to better understand the strengths and weaknesses of the cascade.
+In the presence of an initiator, C3a and C5a formation was sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
 Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
@@
 Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
-Removing supporting molecules could destabilize the convertases.
-For example, Novelmed Therapeutics recently developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
+Removing supporting molecules could also destabilize the convertases.
+For example, Novelmed Therapeutics developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
+Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
 On the other hand, robustness analysis suggested C3a and C5a generation could be significantly attenuated by modulating the free levels of C3 and C5.
-The most commonly used anti-complement drug eculizumab, targets the C5 protein which is cleaved to form C5a \cite{morgan2015complement}.
+The most commonly used anti-complement drug eculizumab, targets the C5 protein \cite{morgan2015complement}.
 A number of other drugs targeting C5 are also being developed. For example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
@@
 as well as small peptides and RNAi could also be used to inhibit C5 \cite{borodovsky2014aln}.
-However, knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
-Thus, targeting C3 and C5 directly will require high dosage.
-It is also well known that eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease that is caused due to uncontrolled complement activation \cite{noris2014dynamics}.
-The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
+The cyclic peptide C3 inhibitor Compstatin and its derivatives [FINISH] \cite{Mastellos:2015aa}.
+However, while the knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate of these proteins make them difficult to target \cite{sissons1977metabolism, swaak1982determination}.
+For example, the eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease involving uncontrolled complement activation \cite{noris2014dynamics}.
 
+%The issue of high dosage can potentially be circumvented by targeting convertases or fragile mechanisms that involve C3, C5 or their activated components.
 % and parameters governing the conversion of C5 to C5a and C5b.

### 2016-06-14 Paper_v4.tex
Updating

@@
 Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
+However, despite its importance, there are very few approved complement specific therapeutics.
 
-The complement cascade involves over 30 soluble and cell surface proteins, receptors and regulators.
+The complement cascade involves many soluble and cell surface proteins, receptors and regulators.
 The molecular connectivity of complement involves both fluid phase and cell surface events, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
@@
 the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of large mechanistic complement models difficult.
-Thus, reduced order modeling approaches, which describe the key biology with a limited number of parameters, will be important for developing validated models of complement, and to fully understandings its role in the body.
+Thus, reduced order modeling approaches, which describe the key biology with a limited number of parameters, will be important for developing validated models of complement, and to fully understanding its role in the body.
 

### 2016-06-15 Paper_v4.tex
Completed text

@@
 Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
-Only dual dual-knockdown of both C3 and C5 consistently reduced C5a formation when all pathways were activated.
+Only dual dual-knockdown of C3 and C5 consistently reduced C3 and C5a formation when all pathways were activated.
 Taken together, we developed a reduced order complement model that was computationally inexpensive,
@@
 Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
-However, research over the past decade has shown that the importance of complement extends well beyond innate immunity.
-For example, complement contributes to tissue homeostasis by inducing tissue repair \cite{ricklin2010complement}.
-Complement has also been linked with several diseases including Alzheimers, Parkinson's disease, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
-Complement plays both positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
-Lastly, several other important biochemical subsystems are integrated with complement including the coagulation cascade, the autonomous nervous system and inflammation \cite{ricklin2013complement}.
+However, research over the past decade has shown the importance of complement extends well beyond innate immunity.
+For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}, and has also been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
+Lastly, several other important biochemical systems are integrated with complement including the coagulation cascade, the autonomous nervous system and inflammation \cite{ricklin2013complement}.
 Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
-However, despite its importance, there are very few approved complement specific therapeutics.
+However, despite its importance, there have been relatively few approved complement specific therapeutics.
 
@@
 the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of large mechanistic complement models difficult.
-Thus, reduced order modeling approaches, which describe the key biology with a limited number of parameters, will be important for developing validated models of complement, and to fully understanding its role in the body.
+Thus, reduced order modeling approaches, which describe the key biology of complement using a limited number of parameters, will be important for developing validated models of complement, and to fully understanding its role in the body.
 
@@
 Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
-On the other hand, robustness analysis suggested C3a and C5a generation could be significantly attenuated by modulating the free levels of C3 and C5.
+On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
 The most commonly used anti-complement drug eculizumab, targets the C5 protein \cite{morgan2015complement}.
@@
 as well as small peptides and RNAi could also be used to inhibit C5 \cite{borodovsky2014aln}.
-The cyclic peptide C3 inhibitor Compstatin and its derivatives [FINISH] \cite{Mastellos:2015aa}.
-However, while the knockdown of C3 and C5 affects C3a and C5a levels downstream, the abundance and turnover rate of these proteins make them difficult to target \cite{sissons1977metabolism, swaak1982determination}.
-For example, the eculizumab dosage needs to be adjusted while treating for Atypical Hemolytic-Uremic Syndrome (aHUS), a disease involving uncontrolled complement activation \cite{noris2014dynamics}.
+The cyclic peptide inhibitor Compstatin and its derivatives, which target C3, are promising approaches for the treatment of X, Y and Z [FINISH] \cite{Mastellos:2015aa}.
+However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
+For example, the eculizumab dosage must be significantly adjusted during the course of treatment for aHUS \cite{noris2014dynamics}.
+A validated complement model, in combination with personalized pharmacokinetic models of immune system function, could be an important development for the field.
 
@@
 A logical progression for this work would be to expand the network to include the classical pathway and the formation of the membrane attack complex (MAC).
-However, it is unclear whether the addition of the classical pathway will decrease the predictive quality of our existing model.
-Liu et al have shown cross-talk between the activation of the classical and lectin pathways that could influence model performance \cite{liu2011computational}.
-One potential approach to address such difficulties would be to incorporate C reactive proteins (CRP) and L-ficolin (LF) into the model,
-both of which are involved with the initiation of classical and lectin pathways.
-Liu et al.
-showed that under inflammation conditions interactions between lectin and classical pathways was mediated through CRP and LF  \cite{liu2011computational}.
-Thus incorporating these two proteins would help us in modeling cross talk.
-Time course measurements of MAC abundance (and MAC formation dynamics) are also scarce, making the inclusion of MAC challenging.
-Next, we should address the under-prediction of C3a. We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
-C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes;
-however, the binding of C4BP with complement proteins is likely reversible and C4BP-bound convertases may have residual activity.
-We also did not capture the maximum concentration of C3a at low initiator levels.
-One possible reasons for this could the C2-by pass pathway, which was not included in the model.
-This pathway further accelerates C3a production without the involvement of a C3 convertase.
-Currently the C3a in the model is generated only through the activity of a C3 convertase.
-Incorporating this additional step within the reduced order modeling framework would be a future direction that we need to consider.
-We should test alternative model structures which include reversible C4BP binding, and partially active convertases.
-Alternatively, we could also perform sensitivity analysis on the C3a prediction residual to determine which parameters controlled the C3a prediction.
+However, time course measurements of MAC abundance (and MAC formation dynamics) are scarce, making the inclusion of MAC challenging.
+On the other hand, inclusion of classical pathway activation is straightforward.
+Liu et al., have shown cross-talk between the activation of the classical and lectin pathways through C reactive proteins (CRP) and L-ficolin (LF) under inflammation conditions \cite{liu2011computational}.
+Thus, inclusion of these species, in addition to a lumped activation term for the classical pathway should allow us to capture classical activation.
+Next, we should address the C3a time scale issue.
+We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
+Tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
+Thus, we skipped an important step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
+The model should be expanded to include the C3wBb intermediate, where C3wBb catalyzes C3 cleavage at a slow rate compared to normal AP or CP C3 convertases.
+We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
+This shortcoming could be addressed by including balances around factor B and D, and including these species in the appropriate kinetic rates.
+The C5a predictions also had an accelerated time scale.
+However, because the C5a time scale depended strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
+Lastly, we should also consider including the C2-by pass pathway, which was not included in the model.
+The C2-bypass mediates lectin pathway activation, without the involvement of MASP-1/2.
+Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement model performance.
+
+% We believe the C3a under-prediction can be attributed to how we modeled C4BP interactions.
+% C4BP interactions were modeled as irreversible binding steps resulting in completely inactive complexes;
+% however, the binding of C4BP with complement proteins is likely reversible and C4BP-bound convertases may have residual activity.
+% We also did not capture the maximum concentration of C3a at low initiator levels.
+% One possible reasons for this could the C2-by pass pathway, which was not included in the model.
+% This pathway further accelerates C3a production without the involvement of a C3 convertase.
+% Currently the C3a in the model is generated only through the activity of a C3 convertase.
+% Incorporating this additional step within the reduced order modeling framework would be a future direction that we need to consider.
+% We should test alternative model structures which include reversible C4BP binding, and partially active convertases.
+% Alternatively, we could also perform sensitivity analysis on the C3a prediction residual to determine which parameters controlled the C3a prediction.
 

### 2016-06-15 Paper_v4.tex
Updates

@@
 Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
-Only dual dual-knockdown of C3 and C5 consistently reduced C3 and C5a formation when all pathways were activated.
+Only the knockdown of both C3 and C5 consistently reduced C3a and C5a formation from all pathways.
 Taken together, we developed a reduced order complement model that was computationally inexpensive,
 and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
-The model described experimental data, and predicted the need for multiple points of therapeutic intervention to disrupt complement activation.
+The model described experimental data, and predicted the need for multiple points of therapeutic intervention to fully disrupt complement activation.
 
@@
 \vspace{0.1in}
-{\noindent \textbf{Keywords:}~Complement system, systems biology, reduced order models, biochemical engineering}
+{\noindent \textbf{Keywords:}~Complement, systems biology, reduced order modeling, biochemical engineering}
 
@@
 Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
-However, research over the past decade has shown the importance of complement extends well beyond innate immunity.
-For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}, and has also been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+However, research over the past decade has shown the importance of complement extends beyond innate immunity.
+For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}, and has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
 Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
@@
 Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
-However, despite its importance, there have been relatively few approved complement specific therapeutics.
+However, despite its importance, there have been relatively few approved complement specific therapeutics, largely because of safety concerns and challenging pharmacokinetic constraints.
 
-The complement cascade involves many soluble and cell surface proteins, receptors and regulators.
-The molecular connectivity of complement involves both fluid phase and cell surface events, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
-The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
+The complement cascade involves many soluble and cell surface proteins, receptors and regulators \cite{Walport:2001ab,Walport:2001aa}.
+The outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
 The membrane attack complex, generated during the terminal phase of the response,
@@
 On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation \cite{sarma2011complement}.
-Complement activation takes places through three pathways: the classical, the lectin binding and the alternate pathways.
+Complement activation takes places through three pathways: the classical, the lectin and the alternate pathways.
 The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
-A multimeric protein complex C1 binds to antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
-This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
-The C4a and C2b fragments bind to form the C4bC2a protease, also known as the classical C3 convertase.
+A multimeric protein complex C1 binds antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
+The activated complex cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
+The C4a and C2b fragments bind to form the C4bC2a protease, also known as the classical pathway C3 convertase (CP C3 convertase).
 The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
-These complexes, in combination with the associated mannose-associated serine proteases
-1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional classical C3 convertase.
-Thus, the classical and lectin pathways, initiated by the recognition of different cues on foreign surfaces,  converge at the classical C3 convertase.
+These complexes, in combination mannose-associated serine proteases
+1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional CP C3 convertase.
+Thus, the classical and lectin pathways, initiated by different cues on foreign surfaces,  converge at the CP C3 convertase.
 However, the alternate pathway works differently. It is activated by a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
 C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb cleaves C3 into C3a and C3b, where
-the C3b fragment further recruits additional factor B and factor D to form C3bBbC3b, the alternate C3 convertase \cite{pangburn1984alternative}.
+the C3b fragment further recruits additional factor B and factor D to form C3bBb, the alternate C3 convertase (AP C3 convertase) \cite{pangburn1984alternative}.
 The role of classical and alternate C3 convertases is varied.
-First, C3 convertases mediate signal amplification.
-C3 convertases cleave C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
+First, AP C3 convertases mediate signal amplification.
+AP C3 convertases cleave C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
 thereby forming a positive feedback loop.
-Next, C3 convertases link complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
-Both classical and alternate C3 convertases can recruit C3b subunits to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
-C5 convertases cleave C5 into the C5a and C5b fragments.
-The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
+Next, AP/CP C3 convertases link complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
+Both classical and alternate C3 convertases can recruit C3b subunits to form the classical pathway C5 convertase (C4bC2aC3b, CP C5 convertase), and the alternate pathway C5 convertase (C3bBbC3b, AP C5 convertase), respectively.
+Both C5 convertases cleave C5 into the C5a and C5b fragments.
+The C5b fragment, along with the complement proteins C6, C7, C8 and multiple C9s, form the membrane attack complex.
 On the other hand, both C3a and C5a are important inflammatory signals involved in several responses \cite{Walport:2001ab,Walport:2001aa}.
-Thus, the complement cascade directly attacks the invading pathogen, and signals for help from the immune system.
+Thus, the complement cascade attacks invading pathogens, while acting as a beacon for adaptive immunity.
 
@@
 %Thus,
-The complement cascade is regulated by plasma and host cell proteins which balance host cell safety with effectiveness.
-The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh).
+The complement cascade is regulated by plasma and host cell proteins which balance host safety with effectiveness.
+The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh);
 C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing chronic complement activation \cite{walker1995complement}.
@@
 Delay accelerating factor (DAF or CD55) also recognizes and dissociates both C3 and C5 convertases on host cell surfaces \cite{lukacik2004complement}.
-More generally, Carboxypeptidase-N, a well known inflammation regulator, has broad activity against the complement proteins C3a, C4a, and C5a, rendering them inactive by cleavage of carboxyl-terminal arginine and lysine residues \cite{liszewski1995control}.
+More generally the well known inflammation regulator Carboxypeptidase-N has broad activity against the complement proteins C3a, C4a, and C5a, rendering them inactive by cleavage of carboxyl-terminal arginine and lysine residues \cite{liszewski1995control}.
 Although Carboxypeptidase-N does not directly influence complement activation, it silences the important inflammatory signals produced by complement.
-Lastly, the assembly of the MAC complex itself can be inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
+Lastly, assembly of the MAC complex itself can be inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
 Thus, there are many points of control which influence complement across the three activation pathways.
@@
 Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
-For example, Hirayama et al. modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
+For example, Hirayama et al., modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
 while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways (up to C3a) as a system of non-linear ODEs \cite{korotaevskiy2009non}.
 More recently, large mechanistic models of sections of complement have also been proposed.
-For example, Liu et al.
-analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
+For example, Liu et al., analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
 Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated between
-the host cell and pathogen \cite{zewde2016quantitative}.
+the fluid, host and pathogen surfaces \cite{zewde2016quantitative}.
 However, these previous modeling studies involved little experimental validation.
@@
 Unlike other important cascades, such as coagulation where there are well developed experimental tools and publicly available data sets,
-the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of large mechanistic complement models difficult.
-Thus, reduced order modeling approaches, which describe the key biology of complement using a limited number of parameters, will be important for developing validated models of complement, and to fully understanding its role in the body.
+the data for complement is relatively sparse. Data sets with missing or incomplete data, and limited dynamic
+data also make the identification of large mechanistic complement models difficult.
+Thus, reduced order approaches which describe the biology of complement using a limited number of species and parameters
+could be important for pharmacokinetic model development, and for our understanding of the varied role of complement in the body.
 
@@
 
-In this study, we developed an ensemble of experimentally validated reduced order complement models.
+In this study, we developed an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
 The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
@@
 After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
-controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention;
+controlled model performance. Sensitivity analysis suggested CP C3 and C5 convertase parameters were critical,
+while robustness analyses suggested complement was robust to any single therapeutic intervention;
 only the knockdown of both C3 and C5 consistently reduced C3a and C5a formation for all cases.
@@
 The complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
-A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
+A trigger event initiated the lectin pathway (encoded as a logical rule), which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
 Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
@@
 Both C5 convertases catalyzed the cleavage of C5 into the C5a and C5b fragments.
-In this initial study, we simplified the model by assuming both Factor B and Factor D were in excess.
-However, we did explicitly account for two control proteins, Factor H and C4BP.
+In this initial study, we simplified the model by assuming both factor B and factor D were in excess.
+However, we did explicitly account for two control proteins, factor H and C4BP.
 Lastly, we did not consider MAC formation, instead we stopped at C5a and C5b.
@@
 A critical challenge for the development of any dynamic model is the estimation of model parameters.
-We estimated an ensemble of complement model parameters in a hierarchical fashion using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
+We estimated an ensemble of complement model parameters using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
 The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) routine,
-starting from an initial random parameter guess. Unless otherwise specified, all initial conditions were assumed to be their mean physiological values.
-A hierarchical approach was taken in which the alternate pathway parameters were estimated first and then fixed during the estimation of the lectin pathway parameters.
+starting from a random parameter guess. The best fit parameter set estimated by DOPS was then used to generate a parameter ensemble using multiobjective optimization.
+Unless otherwise specified, all initial conditions were assumed to be their mean physiological values.
 While we had significant training data, the parameter estimation problem was underdetermined (we were not able to uniquely determine model parameters).
-Thus, instead of using a best-fit yet uncertain parameter set, we estimated an ensemble of probable parameter sets (N = 50, see materials and methods).
-The reduced order complement model ensemble captured the behavior of both the alternative and lectin pathways (Fig. \ref{fig-fit}).
-For the alternative pathway, we used C3a and C5a measurements in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit}A and B).
-On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1g zymosan with alternate pathway parameters fixed (Fig. \ref{fig-fit}C and D).
+Thus, instead of using the best-fit yet uncertain parameter set generated by DOPS, we estimated an ensemble of probable parameter sets to quantify model uncertainty
+(N = 2100, see materials and methods).
+The complement model ensemble captured the behavior of both the alternate and lectin pathways (Fig. \ref{fig-fit}).
+For the alternate pathway, we used C3a and C5a measurements in the absence of zymosan (Fig. \ref{fig-fit}A and B).
+On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1mg/ml zymosan (Fig. \ref{fig-fit}C and D).
 The reduced order model reproduced a panel of alternate and lectin pathway data sets in the neighborhood of physiological factor and inhibitor concentrations.
@@
 
+%A hierarchical approach was taken in which the alternate pathway parameters were estimated first and then fixed during the estimation of the lectin pathway parameters.
 %However, we failed to fully capture the curvature of C5a alternative measureme (Fig. \ref{fig-fit}).
@@
 The rate of C3a formation and C3a peak time were directly proportional to initiator dose.
-Similarly, the C5a plateau and rate of formation were also directly proportional to initiator dose, with the lag time being indirectly proportional to initiator exposure.
+Similarly, the C5a plateau and rate of formation were also directly proportional to initiator dose, with the lag time being indirectly proportional to initiator exposure for both C3a and C5a.
 However, there were shortcomings with model performance.
 First, while the overall C3a trend was captured (within the 99\% confidence interval), the C3a dynamics were too fast with the exception of the low dose case.
-We believe the C3a timescale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
+We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
 We trained the model using either no or 1 mg/ml zymosan, but predicted cases in a different initiator range;
-comparing the training to prediction, the model performance e.g., the shape of the C3a trajectory was biased towards either high or very low initiator doses.
+comparing training to prediction, the model performance e.g., the shape of the C3a trajectory was biased towards either high or very low initiator doses.
 Next, tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
-Thus, we skipped an important step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
+Thus, we skipped an important upstream step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
 We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
@@
 However, because the C5a time scale depends strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
-Despite these shortcomings, we qualitatively predicted unseen experimental measurements within the 99\% confidence of the ensemble, for three inducer levels.
+Despite these shortcomings, we qualitatively predicted unseen experimental measurements typically within the 99\% confidence of the ensemble, for three inducer levels.
 Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.
@@
 well as the rate constant controlling basal C3b formation. The C5a trajectory was sensitive to the AP C5 convertase kinetic parameters (Fig. \ref{fig-SA}A).
-Interestingly, neither the rate nor the saturation constants governing AP C3-convertase activity were sensitive in the absence of zymosan.
-Thus, C3a formation in the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3-convertase activity, in the absence of zymosan.
+Interestingly, neither the rate nor the saturation constant governing AP C3 convertase activity were sensitive in the absence of zymosan.
+Thus, C3a formation in the alternative pathway was more heavily influenced by the spontaneous hydrolysis of C3, rather than AP C3 convertase activity, in the absence of zymosan.
 In the presence of zymosan, the C3a residual was controlled by the formation and activity of the CP C3 convertase, as well as tickover and degradation parameters.
@@
 Thus, sensitivity analysis suggested that CP C3/C5 convertase formation and activity dominated in the presence of zymosan, but tickover parameters and AP C5 convertase were more important
-without initiator. Next, we compared the sensitivity results to current therapeutic approaches;
+without initiator. AP C3 convertase assembly was important, but its activity was not.
+Next, we compared the sensitivity results to current therapeutic approaches;
 pathways involving sensitive parameters have been targeted for clinical intervention (Fig. \ref{fig-SA}C).
@@
 
-Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to a biochemical network.
-Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
 Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
-We calculated robustness indices for C3a and C5a for the parameter ensemble (N = 50) with and without the lectin pathway initiator.
+Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.
+Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
 We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
-This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}E).
+This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}C).
 A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
@@
 Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
-C5a formation was catalyzed by CP C3 and C5 convertases.
-Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a dual-knockdown of C3 and C5.
+even 10\% of the nominal C3 was able to generated enough C5 convertase to maintain C5a formation.
+Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
 For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
@@
 controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
-The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
+The only intervention that consistently reduced C5a formation for all cases was a knockdown of both C3 and C5.
 Taken together, we developed a reduced order complement model that was computationally inexpensive,
@@
 \caption{Reduced order complement model training.
-Model parameters were estimated using Dynamic Optimization with Particle Swarms (DOPS) from C3a and C5a measurements with and without zymosan \cite{morad2015time}.
-The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin pathway initiated with 1g zymosan (\textbf{C}--\textbf{D}).
-The solid red line shows the simulation with the best-fit parameter set, the solid black lines show the simulated mean value of C3a or C5a for the ensemble (N = 50).
-The dark shaded region denotes the 99\% confidence interval of the simulated mean concentrations, while the light shaded region denotes the 99\%
-confidence interval of the best-fit simulation for C3a and C5a. All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.}\label{fig-fit}
+An ensemble of model parameters were estimated using multiobjective optimization from C3a and C5a measurements with and without zymosan \cite{morad2015time}.
+The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin pathway initiated with 1 mg/ml zymosan (\textbf{C}--\textbf{D}).
+The solid black lines show the simulated mean value of C3a or C5a for the ensemble, while the dark shaded region denotes the 99\% confidence interval of mean.
+The light shaded region denotes the 99\% confidence interval of the simulated C3a and C5a concentration.
+All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
+}\label{fig-fit}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final_v3.pdf}
-\caption{Reduced order complement model predictions. The reduced order coagulation model parameter estimates were tested against data not used during model training \cite{morad2015time}.
-Simulations of C3a and C5a generated in the lectin pathway using 0.1g, 0.01g, and 0.001g zymosan were compared with the corresponding experimental measurement.
-The solid red line shows the simulation with the best-fit parameter set, the solid black lines show the simulated mean value of C3a or C5a for the ensemble (N = 50).
-The dark shaded region denotes the 99\% confidence interval of the simulated mean concentrations, while the light shaded region denotes the 99\%
-confidence interval of the best-fit simulation for C3a and C5a. All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
+\caption{Reduced order complement model predictions.
+Simulations of C3a and C5a generated in the lectin pathway using 0.1 mg/ml, 0.01 mg/ml, and 0.001 mg/ml zymosan were compared with the corresponding experimental measurements.
+The solid black lines show the simulated mean value of C3a or C5a for the ensemble, while the dark shaded region denotes the 99\% confidence interval of mean.
+The light shaded region denotes the 99\% confidence interval of the simulated C3a and C5a concentration.
+All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
 }\label{fig-prediction}
@@
 Robustness analysis was conducted on the four cases used for model training:
-C3a alternate (w/o zymosan), C5a alternate (w/o zymosan),  C3a lectin (1g zymosan), and C5a lectin (1g zymosan), by reducing the initial concentration of C3 and/or C5
-by 50\% or 90 \%.
+C3a alternate (w/o zymosan), C5a alternate (w/o zymosan),  C3a lectin (1mg/ml zymosan), and C5a lectin (1mg/ml zymosan), by reducing the initial concentration of C3 and/or C5
+by 50\% or 90\%.
 \textbf{A}: Robustness results for a 50\% decrease in the C3, C5, or C3 and C5 initial condition.

### 2016-06-15 Paper_v4.tex
Many updates

@@
 We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
-This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}C).
+This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
 A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
@@
 controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
-The only intervention that consistently reduced C5a formation for all cases was a knockdown of both C3 and C5.
+The only intervention that consistently reduced C3a and C5a formation for all cases was a knockdown of both C3 and C5.
 Taken together, we developed a reduced order complement model that was computationally inexpensive,
@@
 On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
-The most commonly used anti-complement drug eculizumab, targets the C5 protein \cite{morgan2015complement}.
-A number of other drugs targeting C5 are also being developed. For example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
-while Mubodina, an antibody that targets C5, is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
-Other agents such as the small molecule Coversin \cite{weston2014clinical}, the aptamer Zimura \cite{epstein2007complement},
-as well as small peptides and RNAi could also be used to inhibit C5 \cite{borodovsky2014aln}.
+The most commonly used anti-complement drug Eculizumab, targets the C5 protein \cite{morgan2015complement}.
+Several other antibodies targeting C5 are also being developed; for example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
+while Mubodina is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
+Other agents such as Coversin \cite{weston2014clinical} or the aptamer Zimura \cite{epstein2007complement} could also be used to knockdown C5.
 The cyclic peptide inhibitor Compstatin and its derivatives, which target C3, are promising approaches for the treatment of X, Y and Z [FINISH] \cite{Mastellos:2015aa}.
@@
 The C2-bypass mediates lectin pathway activation, without the involvement of MASP-1/2.
-Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement model performance.
+Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement activation.
 
@@
 \subsection*{Estimating an ensemble of complement model parameters.}
-We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique [REF].
+We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique \cite{DOPS2016}.
 DOPS is a novel hybrid meta-heuristic which combines a multi-swarm particle swarm method with the dynamically dimensioned search approach of Shoemaker and colleagues \cite{Tolson2007}.

### 2016-06-16 Paper_v4.tex
Need to check robustness results
New robustness figure

@@
 Other agents such as Coversin \cite{weston2014clinical} or the aptamer Zimura \cite{epstein2007complement} could also be used to knockdown C5.
-The cyclic peptide inhibitor Compstatin and its derivatives, which target C3, are promising approaches for the treatment of X, Y and Z [FINISH] \cite{Mastellos:2015aa}.
+The peptide inhibitor Compstatin and its derivatives are promising approaches for the inhibition of C3 \cite{Mastellos:2015aa}.
 However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
@@
 \centering
-\includegraphics[width=0.9\textwidth]{./figs/Figure5_RobustnessAnalysis_v2.pdf}
-\caption{Robustness analysis of the complement model with respect to C3 and C5 initial conditions.
-Robustness analysis was conducted on the four cases used for model training:
-C3a alternate (w/o zymosan), C5a alternate (w/o zymosan),  C3a lectin (1mg/ml zymosan), and C5a lectin (1mg/ml zymosan), by reducing the initial concentration of C3 and/or C5
-by 50\% or 90\%.
-\textbf{A}: Robustness results for a 50\% decrease in the C3, C5, or C3 and C5 initial condition.
-\textbf{B}: Robustness results for a 90\% decrease in the C3, C5, or C3 and C5 initial condition.
-The bars denote the log-transformated robustness index while error bars denote one standard deviation. At zero, the perturbed initial concentration has no impact on the measured output.
-A log-transformed robustness index less than zero indicates a negative relation between the perturbed initial concentration and the measured output.}\label{fig-robustness-analysis}
+\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v3.pdf}
+\caption{Robustness analysis of the complement model.
+Robustness coefficients were calculated for a 50\%, 90\% and 99\% reduction in C3, C5, or C3 and C5 initial conditions.
+\textbf{A}: Mean robustness index for C3a and C5a generated from the alternate pathway (w/o zymosan).
+\textbf{B}: Mean robustness index for C3a and C5a generated from the lectin and alternate pathway (1 mg/ml zymosan).
+The color describes the degree of reduction of C3a or C5a following the network perturbation.
+}\label{fig-robustness-analysis}
 \end{figure}

### 2016-06-16 Paper_v4.tex
Submission to BioArXiv

@@
 Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
-We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
-This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
-A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
+We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3, C5 or the combination of C3 and C5 by 50\%, 90\% and 99\%.
+This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig.
+\ref{fig-SA}C).
 The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
-As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
-(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
-However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
+As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A) or lectin pathways
+(Fig. \ref{fig-robustness-analysis}B).
+However, C3a and to a greater extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A).
 This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
-Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig.
-\ref{fig-robustness-analysis}A or B, lane 2).
-Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
-even 10\% of the nominal C3 was able to generated enough C5 convertase to maintain C5a formation.
+Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway.
+Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}B);
+even 1\% of the nominal C3 was able to generated enough AP/CP C5 convertase to maintain C5a formation.
 Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
-For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
-while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).
+For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by an order of magnitude,
+while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B).
 
@@
 If $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)\sim{0}$, perturbation $j$ did not influence the abundance of marker $i$.
-In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\% or 90\% and measured the area under the curve (AUC)
-of C3a or C5a with and without lectin initiator.
-Log-transformed robustness coefficients were calculated for every member of the ensemble, where the mean $\pm$ 1$\times$standard-deviation are reported.
+In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\%, 90\% and 99\% and measured the area under the curve (AUC)
+of C3a or C5a with and without lectin initiator. We computed the robustness coefficients for a subset of the parameter ensemble.
 
@@
 The color describes the degree of reduction of C3a or C5a following the network perturbation.
+Robustness coefficients were calculated using all parameter sets with Pareto rank less than five (N = 65).
 }\label{fig-robustness-analysis}

### 2016-06-16 Paper_v4.tex
Updated w/AS and MM changes

@@
 For example, Hirayama et al., modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
-while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways (up to C3a) as a system of non-linear ODEs \cite{korotaevskiy2009non}.
+while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways as a system of non-linear ODEs \cite{korotaevskiy2009non}.
 More recently, large mechanistic models of sections of complement have also been proposed.
@@
 the fluid, host and pathogen surfaces \cite{zewde2016quantitative}.
-However, these previous modeling studies involved little experimental validation.
+However, these previous modeling studies involved large models with little experimental validation.
 Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict complement measurements.
@@
 Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}B);
-even 1\% of the nominal C3 was able to generated enough AP/CP C5 convertase to maintain C5a formation.
+even 1\% of the nominal C3 was able to generate enough AP/CP C5 convertase to maintain C5a formation.
 Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
@@
 However, because the C5a time scale depended strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
-Lastly, we should also consider including the C2-by pass pathway, which was not included in the model.
+Lastly, we should also consider including the C2-bypass pathway, which was not included in the model.
 The C2-bypass mediates lectin pathway activation, without the involvement of MASP-1/2.
-Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement activation.
+Thus, this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement activation.
 

### 2016-06-16 Paper_v4.tex
Added tau language

@@
 \end{equation}
-where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of protein species in the model.
-The quantity $\tau_{i}$ denotes a timescale parameter for species $i$ which captures unmodeled effects; by default $\tau_{i}$ = 1, for all species unless otherwise specified.
+where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of proteins in the model.
+The quantity $\tau_{i}$ denotes a time scale parameter for species $i$ which captures unmodeled effects.
+For the current study, $\tau$ scaled with the level of initiator ($z$) for C5a and C5b; $\tau_{i}=z/z^{*}$ for i = C5a, C5b where $z^{*}$ was 1mg/ml, $\tau_{i}$ = 1 for all other species.
 The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.

### 2016-06-16 Paper_v4.tex
Added tau language
Made David's changes

@@
 The model also captured the dynamics of C3a and C5a formed from the alternate pathway using only seven ODEs.
-The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al which involved over 100 ODEs \cite{zewde2016quantitative}.
-However, we found that the quantity of C3a produced in the alternate pathway was nearly 1000 times the amount of C5a generated.
+The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al., which involved over 100 ODEs \cite{zewde2016quantitative}.
+However, we found that the C3a produced in the alternate pathway was nearly three orders of magnitude greater than the C5a generated.
 While this was in agreement with the experimental data \cite{morad2015time},
-it differed from the theoretical predictions made by Zewde et al. who showed C3a was ${10^8}$ times greater than the C5a concentration \cite{zewde2016quantitative}.
+it differed from the theoretical predictions made by Zewde et al., who showed C3a was eight orders of magnitude greater than the C5a concentration \cite{zewde2016quantitative}.
 In our model, the time profile of both C3a and C5a generated changed with respect to the quantity of zymosan (the lectin pathway initiator).
@@
 Sensitivity analysis was conducted on the two objectives used for model training.
-\textbf{A}: Sensitivity of the C3a and C5a residual w/o zymosan,
-\textbf{B}: Sensitivity of the C3a and C5a residual with 1mg/ml zymosan.
+\textbf{A}: Sensitivity of the C3a and C5a residual w/o zymosan.
+\textbf{B}: Sensitivity of the C3a and C5a residual with 1 mg/ml zymosan.
 The bars denote the mean total sensitivity index for each parameter, while the error bars denote the 95\% confidence interval.

### 2016-10-05 Paper_v4.tex
Updated figures
Introdution edited

@@
 \vspace{0.10in}
-{\par \centering {School of Chemical and Biomolecular Engineering}}
+{\par \centering {Robert Frederick Smith School of Chemical and Biomolecular Engineering}}
 {\par \centering {Cornell University, Ithaca NY 14853}}
@@
 {\par \centering Jeffrey D. Varner,}
-{\par \centering Professor, School of Chemical and Biomolecular Engineering,}
+{\par \centering Professor, Robert Frederick Smith School of Chemical and Biomolecular Engineering,}
 {\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
@@
 Complement is an important pathway in innate immunity, inflammation, and many disease processes.
-However, despite its importance, there have been few validated mathematical models of complement activation.
+However, despite its importance, there are few validated mathematical models of complement activation.
 In this study, we developed an ensemble of experimentally validated reduced order complement models.
-We combined ordinary differential equations with logical rules to produce a compact yet predictive complement model.
+We combined ordinary differential equations with logical rules to produce a compact yet predictive model of complement activation.
 The model,  which described the lectin and alternative pathways, was an order of magnitude smaller than comparable models in the literature.
@@
 Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
-Only the knockdown of both C3 and C5 consistently reduced C3a and C5a formation from all pathways.
-Taken together, we developed a reduced order complement model that was computationally inexpensive,
+Only the simultaneous knockdown of both C3 and C5 consistently reduced C3a and C5a formation from all pathways.
+Taken together, we developed a validated mathematical model of complement activation that was computationally inexpensive,
 and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
@@
 Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
-However, research over the past decade has shown the importance of complement extends beyond innate immunity.
-For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}, and has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+However, research over the past decade has suggested the importance of complement extends beyond innate immunity.
+For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}.
+It has also has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
 Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
@@
 Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
-However, despite its importance, there have been relatively few approved complement specific therapeutics, largely because of safety concerns and challenging pharmacokinetic constraints.
+Despite its importance, there have been few approved complement specific therapeutics,
+largely because of safety concerns and challenging pharmacokinetic constraints, however, progress is being made \cite{Ricklin:2013aa}.
 
@@
 A multimeric protein complex C1 binds antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
-The activated complex cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
+The activated C1-complex cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
 The C4a and C2b fragments bind to form the C4bC2a protease, also known as the classical pathway C3 convertase (CP C3 convertase).
 The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
-These complexes, in combination mannose-associated serine proteases
+These complexes, in combination with mannose-associated serine proteases
 1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional CP C3 convertase.
 Thus, the classical and lectin pathways, initiated by different cues on foreign surfaces,  converge at the CP C3 convertase.
-However, the alternate pathway works differently. It is activated by a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
+On the other hand, the alternate pathway is activated by a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
 C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb cleaves C3 into C3a and C3b, where
@@
 %Thus,
-The complement cascade is regulated by plasma and host cell proteins which balance host safety with effectiveness.
+The complement cascade is regulated by plasma and host cell surface proteins which balance host safety with effectiveness.
 The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh);
@@
 
-Developing quantitative mathematical models of complement will be crucial to fully understanding its role in the body.
+Developing quantitative mathematical models of complement could be crucial to fully understanding its role in the body.
 Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
@@
 the fluid, host and pathogen surfaces \cite{zewde2016quantitative}.
-However, these previous modeling studies involved large models with little experimental validation.
+However, these previous studies involved large models with little experimental validation.
 Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict complement measurements.
@@
 
-\subsection*{Sensitivity and robustness analysis of complement model performance.}
+\subsection*{Sensitivity and robustness analysis.}
 We conducted global sensitivity and robustness analysis to estimate which parameters and species controlled the performance of the reduced order model.
 We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a/C5a alternate and C3a/C5a lectin objectives
-using the Sobol method \cite{SOBOL_METHOD}. The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
+using the Sobol method \cite{SOBOL_METHOD}.
+sThe sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
 We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
@@
 \section*{Funding}
-This study was supported by an award from the US Army and Systems Biology of Trauma Induced Coagulopathy (W911NF-10-1-0376) to J.V. for the support of A.S.
+This material is based upon work supported by, or in part by, the U. S. Army Research Laboratory and the U. S. Army Research Office under contract/grant number W911NF1020114.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v3.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v4.pdf}
 \caption{Robustness analysis of the complement model.

### 2016-10-05 Paper_v4.tex
Working thru the results section

@@
 
-In this study, we developed an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
+In this study, we estimated an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
 The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
@@
 We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
-Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
-Despite its small size, the model was surprisingly predictive.
+Subsequently, we validated the model on unseen C3a and C5a measurements not used for model training.
+Despite its size, the model was surprisingly predictive.
 After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
@@
 The complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
-A trigger event initiated the lectin pathway (encoded as a logical rule), which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
+A trigger event initiated the lectin pathway (encoded as a logical rule), which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b, respectively.
 Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
-The alternate pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b (not C3w).
-The C3b fragment generated by hydrolysis (or by CP C3 convertase) could then form the alternate pathway (AP) C3 convertase (C3bBb).
+The alternate pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
+The C3b fragments generated by hydrolysis (or by CP C3 convertase) could then form the alternate pathway (AP) C3 convertase (C3bBb).
 We did not consider C3w, nor the formation of the initial alternate C3 convertase (C3wBb).
@@
 Both C5 convertases catalyzed the cleavage of C5 into the C5a and C5b fragments.
-In this initial study, we simplified the model by assuming both factor B and factor D were in excess.
-However, we did explicitly account for two control proteins, factor H and C4BP.
+In this study, we simplified the model by assuming both factor B and factor D were in excess.
+However, we did explicitly account for the action of two other control proteins, factor H and C4BP.
 Lastly, we did not consider MAC formation, instead we stopped at C5a and C5b.
-Lectin pathway activation, and C3/C5 convertase activity was modeled using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
+Lectin pathway activation, and C3/C5 convertase activity were modeled using a combination of saturation kinetics and non-linear transfer functions,
+which resulted in a significant size reduction of the model, while maintaining performance.
 Binding interactions were modeled using mass-action kinetics, where we assumed all binding was irreversible.
@@
 We estimated an ensemble of complement model parameters using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
-The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) routine,
-starting from a random parameter guess. The best fit parameter set estimated by DOPS was then used to generate a parameter ensemble using multiobjective optimization.
-Unless otherwise specified, all initial conditions were assumed to be their mean physiological values.
+The residual between model simulations and experimental measurements was minimized using the Pareto Optimal Ensemble Technique (JuPOETs) [REFs] starting
+from a initial guess generated by the dynamic optimization with particle swarms (DOPS) routine [REF].
+Unless otherwise specified, all initial conditions were assumed to be at their mean physiological values.
 While we had significant training data, the parameter estimation problem was underdetermined (we were not able to uniquely determine model parameters).
-Thus, instead of using the best-fit yet uncertain parameter set generated by DOPS, we estimated an ensemble of probable parameter sets to quantify model uncertainty
+Thus, instead of using the best-fit yet uncertain parameter set, we estimated an ensemble of probable parameter sets to quantify model uncertainty
 (N = 2100, see materials and methods).
 The complement model ensemble captured the behavior of both the alternate and lectin pathways (Fig. \ref{fig-fit}).
-For the alternate pathway, we used C3a and C5a measurements in the absence of zymosan (Fig. \ref{fig-fit}A and B).
+To estimate alternate pathway model parameters, we used C3a and C5a measurements in the absence of zymosan (Fig. \ref{fig-fit}A and B).
 On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1mg/ml zymosan (Fig. \ref{fig-fit}C and D).

### 2016-10-06 Paper_v4.tex
Updated the text

@@
 Next, tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
-Thus, we skipped an important upstream step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
+Thus, we skipped an important upstream step which could influence AP C3 convertase formation by attenuating the rate C3 cleavage into C3a and C3b.
 We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
@@
 However, because the C5a time scale depends strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
-Despite these shortcomings, we qualitatively predicted unseen experimental measurements typically within the 99\% confidence of the ensemble, for three inducer levels.
+Despite these shortcomings, we qualitatively predicted experimental measurements not used for model training typically within the 99\% confidence of the ensemble, for three inducer levels.
 Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.
@@
 In the absence of zymosan (where only the alternative pathway is active), the most sensitive parameter was the rate constant governing the assembly of the AP C3 convertase, as
-well as the rate constant controlling basal C3b formation. The C5a trajectory was sensitive to the AP C5 convertase kinetic parameters (Fig. \ref{fig-SA}A).
+well as the rate constant controlling basal C3b formation via the tickover mechanism. The C5a trajectory was sensitive to the AP C5 convertase kinetic parameters (Fig. \ref{fig-SA}A).
 Interestingly, neither the rate nor the saturation constant governing AP C3 convertase activity were sensitive in the absence of zymosan.
@@
 Thus, there was at least a qualitative overlap between sensitivity and the potential of biochemical efficacy.
-However, sensitivity coefficients are only a local measure of how small changes in parameters affect model performance.
+However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
 To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
@@
 Robustness coefficients were calculated as shown previously \cite{Tasseff:2011aa}.
-Log-transformed robustness coefficients denoted by $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)$ are defined as:
+Log-transformed robustness coefficients denoted by $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)$ were defined as:
 \begin{equation}
@@
 \end{equation}
-Here $t_o$ and $t_f$ denote the initial and final simulation time, while $i$ and $j$ denote the indices for the marker and the perturbation, respectively.
+Here, $t_o$ and $t_f$ denote the initial and final simulation time, while $i$ and $j$ denote the indices for the marker and the perturbation, respectively.
 A value of $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)>0$, indicates increased marker abundance, while $\hat{\alpha}\left(i,j,t_{o},t_{f}\right)<0$ indicates decreased marker abundance following perturbation $j$.
@@
 In this study, we perturbed the initial condition of C3 or C5 or a combination of C3 and C5 by 50\%, 90\% and 99\% and measured the area under the curve (AUC)
-of C3a or C5a with and without lectin initiator. We computed the robustness coefficients for a subset of the parameter ensemble.
+of C3a or C5a with and without lectin initiator. We computed the robustness coefficients for a subset of the parameter ensemble (N = 65) and reported the mean robustness value.
 

### 2016-10-06 Paper_v4.tex
Added the AIC

@@
 JuPOETs is open source, available under an MIT software license.
-The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
-The objective functions used in this study are available in the GitHub model repository available from http://varnerlab.org.
+The JuPOETs source code is freely available from the JuPOETs GitHub repository \cite{JuPOETS-GitHub}.
+The objective functions used in this study are available in the GitHub model repository \cite{Varnerlab}.
+
+The simulation and prediction performance of the complement model was measured using the Akaike information criterion (AIC) \cite{Akaike1974}.
+In this study, we implemented the AIC as:
+\begin{equation}\label{eq:AIC}
+	AIC = 2N_{p}+N_{m}\ln\left(\frac{1}{\norm{\mathcal{M}}}\sum_{\tau}\left(x_{\tau}-y_{\tau}\right)^2\right)
+\end{equation}where $N_{p}, N_{m}$ denotes the number of parameters, and the number of experimental measurements, respectively.
+The summation term in Eq. \eqref{eq:AIC} denotes the residual between the model simulation ($x$) and experimental measurements ($y$),
+where the residual is normalized by the scale of the experimental data ($\norm{\mathcal{M}}$).
+We compared the AIC for the model parameters estimated in this study, with a random parameter control generated to have a similar order of magnitude.
+The mean and standard deviation of the AIC was calculated over the parameter ensemble and the random parameter control were reported in this study.
 

### 2016-10-11 Paper_v4.tex
Updates

@@
 The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
-variable. Each regulatory transfer function took the form:
-\begin{equation}\label{eqn:control-factor}
-	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right)=k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
-\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance of factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
+variable. Each regulatory transfer function was modeled using a Hill function.
 In this study, we used $\mathcal{I}_{j}\in\left\{min,max\right\}$ \cite{pr3010178}. If a process has no modifying factors, $v_{j}=1$.
 The model equations were implemented in Julia and solved using the \texttt{CVODE} routine of the Sundials package \cite{Julia,Sundials}.
-The model code and parameter ensemble is freely available under an MIT software license and can be downloaded from http://www.varnerlab.org.
+The model code and parameter ensemble is freely available under an MIT software license and can be downloaded from the Varnerlab website \cite{Varnerlab}.
 
-\subsection*{Estimating an ensemble of complement model parameters.}
+\subsection*{Estimating complement model parameters.}
 We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique \cite{DOPS2016}.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis_v4.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure5_PairwiseSensitivity_v1.pdf}
+\caption{Pairwise sensitivity and clustering of complement model parameters.
+The response of the complement model was calculated for each parameter combination following a 10\% increase in a model parameter in the presence of 1 mg/ml zymosan.
+The model response was clustered into high (black), medium (gray) and low (white) response based upon the euclidian distance. The blue group of parameters are the most sensitive while
+the parameters in the green cluster are the least sensitive.
+}\label{fig-pairwise-analysis}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1.0\textwidth]{./figs/Figure6_RobustnessAnalysis_v4.pdf}
 \caption{Robustness analysis of the complement model.

### 2016-10-11 Paper_v4.tex
Working on the draft

@@
 \vspace{0.05in}
-{\par \centering \large{Adithya Sagar, Wei Dai$^{\#}$, Mason Minot$^{\#}$, and Jeffrey D. Varner$^{*}$}}
+{\par \centering \large{Adithya Sagar, Wei Dai$^{\#}$, Mason Minot$^{\#}$, Rachel LeCover, and Jeffrey D. Varner$^{*}$}}
 \vspace{0.10in}
@@
 We estimated an ensemble of complement model parameters using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
-The residual between model simulations and experimental measurements was minimized using the Pareto Optimal Ensemble Technique (JuPOETs) [REFs] starting
-from a initial guess generated by the dynamic optimization with particle swarms (DOPS) routine [REF].
+The residual between model simulations and experimental measurements was minimized using the Pareto Optimal Ensemble Technique (JuPOETs) \cite{JuPOETs-BioArXiv}
+starting from a initial guess generated by the dynamic optimization with particle swarms (DOPS) routine.
 Unless otherwise specified, all initial conditions were assumed to be at their mean physiological values.
@@
 Thus, there was at least a qualitative overlap between sensitivity and the potential of biochemical efficacy.
+However, total sensitivity coefficients quantify how simultaneous changes in many parameters e.g., rate or saturation constants affect model performance (in this case model fit).
+To better understand the role of each parameter, and parameter combination, we explored how finite changes in parameter combinations influenced model performance.
+
+Pairwise parameter perturbations identified critical modes of crosstalk which controlled fibrinolysis model performance (Fig. \ref{fig-pairwise-analysis}).
+We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed system state, and the nominal state for each parameter set in the ensemble.
+We clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal cases into low (green), medium (red) and high (blue) response clusters.
+Four of the 28 parameters (or approximately 14\% of the overall parameters) were clustered in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
+Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster). [Finish me].
+Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
 However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
-To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed
-robustness analysis.
+To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed robustness analysis.
 
@@
 We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3, C5 or the combination of C3 and C5 by 50\%, 90\% and 99\%.
-This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig.
-\ref{fig-SA}C).
+This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
 The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
 \includegraphics[width=1.0\textwidth]{./figs/Figure5_PairwiseSensitivity_v1.pdf}
-\caption{Pairwise sensitivity and clustering of complement model parameters.
-The response of the complement model was calculated for each parameter combination following a 10\% increase in a model parameter in the presence of 1 mg/ml zymosan.
-The model response was clustered into high (black), medium (gray) and low (white) response based upon the euclidian distance. The blue group of parameters are the most sensitive while
-the parameters in the green cluster are the least sensitive.
+\caption{Pairwise sensitivity and clustering of complement model parameters in the presence of 1 mg/ml zymosan.
+The response of the complement model was calculated for each parameter combination following a 10\% increase in parameter combinations in the presence of 1 mg/ml zymosan.
+The model parameters were clustered into high (blue), medium (red) and low (green) response clusters based upon the euclidian distance between the perturbed and nominal system state.
 }\label{fig-pairwise-analysis}
@@
 Robustness coefficients were calculated using all parameter sets with Pareto rank less than five (N = 65).
-}\label{fig-robustness-analysis}
+Mean robustness values were reported. }\label{fig-robustness-analysis}
 \end{figure}

### 2016-10-13 Paper_v4.tex
Added some text re double parameter study

@@
 
-Pairwise parameter perturbations identified critical modes of crosstalk which controlled fibrinolysis model performance (Fig. \ref{fig-pairwise-analysis}).
-We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed system state, and the nominal state for each parameter set in the ensemble.
-We clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal cases into low (green), medium (red) and high (blue) response clusters.
-Four of the 28 parameters (or approximately 14\% of the overall parameters) were clustered in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
-Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster). [Finish me].
+Pairwise parameter perturbations identified modes of crosstalk which controlled fibrinolysis model performance (Fig. \ref{fig-pairwise-analysis}).
+We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
+We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green), medium (red) and high (blue) response clusters.
+Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
+These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant governing CP C3 convertase activity (\#22).
+The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
+Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
+Three parameters (\#10,~\#1,~\#27) were especially important in the medium impact cluster.
+The reaction order governing CP C3 convertase activity was important (\#10),
+along with the catalytic rate constant controlling the formation of C4a and C4b in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
 Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
+Many of these parameters influenced complement activation; for example, parameter \#23 (the K$_M$ of CP C3 convertase) was important, just not to the extent of the other parameters.
+The pairwise perturbation study also identified synergistic crosstalk between parameters.
+For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
+Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
+On the other hand, simultaneous changes in parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
 However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
-To more closely simulate a clinical intervention e.g., administration of an anti-complement inhibitor, we performed robustness analysis.
+To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis.
 
-Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
+Robustness analysis in the absence of flow suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig.
+\ref{fig-robustness-analysis}).
 Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.

### 2016-10-13 Paper_v4.tex
Added the AIC blurb

@@
 The reduced order model reproduced a panel of alternate and lectin pathway data sets in the neighborhood of physiological factor and inhibitor concentrations.
-However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
+The model fit for parameter sets estimated by JuPOETs, quantified by the Akaike information criterion (AIC), was statistically significantly different than a random parameter control
+for each case at a 95\% confidence level. However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
 To address this question, we fixed the model parameters and simulated data sets not used for model training.
@@
 We tested the predictive power of the reduced order complement model with data not used during model training (Fig. \ref{fig-prediction}).
-Six validation cases were considered, three for C3a and C5a, respectively.
+Six validation cases were considered, three for C3a and C5a each, respectively.
+Similar to model training, we compared the AIC for each prediction case to a randomized parameter family.
 All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan, and other experimentally mandated changes).
@@
 and C5a formation (Fig. \ref{fig-prediction}, bottom) at three inducer concentrations.
+For each training case, the AIC was statistically significantly different than the random parameter control for a 95\% confidence level.
 The rate of C3a formation and C3a peak time were directly proportional to initiator dose.
@@
 
-Pairwise parameter perturbations identified modes of crosstalk which controlled fibrinolysis model performance (Fig. \ref{fig-pairwise-analysis}).
-We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
-We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green), medium (red) and high (blue) response clusters.
+Pairwise parameter perturbations identified crosstalk within the complement model (Fig. \ref{fig-pairwise-analysis}).
+We perturbed each pairwise combination of parameters by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
+We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green),
+medium (red) and high (blue) response clusters. A low response (white) meant the parameter perturbations did not significantly change the system state compared with the nominal case.
 Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
@@
 Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
-Three parameters (\#10,~\#1,~\#27) were especially important in the medium impact cluster.
+Three parameters (\#10,~\#1,~\#27) were especially important in this cluster;
 The reaction order governing CP C3 convertase activity was important (\#10),
-along with the catalytic rate constant controlling the formation of C4a and C4b in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
+along with the rate constant controlling C4a and C4b formation from C4 in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
 Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
-Many of these parameters influenced complement activation; for example, parameter \#23 (the K$_M$ of CP C3 convertase) was important, just not to the extent of the other parameters.
-The pairwise perturbation study also identified synergistic crosstalk between parameters.
+Many of these parameters influenced complement activation; for example, parameter \#23 (the CP C3 convertase saturation constant) was important, just not to the extent of other model parameters.
+Pairwise synergistic interactions between parameters were also identified.
 For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
 Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
-On the other hand, simultaneous changes in parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
+On the other hand, simultaneously changing parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
+However, the most striking examples of synergy occurred in the medium impact cluster; for example, simultaneously increasing parameters \#13 (rate constant governing AP C3 convertase formation)
+and \#19 (saturation constant governing AP C5 convertase activity) significantly changed the model state.
+Changes in parameter \#3 (rate constant governing C2a and C2b formation from C2) showed both positive and negative synergistic effects
+depending upon the other parameter that was perturbed.
+Taken together, [FINISH ME].
 However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
-To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis.
+To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis in the absence and presence of flow.
 

### 2016-10-14 Paper_v4.tex
Some updates - and flow model

@@
 
+\subsubsection*{Complement activation under flow conditions.}
+We estimated the dynamics of complement activation under flow using a two-compartment model, with variable compartment volumes.
+We considered a main compartment (m), and a wound compartment (w), where complement was activated in the wound compartment by the addition of a pathogenic surface.
+In the main compartment, the balance for species $i$ ($x_{m,i}$) was given by:
+\begin{equation}
+	\frac{V_{m}}{\tau_{i}}\frac{dx_{m,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{m},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{m}-k_{m,w,i}x_{m,i}+k_{w,m,i}x_{w,i}-Cx_{m,i}
+\end{equation}where $V_{m}$ denotes the volume of the main compartment, $k_{p,q,i}$ denotes transfer constant governing the transfer of species $i$ from compartment $p$ to compartment $q$,
+and $C$ denotes the clearance constant from the main compartment. We assumed complement factors were synthesized in the main compartment.
+The balance governing species $i$ in the wound compartment was given by:
+\begin{equation}
+	\frac{V_{w}}{\tau_{i}}\frac{dx_{w,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{w},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{w}-k_{w,m,i}x_{w,i}+k_{m,w,i}x_{m,i}-Bx_{w,i}
+\end{equation}where $V_{w}$ denotes the volume of the wound compartment, and $B$ denotes the rate of blood loss from the wound compartment.
+Lastly, because of the volume loss through clearance in the main compartment, and bleeding from the wound compartment, the volumes of each compartment were dynamic modeled:
+\begin{eqnarray}
+	\frac{dV_{m}}{dt} &=& I_{m}+F_{w,m} - F_{m,w} - C\\
+	\frac{dV_{w}}{dt} &=& I_{w}+F_{m,w} - F_{w,m} - B
+\end{eqnarray}where $I_{m},~I_{w}$ denote the rate of liquid input into the main and wound compartment.
+
 \subsection*{Estimating complement model parameters.}
@@
 In this study, we minimized two training objectives, the total C3a and C5a residual w/o zymosan (O$_{1}$) and the total C3a and C5a residual for 1 mg/ml zymosan (O$_{2}$).
-JuPOETs identified an ensemble of N $\simeq$ 2100 parameter sets which were used for model simulations and uncertainty quantification subsequently.
+JuPOETs identified an ensemble of N = 2100 parameter sets which were used for model simulations and uncertainty quantification subsequently.
 JuPOETs is open source, available under an MIT software license.
@@
 
-\subsection*{Sensitivity and robustness analysis.}
-We conducted global sensitivity and robustness analysis to estimate which parameters and species controlled the performance of the reduced order model.
+\subsection*{Complement model analysis.}
+\subsubsection*{Global sensitivity analysis.}
+We conducted global sensitivity analysis to estimate which parameters and species controlled the performance of the reduced order model.
 We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a/C5a alternate and C3a/C5a lectin objectives
 using the Sobol method \cite{SOBOL_METHOD}.
-sThe sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
+The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
 We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
@@
 
+\subsubsection*{Pairwise sensitivity analysis and clustering.}
+We perturbed each pair of model parameters by 10\% of their nominal value, and then calculated the euclidian distance between the perturbed and nominal system states for
+physiological conditions. We repeated this calculation for each member of the parameter ensemble, and calculated the mean differences between the perturbed and nominal states.
+We then clustered the resulting \texttt{log10} transformed mean distances using the \texttt{Clustergram} routine in MATLAB (The Mathworks, Natick MA).
+We considered three clusters, high, medium and low displacement.
+
+\subsubsection*{Robustness analysis.}
 Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture.
@@
 J.V directed the study. A.S developed the reduced order complement model and the parameter ensemble.
-A.S, W.D and M.M analyzed the model ensemble, and generated figures for the manuscript.
-The manuscript was prepared and edited for publication by A.S, W.D, M.M and J.V.
+A.S, W.D, R.L and M.M analyzed the model ensemble, and generated figures for the manuscript.
+The manuscript was prepared and edited for publication by A.S, W.D, M.M, R.L and J.V.
 

### 2016-12-20 plos_one_submission/Paper_v4.tex
PLoS submission

--- 2016-10-14 Paper_v4.tex
@@
 %\bibliographystyle{chicago}
-\bibliographystyle{plos2009}
+\bibliographystyle{plos2015}
 
@@
 \vspace{0.1in}
-{\par \centering \textbf{To be submitted:}~\emph{Scientific~Reports}}
+{\par \centering \textbf{To be submitted:}~\emph{PLoS~ONE}}
 \vspace{0.5in}
@@
 For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}.
-It has also has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
+It has also has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{Ricklin:2007aa, rittirsch2008harmful}.
 Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
@@
 Pairwise parameter perturbations identified crosstalk within the complement model (Fig. \ref{fig-pairwise-analysis}).
-We perturbed each pairwise combination of parameters by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
+We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
 We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green),
@@
 Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
-These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant governing CP C3 convertase activity (\#22).
+These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant
+governing CP C3 convertase activity (\#22).
 The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
@@
 depending upon the other parameter that was perturbed.
-Taken together, [FINISH ME].
-However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
-To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis in the absence and presence of flow.
+Taken together, sensitivity coefficients quantified how changes in parameters or parameter combinations affected model performance.
+However, individual parameters e.g., rate or saturation constants are not easily druggable.
+To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed knock-down analysis on the initial values of C3 and C5
+in the absence and presence of flow.
 
-Robustness analysis in the absence of flow suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig.
+Knock-down analysis in the absence of flow suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig.
 \ref{fig-robustness-analysis}).
@@
 while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B).
+Taken together, these results suggested that both C3 and C5 inhibitors must be administered in the presence of both intiation pathways.
 
@@
 
-Despite its importance, there has been a paucity of validated mathematical models of complement pathway activation.
+There has been a paucity of validated mathematical models of complement pathway activation.
 To our knowledge, this study is one of the first complement models that combined multiple initiation pathways with experimental validation of important complement products like C5a.
@@
 Complement malfunctions are implicated in a spectrum of diseases, however the
-development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
+development of complement specific therapeutics has been challenging \cite{Ricklin:2007aa,morgan2015complement}.
 Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
@@
 
-\subsubsection*{Complement activation under flow conditions.}
-We estimated the dynamics of complement activation under flow using a two-compartment model, with variable compartment volumes.
-We considered a main compartment (m), and a wound compartment (w), where complement was activated in the wound compartment by the addition of a pathogenic surface.
-In the main compartment, the balance for species $i$ ($x_{m,i}$) was given by:
-\begin{equation}
-	\frac{V_{m}}{\tau_{i}}\frac{dx_{m,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{m},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{m}-k_{m,w,i}x_{m,i}+k_{w,m,i}x_{w,i}-Cx_{m,i}
-\end{equation}where $V_{m}$ denotes the volume of the main compartment, $k_{p,q,i}$ denotes transfer constant governing the transfer of species $i$ from compartment $p$ to compartment $q$,
-and $C$ denotes the clearance constant from the main compartment. We assumed complement factors were synthesized in the main compartment.
-The balance governing species $i$ in the wound compartment was given by:
-\begin{equation}
-	\frac{V_{w}}{\tau_{i}}\frac{dx_{w,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{w},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{w}-k_{w,m,i}x_{w,i}+k_{m,w,i}x_{m,i}-Bx_{w,i}
-\end{equation}where $V_{w}$ denotes the volume of the wound compartment, and $B$ denotes the rate of blood loss from the wound compartment.
-Lastly, because of the volume loss through clearance in the main compartment, and bleeding from the wound compartment, the volumes of each compartment were dynamic modeled:
-\begin{eqnarray}
-	\frac{dV_{m}}{dt} &=& I_{m}+F_{w,m} - F_{m,w} - C\\
-	\frac{dV_{w}}{dt} &=& I_{w}+F_{m,w} - F_{w,m} - B
-\end{eqnarray}where $I_{m},~I_{w}$ denote the rate of liquid input into the main and wound compartment.
+% \subsubsection*{Complement activation under flow conditions.}
+% We estimated the dynamics of complement activation under flow using a two-compartment model, with variable compartment volumes.
+% We considered a main compartment (m), and a wound compartment (w), where complement was activated in the wound compartment by the addition of a pathogenic surface.
+% In the main compartment, the balance for species $i$ ($x_{m,i}$) was given by:
+% \begin{equation}
+% 	\frac{V_{m}}{\tau_{i}}\frac{dx_{m,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{m},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{m}-k_{m,w,i}x_{m,i}+k_{w,m,i}x_{w,i}-Cx_{m,i}
+% \end{equation}where $V_{m}$ denotes the volume of the main compartment, $k_{p,q,i}$ denotes transfer constant governing the transfer of species $i$ from compartment $p$ to compartment $q$,
+% and $C$ denotes the clearance constant from the main compartment. We assumed complement factors were synthesized in the main compartment.
+% The balance governing species $i$ in the wound compartment was given by:
+% \begin{equation}
+% 	\frac{V_{w}}{\tau_{i}}\frac{dx_{w,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{w},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{w}-k_{w,m,i}x_{w,i}+k_{m,w,i}x_{m,i}-Bx_{w,i}
+% \end{equation}where $V_{w}$ denotes the volume of the wound compartment, and $B$ denotes the rate of blood loss from the wound compartment.
+% Lastly, because of the volume loss through clearance in the main compartment, and bleeding from the wound compartment, the volumes of each compartment were dynamic modeled:
+% \begin{eqnarray}
+% 	\frac{dV_{m}}{dt} &=& I_{m}+F_{w,m} - F_{m,w} - C\\
+% 	\frac{dV_{w}}{dt} &=& I_{w}+F_{m,w} - F_{w,m} - B
+% \end{eqnarray}where $I_{m},~I_{w}$ denote the rate of liquid input into the main and wound compartment.
 
@@
 
-\subsection*{Complement model analysis.}
 \subsubsection*{Global sensitivity analysis.}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Fig1_Schematic_v2.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Fig-1_Schematic_v2.pdf}
 \caption{Simplified schematic of the human complement system. The complement cascade is activated through three pathways:  the classical, the lectin, and the alternate pathways.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final_v2.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Fig-2_Fits_final_v2.pdf}
 \caption{Reduced order complement model training.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure3_predictions_final_v3.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Fig-3_predictions_final_v3.pdf}
 \caption{Reduced order complement model predictions.
@@
 \centering
-\includegraphics[width=1.1\textwidth]{./figs/Figure4_Sensitivity_v3.pdf}
+%\includegraphics[width=1.1\textwidth]{./figs/Fig-4_Sensitivity_v3.pdf}
 \caption{Global sensitivity analysis of the reduced order complement model.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure5_PairwiseSensitivity_v1.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Fig-5_PairwiseSensitivity_v1.pdf}
 \caption{Pairwise sensitivity and clustering of complement model parameters in the presence of 1 mg/ml zymosan.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure6_RobustnessAnalysis_v4.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Fig-6_RobustnessAnalysis_v4.pdf}
 \caption{Robustness analysis of the complement model.
